---
title: "Cobalt corrosion Github codebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Table of contents

1) Mass balance recoveries, ICP-MS data for background Co(2+) and abiotic phenazine treatments, and pH data tied to growth assays
- Figure 1A: Cobalt corrosion data for wild-type cells, deletion mutants, spent medium, and abiotic controls in sterile medium
- Figure 1B: Boxplot of log10 [Co(2+)] at 168 hours across abiotic and biotic treatments
- Table S1 (partial): Mass balance recoveries and wire losses
- Figure S1: Contribution of background Co(2+) in medium from "no wire" controls
- Figure S2: Boxplot of [Co(2+)] at 168 hours for abiotic metabolite treatments
-	Figure S3: pH changes over time in sterile and live cell treatments

2) Biosynthetic gene cluster analyses
- Bioinformatics analyses of genome annotations generated with DRAM (used to make Figure S4, Table S2, and Table S3)

3) Metabolomic analyses
- The code underlying principal component analyses in Figure S5 is in an independent R codebook
- Volcano plots in in Figure S6 and boxplots in Figure S7-S8 were saved directly from the FBMN software

4) Cell-associated Co(2+) and toxicity measurements
- Figure 2A: Wild type vs deletion cell growth in corrosion assays as measured by viable plate counts
- Figure 2B: Cobalt dose-response curve reporting growth rate of P. chlororaphis subsp. aureofaciens in medium amended with CoCl2
- Figure 2C: Stacked barplot of cobalt localization in dose-response experiments
-	Figure S9: Log(µmol Co(2+)/CFU) over Co(2+) exposure from spike concentrations in corrosion assays compared to dose-response experiments (processed in Excel)
-	Figure S10: Boxplot of [Co(2+)] between A-F and F-A preservation for wild-type cells

5) Additional data for anoxic treatments
- Figure 3: Boxplot of [Co(2+)] at 168 hours for oxic vs anoxic wild-type live cells, wild-type spent medium, and sterile medium (with and without supplied nitrate)
-	Figure S11: Comparison of SEM images of oxic vs anoxic wires after 168 hours (image files)
-	Figure S12: Boxplot of [Co(2+)] at 168 hours for spent medium oxic vs anoxic treatments
-	Figure S13: Boxplot of [Co(2+)] at 168 hours for all corrosion assay treatments
-	Figure S14: Wild type live cell growth over 168 hours in anoxic conditions
-	Figure S15: Nitrate measurements over 168 hours in abiotic and biotic anoxic treatments (processed in Excel)
**Analysis for genetic basis of DET
- Example bash code used to assemble long read data using Nanophase
- Example bash code used to anontate genomes with DRAM
- Fig S4: DET corrosion gene annotation

6) Wire diameters, EDX data, and SEM of non-rinsed wires (not included in this codebook)
- These are largely image files used in Figure 5, Table S4, Table S5, and Figure S16 that were not analyzed in R

7) Passivation layer characterization by PXRD (not included in this codebook)
-	Figure 6A: Geochemical modelling (code can be found on dedicated Github repository referenced in manuscript)
- Figure 6B: PXRD data for wires post-168 hour corrosion assays (processed in Excel)
- Figures S17-S19 were taken directly from the PXRD instrument, not analyzed in R

8) Open circuit potential and Co(2+) concentrations for electrochemical assays
- Figure 7: Cyclic voltammetry for abiotic and biotic treatments over 72 hours
-	Figure S20: Open circuit potential for abiotic and biotic treatments over 72 hours
-	Figure S21: [Co(2+)] over time in abiotic and biotic treatments over 72 hours

9) Thermodynamic calculations (not included in this codebook)
- These are theoretical calculations summarized in Table S6 were not performed in R

10) Genetic manipulations and validation
- Examples of nanophase code used in genome assembly are included here (see the DET part of section 5)
- All other figures (S22-S28) and alignment files were performed outside of R through other open source software

11) LC-HRMS QA/QC and phenazine quantification
- Example chromatograms, m/z values, and retention times that provided calibration curves were processed manually rather than in R (Figure S29 and Tables S7-S8)
-	Figure S30: Stacked barplots for phenazine measurements via LC-HRMS


# Packages required

Load all of these packages before running any of the chunks of code in the subsections below.
```{r}
library(tidyverse)
library(ggplot2)
library(readxl)
library(data.table)
library(plotly)
library(Cairo)
library(lmtest)
library(car)
library(wesanderson)
library(ggbreak)
library(scales)
library(ggh4x)
library(ggtext)
library(drc)
library(RColorBrewer)
library(forcats)
library(markdown)
library(patchwork)
```


# Figure 1A: Cobalt corrosion data for wild-type cells, deletion mutants, spent medium, and abiotic controls in sterile medium

1. Use the code below to load in source data.
```{r}
#Load the data frame from scratch to test
All_cobalt_data_combined_revised <- fread("All_cobalt_ICPMS_data_cleaned_formatted_revised_mass_balance_revised_20260615.txt", sep = "\t", header = "auto")
```

2. Summarize the mass balance data.
```{r}
#Summarize the mass balance data
Mass_balance_data_summary <- All_cobalt_data_combined_revised %>%
                             filter(Wire_treatment == "Wire included") %>% #only include wire treatments
                             group_by(Abiotic_vs_biotic, Wire_treatment, DMSO_treatment, Metabolite_treatment, Sample_preservation, Time) %>%
                             mutate(Mean_recovery = mean(Recovery),
                                    Median_recovery = median(Recovery),
                                    Stdev_recovery = sd(Recovery),
                                    Min_recovery = min(Recovery),
                                    Max_recovery = max(Recovery)) %>%
                             filter(Wire_treatment == "Wire included" & Time == 168 & Replicate == 1) %>%
                             dplyr::select(c(Abiotic_vs_biotic, DMSO_treatment, Metabolite_treatment, Sample_preservation, Mean_recovery:Max_recovery))

write.table(Mass_balance_data_summary, file = "TableS1_Mass_balance_data_summary_20260820.txt", sep = "\t", 
            row.names = FALSE, col.names = TRUE) 

Mass_balance_data_summary_count <- All_cobalt_data_combined_revised %>%
                                   count(Abiotic_vs_biotic, Wire_treatment, DMSO_treatment, Metabolite_treatment, Sample_preservation, Time)
```

3. Use the graphing script and call the *corrected* cobalt concentration in this case and save it as new dedicated files.
```{r}
#Adjust the presentation aesthetics
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
        size = 34, angle = 0, hjust = .5, vjust = 0, face = "bold"),
        axis.title.y = element_text(color = "grey20", 
        size = 34, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 32, colour = "grey20", face = "bold"),
        strip.text.y = element_text(size = 32, colour = "grey20"),
        plot.title = element_text(size = 30, face = "bold", hjust = .5),
        legend.title=element_text(size=32, face = "bold"), 
        legend.text=element_text(size=32),
        panel.border = element_rect(colour = "black", fill=NA, size=1))

dark2_colors <- brewer.pal(n = 8, name = "Dark2") #make palette with desired levels


#Create custom display orders
All_cobalt_data_combined_revised$Abiotic_vs_biotic_ordered <- factor(All_cobalt_data_combined_revised$Abiotic_vs_biotic,
                                                             levels = c("Wild type",
                                                                        "∆phz",
                                                                        "∆hcn",
                                                                        "∆phz∆hcn",
                                                                        "Spent medium",
                                                                        "Spent medium from ∆phz",
                                                                        "Heat kill", 
                                                                        "Sterile medium"))

All_cobalt_data_combined_revised$Metabolite_treatment_ordered <- factor(All_cobalt_data_combined_revised$Metabolite_treatment,
                                                                 levels = c("Nothing added",
                                                                            "PCA",
                                                                            "MeOH"))

Abiotic_vs_biotic_treatment_scatterplot_recoveries <- All_cobalt_data_combined_revised %>%
                                           filter(Wire_treatment == "Wire included") %>% #filter to wire only experiments
                                           filter(DMSO_treatment == "No DMSO") %>% #use this to only focus on experiments that did not receive DMSO
                                           filter(Sample_preservation == "Acidified, filtered") %>% #216 observations
                                           filter(!Abiotic_vs_biotic == "Sterile medium, anoxic",  !Abiotic_vs_biotic == "Spent medium from ∆phz, anoxic", !Abiotic_vs_biotic == "Sterile medium, anoxic, nitrate", !Abiotic_vs_biotic == "Spent medium from WT, anoxic", !Abiotic_vs_biotic == "Spent medium from WT, anoxic, nitrate", !Abiotic_vs_biotic == "Wild type, anoxic", !Abiotic_vs_biotic == "Wild type, anoxic, nitrate") %>%
                                           filter(Metabolite_treatment_ordered == "Nothing added") %>% #remove all of the other metabolite treatments

                                          #plotting script
            
                                           ggplot(aes(x=Time, 
                                                      y = Mean_cobalt_concentration_uM_corrected, 
                                                      fill = Abiotic_vs_biotic_ordered)) +
                                                  geom_errorbar(aes(ymin = Mean_cobalt_concentration_uM_corrected - Stdev_cobalt_concentration_uM_corrected,
                                                                    ymax = Mean_cobalt_concentration_uM_corrected + Stdev_cobalt_concentration_uM_corrected,
                                                                    colour = Abiotic_vs_biotic_ordered)) +
                                                  geom_point(shape = 21, color = "black", stroke = 0.8, size = 6) +
                                                  geom_line(aes(group = Bottle_number, color = Abiotic_vs_biotic_ordered))+
                                                  scale_x_continuous(breaks = seq(0, 168, by = 24)) +
                                                  ylim(0, 6500) +
                                                  facet_grid(~ Cell_presence,
                                                             labeller = labeller(Cell_presence = 
                                                                                   c("No cells" = "Abiotic",
                                                                                     "Live cells" = "Biotic"))) +
                                                  ylab("Co(2+) concentration (µM)\nrecoveries applied") +
                                                  xlab("Time elapsed (hours)") + 
                                                  scale_fill_manual(values = dark2_colors,
                                                                    name = "Treatment", 
                                                                    labels = c("Wild type", 
                                                                             expression(italic("∆phz")),
                                                                             expression(italic("∆hcn")),
                                                                             expression(italic("∆phz∆hcn")),
                                                                             "Spent medium",
                                                                           expression(plain("Spent medium ") * italic("∆phz")),
                                                                           "Heat kill", 
                                                                             "Sterile medium"))+
  
                                                     scale_colour_manual(values = dark2_colors,
                                                                    name = "Treatment", 
                                                                    labels = c("Wild type", 
                                                                             expression(italic("∆phz")),
                                                                             expression(italic("∆hcn")),
                                                                             expression(italic("∆phz∆hcn")),
                                                                             "Spent medium",
                                                                             expression(plain("Spent medium ") * italic("∆phz")),
                                                                             "Heat kill", 
                                                                             "Sterile medium"))+
                                                  theme +
                                                  theme(plot.margin = unit(c(1,1,1,1), "cm"),
                                                        axis.title.x = element_text(vjust = -2),
                                                        axis.title.y = element_text(vjust = 2))
          
#Add a layer to angle the text
Abiotic_vs_biotic_treatment_scatterplot_angled_recoveries <- Abiotic_vs_biotic_treatment_scatterplot_recoveries + theme(axis.text.x = element_text(angle = 45))

#Save in different image formats at 300 dpi
ggsave("Fig1A_Cobalt_ICPMS_data_abiotic_biotic_and_metabolite_treatments_no_DMSO_scatterplot_angled_recoveries_applied_20260820.jpg", 
       dpi = 300, height=10, width=20, device= jpeg)
ggsave("Fig1A_Cobalt_ICPMS_data_abiotic_biotic_and_metabolite_treatments_no_DMSO_scatterplot_angled_recoveries_applied_20260820.tiff", 
       dpi = 300, height=10, width=20, device= tiff)
ggsave("Fig1A_Cobalt_ICPMS_data_abiotic_biotic_and_metabolite_treatments_no_DMSO_scatterplot_angled_recoveries_applied_20260820.pdf", 
       dpi = 300, height=8, width=22, device= cairo_pdf)
```


# Figure 1B: Boxplot of log10 [Co(2+)] at 168 hours across abiotic and biotic treatments

1. Load in the data from the manually adjusted tab-delimited file that contains recoveries. Generate a new variable where the recovery has been applied.
```{r}
All_cobalt_data_combined_revised_recoveries <- fread("All_cobalt_ICPMS_data_cleaned_formatted_revised_mass_balance_revised_20260615.txt", sep = "\t", header = "auto")
```

2. Run the model applying the same outlier removal and log-transformations used on the data that was not corrected for recovery.
```{r}
Abiotic_vs_biotic_wire_only_ANOVA_data_outlier_removed_recoveries <- All_cobalt_data_combined_revised_recoveries %>%
                                                                     filter(DMSO_treatment == "No DMSO") %>% 
                                                                     filter(Sample_preservation == "Acidified, filtered") %>% 
                                                                     filter(Time == 168) %>%
                                                                     filter(Metabolite_treatment == "Nothing added") %>%
                                                                     filter(Wire_treatment == "Wire included") %>%
                                                                     filter(!Abiotic_vs_biotic == "Sterile medium, anoxic",  !Abiotic_vs_biotic == "Spent medium from ∆phz, anoxic", !Abiotic_vs_biotic == "Sterile medium, anoxic, nitrate", !Abiotic_vs_biotic == "Spent medium from WT, anoxic", !Abiotic_vs_biotic == "Spent medium from WT, anoxic, nitrate", !Abiotic_vs_biotic == "Wild type, anoxic", !Abiotic_vs_biotic == "Wild type, anoxic, nitrate") #remove anoxic

##Generate model
Abiotic_vs_biotic_model_log10_outlier_removed_recoveries <- lm(log10(Cobalt_concentration_uM_corrected) ~ Abiotic_vs_biotic, 
                                                            data = Abiotic_vs_biotic_wire_only_ANOVA_data_outlier_removed_recoveries, 
                                                            contrasts = list(Abiotic_vs_biotic = contr.sum), type =3)

Abiotic_vs_biotic_model_log10_outlier_removed_recoveries
par(mfrow = c(2,2))
plot(Abiotic_vs_biotic_model_log10_outlier_removed_recoveries)

#Run Breusch Pagan test for homoscedasticity
bptest(Abiotic_vs_biotic_model_log10_outlier_removed_recoveries)

#Run Shapiro-Wilk normality test
shapiro.test(Abiotic_vs_biotic_model_log10_outlier_removed_recoveries$residuals)

#Run Durbin-Watson test for autocorrelation
dwtest(Abiotic_vs_biotic_model_log10_outlier_removed_recoveries)
```

## Figure 1B ANOVA Interpretation
- Homoscedasticity looks okay as does the normal distribution of residuals.
- BP test returns a p-value of 0.2491, so we cannot reject the null hypothesis of homoscedasticity.
- SW test returns a p-value of 0.5285, so we cannot reject the null hypothesis that residuals are normally distributed.
- DW test returns a p-value of 0.129, so we cannot reject the null hypothesis that there is no autocorrelation between data points.

**All linear model assumptions are met**

3. Generate the actual output of anova model generated with the data where the outlier removed and a log10 transformation was applied. This is returning a type 3 sum of squares because we are not longer considering the interaction having controlled for wire treatment
```{r}
Abiotic_vs_biotic_type_III_ANOVA_results_recoveries <- Anova(Abiotic_vs_biotic_model_log10_outlier_removed_recoveries, type = 3)
Abiotic_vs_biotic_type_III_ANOVA_results_recoveries
```
- F-value = 84.124
- p-value = < 2.2e-16 ***

- There are significant differences in cobalt concentration at 168 hours that are attributable to the different abiotic vs biotic treatments even after correcting for recoveries.
- The p-values are different for the recovery corrected data, but both version of the data have p < 0.001.

4. Run a Tukey post-hoc test to figure out where the significant differences are.
```{r}
#First you need to convert the model to an aov object
Abiotic_vs_biotic_model_log10_outlier_removed_converted_recoveries <- aov(Abiotic_vs_biotic_model_log10_outlier_removed_recoveries)
Tukey_comparison_abiotic_vs_biotic_model_log10_outlier_removed_recoveries <- TukeyHSD(Abiotic_vs_biotic_model_log10_outlier_removed_converted_recoveries, conf.level = 0.95)
Tukey_comparison_abiotic_vs_biotic_model_log10_outlier_removed_recoveries
Tukey_comparison_abiotic_vs_biotic_data_frame_recoveries <- as.data.frame(Tukey_comparison_abiotic_vs_biotic_model_log10_outlier_removed_recoveries$Abiotic_vs_biotic)

letters_AvB <- multcompLetters4(Abiotic_vs_biotic_model_log10_outlier_removed_converted_recoveries,
                                Tukey_comparison_abiotic_vs_biotic_model_log10_outlier_removed_recoveries)
letters_AvB
```

## Figure 1B multiple comparison interpretation

**Significant results**
                                              diff         lwr          upr     p adj
∆phz∆hcn-∆hcn                          0.576930931  0.31042746  0.843434402 0.0000023
Heat kill-∆hcn                        -0.688168696 -0.97602535 -0.400312046 0.0000004
Spent medium-∆hcn                     -0.370403748 -0.63690722 -0.103900277 0.0020801
Sterile medium-∆hcn                   -0.768244888 -0.99472929 -0.541760488 0.0000000
∆phz∆hcn-∆phz                          0.378550119  0.11204665  0.645053591 0.0015955
Heat kill-∆phz                        -0.886549507 -1.17440616 -0.598692857 0.0000000
Spent medium-∆phz                     -0.568784560 -0.83528803 -0.302281088 0.0000030
Spent medium from ∆phz-∆phz           -0.297767620 -0.58562427 -0.009910971 0.0386296
Sterile medium-∆phz                   -0.966625700 -1.19311010 -0.740141299 0.0000000
Heat kill-∆phz∆hcn                    -1.265099626 -1.55295628 -0.977242977 0.0000000
Spent medium-∆phz∆hcn                 -0.947334679 -1.21383815 -0.680831207 0.0000000
Spent medium from ∆phz-∆phz∆hcn       -0.676317740 -0.96417439 -0.388461090 0.0000005
Sterile medium-∆phz∆hcn               -1.345175819 -1.57166022 -1.118691418 0.0000000
Wild type-∆phz∆hcn                    -0.388218663 -0.63150194 -0.144935391 0.0003423
Spent medium-Heat kill                 0.317764947  0.02990830  0.605621597 0.0225589
Spent medium from ∆phz-Heat kill       0.588781887  0.28105018  0.896513589 0.0000207
Wild type-Heat kill                    0.876880963  0.61037749  1.143384435 0.0000000
Sterile medium-Spent medium           -0.397841140 -0.62432554 -0.171356739 0.0000825
Wild type-Spent medium                 0.559116016  0.31583274  0.802399288 0.0000008
Sterile medium-Spent medium from ∆phz -0.668858079 -0.92011996 -0.417596196 0.0000000
Wild type-Spent medium from ∆phz       0.288099077  0.02159560  0.554602549 0.0269924
Wild type-Sterile medium               0.956957156  0.75831720  1.155597116 0.0000000

**Non-significant results**
∆phz-∆hcn                              0.198380811 -0.06812266  0.464884283 0.2659853
Spent medium from ∆phz-∆hcn           -0.099386809 -0.38724346  0.188469840 0.9455226
Wild type-∆hcn                         0.188712267 -0.05457100  0.431995540 0.2229459
Wild type-∆phz                        -0.009668544 -0.25295182  0.233614728 1.0000000
Sterile medium-Heat kill              -0.080076193 -0.33133808  0.171185690 0.9640108
Spent medium from ∆phz-Spent medium    0.271016939 -0.01683971  0.558873589 0.0763504

### Letters to add to Figure 1B

- Wild type = a
- ∆phz = a
- ∆hcn = ad
- ∆phz∆hcn = b
- Spent medium = c
- Spent medium from ∆phz = cd
- Heat kill = e
- Sterile medium = e

5. Generate boxplots that will be used to overlay the Tukey annotations manually.
```{r}
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 16, angle = 45, hjust = 0.5, vjust = 0.55, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 16, angle = 0, hjust = 0.5, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 18, angle = 0, hjust = 0.5, vjust = -1, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 18, angle = 90, face = "plain"),
        strip.text.x = element_text(size = 20, colour = "grey20"),
        strip.text.y = element_text(size = 20, colour = "grey20"),
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5, vjust = 1),
        legend.title=element_text(size=18), 
        legend.text=element_text(size=18),
        legend.position = "right",
        panel.border = element_rect(colour = "black", fill=NA, size=1))


#Create custom order to display x-axis

All_cobalt_data_combined_revised_recoveries$Abiotic_vs_biotic_ordered <- factor(All_cobalt_data_combined_revised_recoveries$Abiotic_vs_biotic,
                                                                 levels = c("Wild type",
                                                                        "∆phz",
                                                                        "∆hcn",
                                                                        "∆phz∆hcn",
                                                                        "Spent medium",
                                                                        "Spent medium from ∆phz",
                                                                        "Heat kill", 
                                                                        "Sterile medium"))

Abiotic_vs_biotic_wire_only_boxplot_log10_recoveries <- All_cobalt_data_combined_revised_recoveries %>%
                                                        filter(DMSO_treatment == "No DMSO") %>% 
                                                        filter(Sample_preservation == "Acidified, filtered") %>% 
                                                        filter(Time == 168) %>%
                                                        filter(Metabolite_treatment == "Nothing added") %>%
                                                        filter(Wire_treatment == "Wire included") %>%
                                                        filter(!Abiotic_vs_biotic == "Sterile medium, anoxic",  !Abiotic_vs_biotic == "Spent medium from ∆phz, anoxic", !Abiotic_vs_biotic == "Sterile medium, anoxic, nitrate", !Abiotic_vs_biotic == "Spent medium from WT, anoxic", !Abiotic_vs_biotic == "Spent medium from WT, anoxic, nitrate", !Abiotic_vs_biotic == "Wild type, anoxic", !Abiotic_vs_biotic == "Wild type, anoxic, nitrate") %>%
                                                        mutate(Abiotic_vs_biotic_ordered = fct_recode(Abiotic_vs_biotic_ordered, "Spent medium\nfrom ∆phz" = "Spent medium from ∆phz")) 

#Create vector of letters in order to add to boxplot:
letters_vec_Fig1 <- c("a","a","ad", "b", "c", "cd", "e", "e")

#Make dataframe to integrate into plot:
annotation_df_Fig1 <- data.frame(Abiotic_vs_biotic_ordered = 
                       levels(Abiotic_vs_biotic_wire_only_boxplot_log10_recoveries$Abiotic_vs_biotic_ordered),
                       letter = letters_vec_Fig1)

Abiotic_vs_biotic_boxplot <- Abiotic_vs_biotic_wire_only_boxplot_log10_recoveries%>%
                                                        #Graphing script
                                                        ggplot(aes(x=Abiotic_vs_biotic_ordered, y=log10(Cobalt_concentration_uM_corrected))) + 
                                                        geom_boxplot() +
                                                        geom_jitter(fill = "grey", colour = "black", width = 0.1, pch =21, stroke =1)+
                                                        #ylim(2, 4)+
                                                        coord_cartesian(clip = "off") +
                                                        geom_text(data = annotation_df_Fig1,
                                                        aes(x = Abiotic_vs_biotic_ordered, y = Inf, label = letter),
                                                        vjust = -1.5,
                                                        size = 12, face = "bold", 
                                                        inherit.aes = FALSE) +
                                                        labs(title = "" ,
                                                             y = "log10[Co(2+) concentration (µM)]\ncorrected for recovery", 
                                                             x = "Abiotic and biotic treatments") +
                                                        theme + #this calls the aesthetic theme variable
                                                        theme(plot.margin = margin(t = 40, r = 10, b = 10, l = 10),,
                                                              axis.title.x = element_text(vjust = -2),
                                                              axis.title.y = element_text(vjust = 2))

Abiotic_vs_biotic_boxplot
ggsave("Fig1B_Cobalt_ICPMS_data_biotic_treatments_no_DMSO_boxplot_log10_recoveries_applied_20260820.jpg", dpi = 300, height=9, width=12, device=jpeg)
ggsave("Fig1B_Cobalt_ICPMS_data_biotic_treatments_no_DMSO_boxplot_log10_recoveries_applied_20260820.tiff", dpi = 300, height=9, width=12, device=tiff)
ggsave("Fig1B_Cobalt_ICPMS_data_biotic_treatments_no_DMSO_boxplot_log10_recoveries_applied_20260820.pdf", dpi = 300, height=9, width=12, device=cairo_pdf)
```

## Combining Figures 1A and 1B

1. Replot 1B with the same colour coding as 1A.
```{r}
dark2_colors <- brewer.pal(n = 8, name = "Dark2") #make palette with desired levels

theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
        size = 32, angle = 45, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
        size = 34, angle = 0, hjust = .5, vjust = 0, face = "bold"),
        axis.title.y = element_text(color = "grey20", 
        size = 34, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 32, colour = "grey20", face = "bold"),
        strip.text.y = element_text(size = 32, colour = "grey20"),
        plot.title = element_text(size = 30, face = "bold", hjust = .5),
        legend.title=element_text(size=32, face = "bold"), 
        legend.text=element_text(size=32),
        panel.border = element_rect(colour = "black", fill=NA, size=1))

Abiotic_vs_biotic_wire_only_boxplot_log10_recoveries_colour_coded <- All_cobalt_data_combined_revised_recoveries %>%
                                                        filter(DMSO_treatment == "No DMSO") %>% 
                                                        filter(Sample_preservation == "Acidified, filtered") %>% 
                                                        filter(Time == 168) %>%
                                                        filter(Metabolite_treatment == "Nothing added") %>%
                                                        filter(Wire_treatment == "Wire included") %>%
                                                        filter(!Abiotic_vs_biotic == "Sterile medium, anoxic",  !Abiotic_vs_biotic == "Spent medium from ∆phz, anoxic", !Abiotic_vs_biotic == "Sterile medium, anoxic, nitrate", !Abiotic_vs_biotic == "Spent medium from WT, anoxic", !Abiotic_vs_biotic == "Spent medium from WT, anoxic, nitrate", !Abiotic_vs_biotic == "Wild type, anoxic", !Abiotic_vs_biotic == "Wild type, anoxic, nitrate") %>%
                                                        mutate(Abiotic_vs_biotic_ordered = fct_recode(Abiotic_vs_biotic_ordered, 
                                                                                                  "Spent medium\nfrom ∆phz" = "Spent medium from ∆phz")) %>%
                                                        #Graphing script
                                                        ggplot(aes(x=Abiotic_vs_biotic_ordered, y=log10(Cobalt_concentration_uM_corrected), fill = Abiotic_vs_biotic_ordered)) + 
                                                        geom_boxplot() +
                                                        geom_jitter(fill = "grey", colour = "black", width = 0.1, pch =21, stroke =1)+
                                                        scale_fill_manual(values = dark2_colors) +
                                                        scale_x_discrete(labels = c("Wild type" = "Wild type",
                                                                                    "∆phz" = "<i>∆phz</i>",
                                                                                    "∆hcn" = "<i>∆hcn</i>",
                                                                                    "∆phz∆hcn" = "<i>∆phz∆hcn</i>",
                                                                                    "Spent medium" = "Spent medium",
                                                                                    "Spent medium\nfrom ∆phz" = "Spent medium<br>from <i>∆phz</i>",
                                                                                    "Heat kill" = "Heat kill",
                                                                                    "Sterile medium" = "Sterile medium")) +
                                                        #ylim(2, 4)+
                                                        coord_cartesian(clip = "off") +
                                                        geom_text(data = annotation_df_Fig1,
                                                        aes(x = Abiotic_vs_biotic_ordered, y = Inf, label = letter),
                                                        vjust = -0.5,
                                                        size = 10, fontface = "bold", 
                                                        inherit.aes = FALSE) +
                                                        labs(title = "" ,
                                                             y = "log10[Co(2+) concentration (µM)]\nrecoveries applied", 
                                                             x = "Treatment") +
                                                        theme + #this calls the aesthetic theme variable
                                                        theme(plot.margin = margin(t = 105, r = 40, b = 18, l = 10),
                                                              axis.title.x = element_text(vjust = -2),
                                                              axis.title.y = element_text(vjust = 2),
                                                              axis.text.x = ggtext::element_markdown(angle = 55,
                                                                                                     size = 32,
                                                                                                     colour = "grey20"),
                                                              legend.position = "none",
                                                              guides(fill = "none"))

Abiotic_vs_biotic_wire_only_boxplot_log10_recoveries_colour_coded #annotate this manually in an image editor
ggsave("Fig1B_Cobalt_ICPMS_data_biotic_treatments_no_DMSO_boxplot_log10_recoveries_applied_colour_coded_20260820.jpg", dpi = 300, height=9, width=12, device=jpeg)
ggsave("Fig1B_Cobalt_ICPMS_data_biotic_treatments_no_DMSO_boxplot_log10_recoveries_applied_colour_coded_20260820.tiff", dpi = 300, height=9, width=12, device=tiff)
ggsave("Fig1B_Cobalt_ICPMS_data_biotic_treatments_no_DMSO_boxplot_log10_recoveries_applied_colour_coded_20260820.pdf", dpi = 300, height=9, width=12, device=cairo_pdf)
```

2. Use the patchwork package to combine, adjust dimensions, and aesthetics.
```{r}
Fig1A_Fig1B_combined <- (Abiotic_vs_biotic_treatment_scatterplot_angled_recoveries / Abiotic_vs_biotic_wire_only_boxplot_log10_recoveries_colour_coded) + 
                         plot_layout(ncol = 1)

Fig1A_Fig1B_combined

ggsave("Fig1AB_combined_colour_coded_20260820.jpg", dpi = 300, height=16, width=20, device=jpeg)
ggsave("Fig1AB_combined_colour_coded_20260820.tiff", dpi = 300, height=16, width=20, device=tiff)
ggsave("Fig1AB_combined_colour_coded_20260820.pdf", dpi = 300, height=16, width=20, device=cairo_pdf)
```

# Figure S1: Contribution of background Co(2+) in medium from "no wire" controls

1. Reuse the combined data frame but filter it only to the no wire treatments. Facet the data so each Abiotic_vs_biotic treatment ordered is in its own panel, preserve the colour coding but eliminate the legend, and fix the y-axis to the same range as the main figure.
```{r}
All_cobalt_data_combined_revised <- fread("All_cobalt_ICPMS_data_cleaned_formatted_revised_mass_balance_revised_20260615.txt", sep = "\t", header = "auto")

#Adjust the presentation aesthetics
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
        size = 34, angle = 0, hjust = .5, vjust = 0, face = "bold"),
        axis.title.y = element_text(color = "grey20", 
        size = 34, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 32, colour = "grey20", face = "bold"),
        strip.text.y = element_text(size = 32, colour = "grey20"),
        plot.title = element_text(size = 30, face = "bold", hjust = .5),
        legend.title=element_text(size=32, face = "bold"), 
        legend.text=element_text(size=32),
        panel.border = element_rect(colour = "black", fill=NA, size=1))


#Create custom display orders
#All_cobalt_data_combined_revised$Abiotic_vs_biotic_ordered <- factor(All_cobalt_data_combined_revised$Abiotic_vs_biotic,
                                                             #levels = c("Wild type",
                                                                        #"∆phz",
                                                                        #"∆hcn",
                                                                        #"Spent medium",
                                                                        #"Spent medium from ∆phz",
                                                                        #"Sterile medium"))


All_cobalt_data_combined_revised$Metabolite_treatment_ordered <- factor(All_cobalt_data_combined_revised$Metabolite_treatment,
                                                                 levels = c("Nothing added",
                                                                            "PCA",
                                                                            "MeOH"))
No_wire_scatterplot <- All_cobalt_data_combined_revised %>%
                       filter(Wire_treatment == "No wire") %>% #filter to no wire experiments
                       filter(DMSO_treatment == "No DMSO") %>% #use this to only focus on experiments that did not receive DMSO
                       filter(Sample_preservation == "Acidified, filtered") %>%
                       filter(!Abiotic_vs_biotic == "Sterile medium, anoxic",  !Abiotic_vs_biotic == "Spent medium from ∆phz, anoxic", !Abiotic_vs_biotic == "Sterile medium, anoxic, nitrate", !Abiotic_vs_biotic == "Spent medium from WT, anoxic", !Abiotic_vs_biotic == "Spent medium from WT, anoxic, nitrate", !Abiotic_vs_biotic == "Wild type, anoxic", !Abiotic_vs_biotic == "Wild type, anoxic, nitrate", !Abiotic_vs_biotic == "∆phz∆hcn", !Abiotic_vs_biotic == "Heat kill") %>%
                       group_by(Abiotic_vs_biotic, Wire_treatment, DMSO_treatment, Metabolite_treatment, Time) %>%
                       filter(Metabolite_treatment_ordered == "Nothing added") %>% #remove all of the other metabolite treatments
                       mutate(Abiotic_vs_biotic = fct_recode(Abiotic_vs_biotic, "Spent medium\nfrom ∆phz" = "Spent medium from ∆phz")) %>%
  mutate(facet_label = dplyr::recode(as.character(Abiotic_vs_biotic),
      "Wild type" = "Wild type",
      "∆phz" = "<i>∆phz</i>",
      "∆hcn" = "<i>∆hcn</i>",
      "Spent medium" = "Spent medium",
      "Spent medium\nfrom ∆phz" = "Spent medium<br>from <i>∆phz</i>",
      "Sterile medium" = "Sterile medium"), 
      facet_label = factor(facet_label,levels = c("Wild type",
                                                  "<i>∆phz</i>",
                                                  "<i>∆hcn</i>",
                                                  "Spent medium",
                                                  "Spent medium<br>from <i>∆phz</i>",
                                                  "Sterile medium"))) %>% 
                      #plotting script
                      ggplot(aes(x=Time, y = Mean_cobalt_concentration_uM_corrected)) +
                      geom_errorbar(aes(ymin = Mean_cobalt_concentration_uM_corrected - Stdev_cobalt_concentration_uM_corrected,
                                        ymax = Mean_cobalt_concentration_uM_corrected + Stdev_cobalt_concentration_uM_corrected)) +
                      geom_point(shape = 21, color = "black", stroke = 0.8, size = 6) +
                      geom_line(aes(group = Bottle_number))+
                      geom_hline(yintercept = 1.68, linetype = "dashed", color = "orange", size = 1) +
                      scale_x_continuous(breaks = seq(0, 168, by = 24)) +
                      #ylim(0, 6500) +
                      facet_grid(~ facet_label, drop = FALSE) +
                      ylab("Co(2+) concentration (µM)\n recoveries applied") +
                      xlab("Time elapsed (hours)") +
                      theme +
                      theme(plot.margin = unit(c(1,1,1,1), "cm"),
                            axis.title.x = element_text(vjust = -2),
                            axis.title.y = element_text(vjust = 2),
                            legend.position = "none", 
                            strip.text.x = ggtext::element_markdown(size = 32,face = "bold"))
          
#Add a layer to angle the text
No_wire_scatterplot_angled <- No_wire_scatterplot + theme(axis.text.x = element_text(angle = 45))

#Save in different image formats at 300 dpi
ggsave("FigS1_Cobalt_ICPMS_data_no_wire_treaments_scatterplot_angled_20260820.jpg", dpi = 300, height=8, width=34, device= jpeg)
ggsave("FigS1_Cobalt_ICPMS_data_no_wire_treaments_scatterplot_angled_20260820.tiff", dpi = 300, height=8, width=34, device= tiff)
ggsave("FigS1_Cobalt_ICPMS_data_no_wire_treaments_scatterplot_angled_20260820.pdf", dpi = 300, height=8, width=34, device= cairo_pdf)
```


# Figure S2: Boxplot of [Co(2+)] at 168 hours for abiotic metabolite treatments

1. Load the data frame from the compiled data.
```{r}
All_cobalt_data_combined_revised <- fread("All_cobalt_ICPMS_data_cleaned_formatted_revised_mass_balance_revised_20260615.txt", sep = "\t", header = "auto")

All_cobalt_data_combined_revised$Metabolite_treatment_ordered <- factor(All_cobalt_data_combined_revised$Metabolite_treatment,
                                                                 levels = c("Nothing added",
                                                                            "PCA",
                                                                            "MeOH"))
Abiotic_no_outlier <- All_cobalt_data_combined_revised %>%
                      filter(DMSO_treatment == "No DMSO") %>% #use this to only focus on experiments that did not receive DMSO
                      filter(Sample_preservation == "Acidified, filtered") %>% #216 observations
                      filter(Time == 168 & Wire_treatment == "Wire included") %>% #Only look at no metabolite spikes + 168 hours
                      filter(Abiotic_vs_biotic == "Sterile medium")
```

2. Generate the linear models and verify assumptions for the ANOVA using type III sum of squares.
```{r}
##Generate model
Abiotic_model_outlier_removed <- lm(Cobalt_concentration_uM_corrected ~ Metabolite_treatment, 
                                    data = Abiotic_no_outlier, 
                                    contrasts = list(Metabolite_treatment = contr.sum), type =3)

Abiotic_model_outlier_removed
par(mfrow = c(2,2))
plot(Abiotic_model_outlier_removed)

#Run Breusch Pagan test for homoscedasticity
bptest(Abiotic_model_outlier_removed)

#Run Shapiro-Wilk normality test
shapiro.test(Abiotic_model_outlier_removed$residuals)

#Run Durbin-Watson test for autocorrelation
dwtest(Abiotic_model_outlier_removed)
```
## Figure S3 ANOVA Interpretation
- Homoscedasticity may be an issue, it looks like samples associated with higher cobalt concentrations ~ 240 µM have higher variance.
- Normal distribution of residuals looks good.
- BP test returns a p-value of *0.268* --> we cannot reject the null hypothesis of homoscedasticity.
- SW test returns a p-value of *0.4308* --> we cannot reject the null hypothesis of normally distributed residuals.
- DW test returns a p-value of *0.9208* --> we cannot reject the null hypothesis that there is no autocorrelation between data points.

**All assumptions met, okay to proceed with ANOVA**

3. Run the ANOVA and the Tukey comparison.
```{r}
Abiotic_type_III_ANOVA_results <- Anova(Abiotic_model_outlier_removed, type = 3)
Abiotic_type_III_ANOVA_results
```
- Metabolite treatment did not emerge as a significant variable, so there is no significant differences between groups.
- F-value 0.8954
- p-value 0.4341

4. Generate the boxplot. There is no need to annotate the boxes later because there are no significant differences.
```{r}
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 16, angle = 45, hjust = 0.5, vjust = 0.55, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 16, angle = 0, hjust = 0.5, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 18, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
        axis.title.y = element_text(color = "grey20", size = 18, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 20, colour = "grey20"),
        strip.text.y = element_text(size = 20, colour = "grey20"),
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5, vjust = 1),
        legend.title=element_text(size=18), 
        legend.text=element_text(size=18),
        legend.position = "right",
        panel.border = element_rect(colour = "black", fill=NA, size=1))

Abiotic_outlier_removed_boxplot <- Abiotic_no_outlier %>%
                                   ggplot(aes(x=Metabolite_treatment_ordered, y=Cobalt_concentration_uM_corrected)) + 
                                   geom_boxplot() +
                                   geom_jitter(fill = "grey", colour = "black", width = 0.1, pch =21, stroke =1)+
                                   ylim(0, 400) +
                                   labs(title = "" ,
                                   y = "Co(2+) concentration (µM)\nrecoveries applied", 
                                   x = "Metabolic treatments in sterile medium") +
                                   theme +
                                   theme(plot.margin = unit(c(1,1,1,1), "cm"),
                                         axis.title.x = element_text(vjust = -2),
                                         axis.title.y = element_text(vjust = 2))
Abiotic_outlier_removed_boxplot
ggsave("FigS2_Cobalt_ICPMS_data_abiotic_treatments_no_DMSO_boxplot_20260820.jpg", dpi = 300, height=6, width=6, device=jpeg)
ggsave("FigS2_Cobalt_ICPMS_data_abiotic_treatments_no_DMSO_boxplot_20260820.tiff", dpi = 300, height=6, width=6, device=tiff)
ggsave("FigS2_Cobalt_ICPMS_data_abiotic_treatments_no_DMSO_boxplot_20260820.pdf", dpi = 300, height=6, width=6, device=cairo_pdf)
```
#  Biosynthetic gene cluster analyses

These analyses underpin *Figure S4,* *Table S2*, and *Table S3*, which were extracted from the notes in this R notebook.

## Run DRAM for annotations and distillation

1. Activate the conda environment containing DRAM and run the annotate command. Replace the directory and fasta file with the genome file for Pseudomonas chlororaphis subspecies aurefoaciens (or any genome or genoems of your choice)
```{bash}
DRAM.py annotate -i 'Pseudomonas_genome_analyses/*.fasta' --use_uniref -o Pseudomonas_annotation_with_uniref --threads 8
```

2. Once the annotation is complete, use the distillation command. Make sure to use the same output directory from the annotate step.
```{bash}
DRAM.py distill -i Pseudomonas_annotation_with_uniref/annotations.tsv -o Pseudomonas_genome_summaries --trna_path Pseudomonas_annotation_with_uniref/trnas.tsv --rrna_path Pseudomonas_annotation_with_uniref/rrnas.tsv
```

## Table S2 raw data: Manually search annotations for genes of interest corresponding to antiSMASH predictions and potential corrosion pathways

These files were used to manually create **Table S2**.

1. Load in the tab-delimited file containing the annotation data. Make sure to convert the MS Excel file to a flat format, i.e., .tsv manually.
```{r}
Pseudomonas_chlororaphis_aureofaciens_annotations <- fread("Supporting_Data_1B.txt", sep = "\t", header = "auto") %>% slice (-1)
colnames(Pseudomonas_chlororaphis_aureofaciens_annotations) <- as.character(Pseudomonas_chlororaphis_aureofaciens_annotations[1, ])
Pseudomonas_chlororaphis_aureofaciens_annotations <- Pseudomonas_chlororaphis_aureofaciens_annotations[-1,]
```

2. Manually search the annotation file using different strings of test for functions relevant to cobalt and microbially-induced corrosion.
```{r}
corrosion_pathways <- Pseudomonas_chlororaphis_aureofaciens_annotations %>%
                    
                   #Simple text search for "cobalt"
                   filter(
                     (if_any(c(kegg_hit:pfam_hits), ~ grepl("cobalt", .))) |
                     (if_any(c(kegg_hit:pfam_hits), ~ grepl("cob", .))) |
                     (if_any(c(kegg_hit:pfam_hits), ~ grepl("cyanide", .)))|
                     (if_any(c(kegg_hit:pfam_hits), ~ grepl("hcn", .)))|
                     (if_any(c(kegg_hit:pfam_hits), ~ grepl("cytochrome", .)))|
                     #(if_any(c(kegg_hit:pfam_hits), ~ grepl("hydrogenase", .)))| this is non-specific and picks up anything with dehydrogenase, comment out
                     (if_any(c(kegg_hit:pfam_hits), ~ grepl("siderophore", .)))) %>%
                    dplyr::select(scaffold,
                           gene_position,
                           start_position,
                           end_position,
                           ko_id,
                           kegg_hit,
                           uniref_id,
                           uniref_hit)

#Write the table to save the source data
write.table(corrosion_pathways, file = "SuppData_Pseudomonas_chlororaphis_aureofaciens_putative_corrosion_pathways.tsv", sep = "\t", row.names = FALSE)
``` 

*key biosynthetic genes* are denoted by asterisks in markdown format.

## Table S3 raw data: Cross-referencing annotations with potential metabolites

### Region 2 (hydrogen cyanide cluster): 2951483 to 2964455
```{r}
Region2 <- Pseudomonas_chlororaphis_aureofaciens_annotations %>% filter(start_position >= 2951483 & end_position <=2964455)
```

Genes in hydrogen cyanide-cluster:
- *ctg1_2591*: 2956483 - 2956800 --> hydrogen cyanide synthase HcnA [EC:1.4.99.5] (C grade, KEGG ID = K10814)
- *ctg1_2592*: 2956797 - 2958206 --> hydrogen cyanide synthase HcnB [EC:1.4.99.5] (B grade, KEGG ID = K10815)
- *ctg1_2593*: 2958199 - 2959455 --> hydrogen cyanide synthase HcnC [EC:1.4.99.5] (B grade, KEGG ID = K10816)
- ctg1_2594: 2959618 - 2960223 --> glutathione S-transferase [EC:2.5.1.18] (B grade, KEGG ID = K00799)

### Region 6 (NI-siderophore): 3832837 to 3869821
```{r}
Region6 <- Pseudomonas_chlororaphis_aureofaciens_annotations %>% filter(start_position >= 3832837 & end_position <=3869821)
```

Genes in NI-siderophore cluster:

- *ctg1_3372*: 3846837 - 3848735 --> *staphyloferrin* B synthase [EC:6.3.2.56] (B grade KEGG ID = K23375)
- ctg1_3373: 3848740 - 3849513 --> *staphyloferrin* B biosynthesis citrate synthase (B grade, KEGG ID = K23376)
- *ctg1_3374*: 3849507 - 3851369 --> 2-[(L-alanin-3-ylcarbamoyl)methyl]-3-(2-aminoethylcarbamoyl)-2-hydroxypropanoate synthase [EC:6.3.2.55] (B grade, KEGG ID = K23374)
- ctg1_3375: 3851394 - 3852812 --> UniRef90_A0A554P7Y5 DHA2 family efflux MFS transporter permease subunit n=30 Tax=Pseudomonadaceae TaxID=135621 RepID=A0A554P7Y5_9PSED (B grade, no hit in KEGG)
- ctg1_3376: 3852809 - 3854029 --> diaminopimelate decarboxylase [EC:4.1.1.20] (B grade, KEGG ID = K01586)
- *ctg1_3377*: 3854022 - 3855821 --> UniRef90_A0A0C5EL69 AcsD protein n=70 Tax=Pseudomonadaceae TaxID=135621 RepID=A0A0C5EL69_9PSED (B grade, no hit in KEGG)

### Region 7 (hydrogen cyanide cluster): 3888854 to 3901701
```{r}
Region7 <- Pseudomonas_chlororaphis_aureofaciens_annotations %>% filter(start_position >= 3888854 & end_position <=3901701)
```

- ctg1_3405: 3890984 - 3892441 --> succinate-semialdehyde dehydrogenase / glutarate-semialdehyde dehydrogenase [EC:1.2.1.16 1.2.1.79 1.2.1.20] (B grade, KEGG ID = K00135)
- ctg1_3406: 3892700 - 3893857 --> acetylornithine deacetylase [EC:3.5.1.16] (B grade, KEGG ID = K01438)
- *ctg1_3407*: 3893854 - 3895008 --> UniRef90_A0AA45XIN0 Sarcosine oxidase subunit beta n=5 Tax=Pseudomonas TaxID=286 RepID=A0AA45XIN0_9PSED (no KEGG hit, B grade, UniRef ID = A0AA45XIN0_9PSED)
  - Catalyzes oxidative demethylation of sarcosine (N-methylglycine) to yield glycine
- *ctg1_3408*: 3895005 - 3896411 --> UniRef90_A0A3G7K0E2 Opine oxidase subunit A n=39 Tax=Pseudomonas TaxID=286 RepID=A0A3G7K0E2_9PSED, (no KEGG hit, B grade, Uniref ID = A0A3G7K0E2_9PSED)
  - Cleaves opines into amino acids and pyruvate
- *ctg1_3409*: 3896393 - 3896701 --> UniRef90_A0A5M7CTD1 (2Fe-2S)-binding protein n=14 Tax=Pseudomonadaceae TaxID=135621 RepID=A0A5M7CTD1_9PSED (C grade, no KEGG hit, UniRef ID = A0A5M7CTD1_9PSED)
- ctg1_3410: 3896698 - 3898539 --> peptide/nickel transport system ATP-binding protein (B grade, KEGG ID = K02031)
- ctg1_3411: 3898541 - 3899416 --> UniRef90_A0A0C1ZSU6 DNA-directed RNA polymerase subunit alpha n=30 Tax=Pseudomonadaceae TaxID=135621 RepID=A0A0C1ZSU6_PSEFL (no KEGG hit, B grade, Uniref ID = A0A0C1ZSU6_PSEFL)
- ctg1_3412: 3899413 - 3900369 --> peptide/nickel transport system permease protein (B grade, KEGG ID = K02033)

### Region 10 (NRP-metallophore): 5056434 to 5155407
```{r}
Region10 <- Pseudomonas_chlororaphis_aureofaciens_annotations %>% filter(start_position >= 5056434 & end_position <=5155407)
```

- *ctg1_4474*: 5,076,434 - 5,077,768 --> L-ornithine N5-monooxygenase [EC:1.14.13.195 1.14.13.196] (B grade, KEGG ID = K10531)
- ctg1_4475: 5,077,900 - 5,078,496 --> UniRef90_A0A165YGX4 RNA polymerase subunit sigma n=4 Tax=Pseudomonas TaxID=286 RepID=A0A165YGX4_PSEFL (C grade, no KEGG ID)
- ctg1_4476: 5,078,716 - 5,079,888 --> membrane fusion protein, macrolide-specific efflux system (B grade, KEGG ID = K13888)
- ctg1_4477: 5,079,889 - 5,081,862 --> macrolide transport system ATP-binding/permease protein [EC:7.6.2.-] (B grade, KEGG ID = K05685)
- ctg1_4478: 5,081,870 - 5,083,279 --> outer membrane protein, multidrug efflux system (B grade, KEGG ID = K18139)
- *ctg1_4479*: 5,083,355 - 5,084,977 --> UniRef90_A0A554PAC0 Twin-arginine translocation signal domain-containing protein n=19 Tax=Pseudomonadaceae TaxID=135621 RepID=A0A554PAC0_9PSED (B grade, no KEGG hit)
- ctg1_4480: 5,085,186 - 5,086,595 --> membrane dipeptidase [EC:3.4.13.19] (B grade, KEGG ID = K01273)
- ctg1_4481: 5,086,617 - 5,087,915 --> UniRef90_A0A2C9EQN1 Isopenicillin N epimerase CefD n=56 Tax=Pseudomonadaceae TaxID=135621 RepID=A0A2C9EQN1_PSEPH (B grade, no KEGG hit)
- *ctg1_4482*: 5,087,959 - 5,088,855 --> UniRef90_Q4K997 Chromophore maturation protein *PvdO* n=55 Tax=Pseudomonadaceae TaxID=135621 RepID=Q4K997_PSEF5 (B grade, no KEGG hit)
- ctg1_4483: 5,088,910 - 5,089,773 --> UniRef90_Q4K996 phosphoribosylglycinamide formyltransferase 1 n=49 Tax=Pseudomonadaceae TaxID=135621 RepID=Q4K996_PSEF5 (B grade, no KEGG hit)
- ctg1_4484: 5,089,961 - 5,091,613 --> putative pyoverdin transport system ATP-binding/permease protein (B grade, KEGG ID = K06160)
- ctg1_4485: 5,091,714 - 5,094,182 --> outer-membrane receptor for ferric coprogen and ferric-rhodotorulic acid (B grade, KEGG ID = K16088)
- *ctg1_4486*: 5,094,570 - 5,107,805 --> UniRef90_A0A554PAB6 Amino acid adenylation domain-containing protein n=22 Tax=Pseudomonas TaxID=286 RepID=A0A554PAB6_9PSED (B grade, no KEGG hit)
- *ctg1_4487*: 5,107,818 - 5,114,414 --> UniRef90_A0A554PAC2 Non-ribosomal peptide synthetase n=26 Tax=Pseudomonas TaxID=286 RepID=A0A554PAC2_9PSED (B grade, no KEGG hit)
- *ctg1_4488*: 5,114,414 - 5,125,576 --> UniRef90_UPI002407C9DD amino acid adenylation domain-containing protein n=4 Tax=Pseudomonas chlororaphis TaxID=587753 RepID=UPI002407C9DD (B grade, no KEGG hit)

### Region 11 (NRPS): 5189456 to 5242472
```{r}
Region11 <- Pseudomonas_chlororaphis_aureofaciens_annotations %>% filter(start_position >= 5189456 & end_position <=5242472)
```

NRPS region here is really only one gene:
- *ctg1_4563*: 5209456 - 5222472 --> UniRef90_A0A554PEL7 Non-ribosomal peptide synthetase n=29 Tax=Pseudomonadaceae TaxID=135621 RepID=A0A554PEL7_9PSED (B grade, no KEGG hit)



# Figure 2A: Wild type vs deletion cell growth in corrosion assays as measured by viable plate counts

1. Load the plate count data.
```{r}
Plate_counts_combined <- fread("Plate_counts_combined_20250915.txt", sep = "\t", header = "auto")
```

2. Add new variables to the data frame parsing the Comment variable that contains detailed informtion on the sample
```{r}
Plate_counts_combined_treatments_added <- Plate_counts_combined %>%
                                          #Add time as a numeric variable
                                          mutate(Time = case_when(grepl("0hr", Comment) ~ "0",
                                                                  grepl("24hr", Comment) ~ "24",
                                                                  grepl("72hr", Comment) ~ "72",
                                                                  grepl("168hr", Comment) ~ "168")) %>%
  
                                          mutate(Time = as.numeric(Time)) %>%
        
                                          #Add wire treatment as a dedicated variable and leave as character
                                          mutate(Wire_treatment = case_when(grepl("_Wire", Comment) ~ "Wire included",
                                                                            grepl("NoWire", Comment) ~ "No wire")) %>%
  
  
                                          #Add strain information
                                          mutate(Strain = case_when(grepl("+10%inoc_", Comment) ~ "Wild type",
                                                                    grepl("DMSO", Comment) ~ "Wild type",
                                                                    grepl("0%inoc∆phz", Comment) ~ "∆phz",
                                                                    grepl("0%inoc∆hcn", Comment) ~ "∆hcn",
                                                                    grepl("DD", Comment) ~ "∆phz∆hcn")) %>%
  
                                          #Add DMSO information
                                          mutate(DMSO_treatment = case_when(!grepl("DMSO", Comment) ~ "No DMSO",
                                                                            grepl("+DMSO", Comment) ~ "DMSO")) 

Plate_counts_combined_summary_statistics <- Plate_counts_combined_treatments_added %>%
                                            group_by(Strain, Wire_treatment, DMSO_treatment, Flask_number, Time) %>%
                                            filter(CFU_mL >0) %>% #remove any values that had 0 counts on the plate for graphing
            
                                            #Calculate plate counts using technical replicates for each flask at each time point
                                            mutate(Technical_mean_CFU_mL = mean(CFU_mL),
                                                   Technical_stdev_CFU_mL = sd(CFU_mL)) %>%
                              
                                             ungroup() %>%
                            
                                             #Calculate plate counts using average of technical replicates for all flasks (biological replicates)
                                             group_by(Strain, Wire_treatment, DMSO_treatment, Time) %>%
                                             
                                             mutate(Biological_mean_CFU_mL = mean(Technical_mean_CFU_mL),
                                                    Biological_stdev_CFU_mL = sd(Technical_mean_CFU_mL))

write.table(Plate_counts_combined_summary_statistics, file = "Plate_counts_combined_summary_statistics_20251009.txt", sep = "\t", 
            row.names = FALSE, col.names = TRUE) 
```

3. Plot the growth data, summarizing the means and standard deviations prior to the ggplot command.
```{r}
#Adjust the presentation aesthetics
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
        size = 48, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", 
        size = 48, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
        size = 50, angle = 0, hjust = .5, vjust = 0, face = "bold"),
        axis.title.y = element_text(color = "grey20", 
        size = 50, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 48, colour = "grey20", face = "bold"),
        strip.text.y = element_text(size = 48, colour = "grey20"),
        plot.title = element_text(size = 48, face = "bold", hjust = .5),
        legend.title=element_text(size=48, face = "bold"), 
        legend.text=element_text(size=48),
        panel.border = element_rect(colour = "black", fill=NA, size=1))

dark2_colors <- brewer.pal(n = 7, name = "Dark2") #generate colour palette with desired levels to match ICP-MS data

#Create custom display orders
Plate_counts_combined_treatments_added$Strain_ordered <- factor(Plate_counts_combined_treatments_added$Strain,
                                                             levels = c("Wild type",
                                                                        "∆phz",
                                                                        "∆hcn",
                                                                        "∆phz∆hcn"))


Growth_curves_log_scale <- Plate_counts_combined_treatments_added %>%
                           filter(DMSO_treatment == "No DMSO") %>% #use this to only focus on experiments that did not receive DMSO
                           group_by(Strain, Wire_treatment, DMSO_treatment, Flask_number, Time) %>%
  
                           filter(CFU_mL >0) %>% #remove any values that had 0 counts on the plate for graphing
            
                           #Calculate plate counts using technical replicates for each flask at each time point
                           mutate(Technical_mean_CFU_mL = mean(CFU_mL),
                                  Technical_stdev_CFU_mL = sd(CFU_mL)) %>%
            
                           ungroup() %>%
          
                           #Calculate plate counts using average of technical replicates for all flasks (biological replicates)
                           group_by(Strain, Wire_treatment, DMSO_treatment, Time) %>%
                           
                           mutate(Biological_mean_CFU_mL = mean(Technical_mean_CFU_mL),
                                  Biological_stdev_CFU_mL = sd(Technical_mean_CFU_mL)) %>%
  
                           
                           #plotting script
                      
                           ggplot(aes(x=Time, y = log10(Biological_mean_CFU_mL), fill = Strain_ordered)) +
                           geom_errorbar(aes(ymin = log10(Biological_mean_CFU_mL - Biological_stdev_CFU_mL),
                                             ymax = log10(Biological_mean_CFU_mL + Biological_stdev_CFU_mL),
                                             colour = Strain_ordered), linewidth = 1) +
                           geom_point(shape = 21, color = "black", stroke = 1.25, size = 10) +
                           
                           geom_line(aes(group = Flask_number, color = Strain_ordered), linewidth = 1)+
                           scale_x_continuous(breaks = seq(0, 168, by = 24))+
                           facet_wrap(~ Wire_treatment, ncol = 1) +
                           ylab("log10[CFU/mL]") +
                           scale_y_continuous(labels = label_math(), limits = c(6, 10.5))+
                           xlab("Time elapsed (hours)") + 
                           scale_color_manual(name = "Strain", 
                                               values = dark2_colors,
                                               labels = c("Wild type", 
                                                        expression(italic("∆phz")),
                                                        expression(italic("∆hcn")),
                                                        expression(italic("∆phz∆hcn"))))+
                           scale_fill_manual(name = "Strain", 
                                             values = dark2_colors,
                                             labels = c("Wild type", 
                                                      expression(italic("∆phz")),
                                                      expression(italic("∆hcn")),
                                                      expression(italic("∆phz∆hcn")))) +
                           theme +
                           theme(plot.margin = unit(c(1,1,1.5,1), "cm"),
                                 axis.title.x = element_text(vjust = -2),
                                 axis.title.y = element_text(vjust = 2))
          
Growth_curves_log_scale

Growth_curves_log_scale_no_legend_angled <- Growth_curves_log_scale + theme(axis.text.x = element_text(angle = 45))

ggsave("Fig2A_plate_count_data_WT_phz_mutant_scatterplot_log10_20260820.jpg", dpi = 300, height=16, width=20, device=jpeg)
ggsave("Fig2A_plate_count_data_WT_phz_mutant_scatterplot_log10_20260820.tiff", dpi = 300, height=16, width=20, device=tiff)
ggsave("Fig2A_plate_count_data_WT_phz_mutant_scatterplot_log10_20260820.pdf", dpi = 300, height=16, width=20, device=cairo_pdf)
```


# Figure 2B: Cobalt dose-response curve reporting growth rate of P. chlororaphis subsp. aureofaciens in medium amended with CoCl2

1. Load the data where growth rates were manually calculated and refitting the dose response curve.
```{r}
#Load the data
library(tidyverse)
library(growthcurver)
library(drc)
library(readxl)

Growth_data <- read_excel("Fig2B_source_data_20260619.xlsx", sheet = "Growth rates")

Growth_rate_4_param_model_manual<- drm(Growth_rate ~ Total_cobalt_recovered_uM, 
                                data = Growth_data, 
                                fct = LL.4())

summary(Growth_rate_4_param_model_manual)

Growth_rate_dose_response_curve_summary_manual <- ED(Growth_rate_4_param_model_manual, c(5, 10, 50), interval = "delta")
Growth_rate_dose_response_curve_summary_manual

# Generate smooth prediction curve
prediction_df <- data.frame(Total_cobalt_recovered_uM = seq(
    min(Growth_data$Total_cobalt_recovered_uM[Growth_data$Total_cobalt_recovered_uM > 0]),
    max(Growth_data$Total_cobalt_recovered_uM, na.rm = TRUE),
    length.out = 1000))

prediction_df$Predicted_growth_rate <- predict(Growth_rate_4_param_model_manual, newdata = prediction_df, type = "response")

#Adjust the presentation aesthetics
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
        size = 48, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", 
        size = 48, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
        size = 50, angle = 0, hjust = .5, vjust = 0, face = "bold"),
        axis.title.y = element_text(color = "grey20", 
        size = 50, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 50, colour = "grey20", face = "bold"),
        strip.text.y = element_text(size = 50, colour = "grey20"),
        plot.title = element_text(size = 48, face = "bold", hjust = .5),
        legend.title=element_text(size=50, face = "bold"), 
        legend.text=element_text(size=50),
        panel.border = element_rect(colour = "black", fill=NA, size=1))

# Create ggplot version
Growth_rate_dose_response_plot_manual <- ggplot() +
                                         geom_point(data = Growth_data, aes(x = Total_cobalt_recovered_uM,
                                                                            y = Growth_rate), 
                                                    size = 10, stroke = 2, shape = 21) +
                                         geom_line(data = prediction_df, aes(x = Total_cobalt_recovered_uM,
                                                                             y = Predicted_growth_rate), 
                                                   linewidth = 1.5) +
                                         scale_x_continuous(trans = "log10") +
                                         scale_y_continuous(limits = NULL,
                                                            expand = expansion(mult = c(0.1, 0.2))) +
                                         coord_cartesian(ylim = c(0, 0.2)) +
                                         labs(x = "Co(2+) concentration (µM)",
                                              y = "Growth rate (hr<sup>-1</sup>)") +
                                         theme +
                                         theme(panel.border = element_rect(colour = "black", fill = NA,
                                                                           linewidth = 1.5), 
                                               axis.title.x = element_text(vjust = -1.5),
                                               axis.title.y = ggtext::element_markdown(face = "bold", 
                                                                                       margin = margin(r = 20)),
                                               panel.background = element_rect(fill = "white", colour = NA),
                                               panel.grid = element_blank(),
                                               axis.ticks = element_line(colour = "black", linewidth = 1.2),
                                               axis.text = element_text(size = 10), 
                                               plot.margin = unit(c(1,1,1,1), "cm"))
Growth_rate_dose_response_plot_manual

ggsave("Fig2B_ggplot_20260820.jpg", dpi = 300, height=16, width=20, device=jpeg)
ggsave("Fig2B_ggplot_20260820.tiff", dpi = 300, height=16, width=20, device=tiff)
ggsave("Fig2B_ggplot_20260820.pdf", dpi = 300, height=16, width=20, device=cairo_pdf)

```


#Figure 2C: Stacked barplot of cobalt localization in dose-response experiments.

1. Import source data and trim to necessary columns.
```{r}
#Load the data
Uptake_data <- read_excel("Fig2C-S9B_source_data_20260619.xlsx", sheet = "Figure 2C", range = "A2:M18") %>% 
  dplyr::select(`Sample Name`, Comment, `Total Co(2+) recovered [ uM ]`, `Free Co(2+) recovered [ uM ]`, `Cell-adsorbed Co(2+) [ uM ]`, `Intracellular Co(2+) [ uM ]`)

Uptake_data_reformatted <- Uptake_data %>% 
  rename(Short_name = `Sample Name`) %>%
  rename(Sample_name = Comment) %>% 
  mutate(Replicate = case_when(grepl(".1", Short_name) ~ "1",
                               grepl(".2", Short_name) ~ "2",
                               grepl(".3", Short_name) ~ "3")) %>%
  mutate(Spiked_cobalt_uM = case_when(grepl("2000µM", Sample_name) ~ "2000", 
                                      grepl("200µM", Sample_name) ~ "200", 
                                      grepl("20µM", Sample_name) ~ "20", 
                                      grepl("2µM", Sample_name) ~ "2", 
                                      grepl("0µM", Sample_name) ~ "0")) %>% 
  pivot_longer(cols = c(`Total Co(2+) recovered [ uM ]`,
                        `Free Co(2+) recovered [ uM ]`,
                        `Cell-adsorbed Co(2+) [ uM ]`,
                        `Intracellular Co(2+) [ uM ]`),
               names_to = "Cobalt_fraction",
               values_to = "Concentration_uM") %>%
  mutate(Cobalt_fraction = case_when(Cobalt_fraction == "Total Co(2+) recovered [ uM ]" ~ "Total",
                                     Cobalt_fraction == "Free Co(2+) recovered [ uM ]" ~ "In solution",
                                     Cobalt_fraction == "Cell-adsorbed Co(2+) [ uM ]" ~ "Cell-adsorbed",
                                     Cobalt_fraction == "Intracellular Co(2+) [ uM ]" ~ "Intracellular")) %>%
  filter(!is.na(Replicate)) %>% 
  group_by(Sample_name, Replicate) %>%
  mutate(Percent_of_total = 100 * Concentration_uM / Concentration_uM[Cobalt_fraction == "Total"]) %>%
  ungroup()
  
```


2. Create a stacked barplot for 20µM to 2000µM Co(2+) treatments.
```{r}
barplot_theme <-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
                 panel.background = element_blank(), axis.line = element_line(colour = "black"),
                 axis.text.x = element_text(color = "black", size = 48),
                 axis.text.y = element_text(color = "black", size = 48),
                 axis.title.x = element_text(color = "black", size = 50, vjust = -3, face = "bold"),
                 axis.title.y = element_text(color = "black", size = 50, vjust = 4, face = "bold"),
                 strip.text.x = element_text(color = "black", size = 50, face = "bold"),
                 strip.text.y = element_text(color = "black", size = 48, face = "bold"),
                 legend.title=element_text(size=50, vjust = 2, face = "bold"), 
                 legend.text=element_text(size=50),
                 legend.key = element_rect(fill = "white", color = NA),
                 plot.title = element_text(size = 48, face = "bold", hjust = 0.5),
                 panel.border = element_rect(colour = "black", fill=NA, size=1),
                 plot.margin = unit(c(2, 2, 2, 2), "cm"))

Uptake_data_barplot <- Uptake_data_reformatted %>%
                       filter(Spiked_cobalt_uM %in% c(20, 200, 2000)) %>%
                       filter(Cobalt_fraction %in% c("In solution", "Cell-adsorbed", "Intracellular")) %>%
                       group_by(Sample_name) %>%
                       ggplot(aes(x = Replicate, y = Percent_of_total, fill = Cobalt_fraction)) +
                       geom_bar(aes(group = Replicate), position="stack", stat="identity", 
                                colour = "black", linewidth = 1) +
                       facet_wrap(~ Spiked_cobalt_uM,
                                  scales = "free_y",
                                  nrow = 1,
                                  labeller = labeller(Spiked_cobalt_uM = function(x) 
                                    paste0(x, " µM"))) +
                       scale_fill_manual(name = "Localization", 
                                         breaks = c("Intracellular", "Cell-adsorbed", "In solution"), 
                                         values = c("In solution" = "grey90",   
                                                    "Cell-adsorbed" = "grey45",  
                                                    "Intracellular" = "grey10")) +
                       barplot_theme +
                       labs(x = "Replicate" , y = "Percent of total Co(2+) recovered (%)") +
                       theme(legend.key.size = unit(2, 'cm'))
Uptake_data_barplot

ggsave("Fig2C_Cobalt_localization_stacked_barplot_faceted_20260820.jpg", dpi = 300, height=16, width=42, device=jpeg)
ggsave("Fig2C_Cobalt_localization_stacked_barplot_faceted_20260820.tiff", dpi = 300, height=16, width=42, device=tiff)

```

# Figure S9: Log(µmol Co(2+)/CFU) over Co(2+) exposure from spike concentrations in corrosion assays compared to dose-response experiments

**Note:** These analyses were conducted manually in MS Excel and are tied to the source files:
- "FigS9A_source_data_20260619.xlsx" on the sheet "Figure S9A"
- "Fig2C-FigS9B_source_data_20260619.xlsx" on the sheet "Figure S9B"


#Figure S10: Boxplots of [Co(2+)] between A-F and F-A preservation in a Co(2+) dose-response curve with wild-type cells

1. Import the source data.
```{r}
Cobalt_data_treatments_added <- fread("FigS10_source_data_20260620.txt", sep = "\t", header = "auto")

```

2. Create a boxplot with the full data set and another version with the extreme values from the 2 µM treatment removed.
```{r}
Cobalt_boxplot_free <- ggplot(Cobalt_data_treatments_added,
                   aes(x = Sample_preservation,
                       y = Cobalt_concentration_uM,
                       fill = Sample_preservation)) +
  geom_boxplot(width = 0.6, outlier.shape = NA, alpha = 0.8) +
  
  geom_jitter(
    shape = 21,           # filled circle
    colour = "black",     # optional outline
    width = 0.15,
    height = 0,
    alpha = 0.6,
    size = 1.8
  ) +
  
  facet_wrap(
    ~ Spike_concentration,
    scales = "free_y",
    nrow = 1,
    labeller = labeller(
      Spike_concentration = function(x) paste0(x, " µM")
    )
  ) +
  labs(
    x = "Preservation Method",
    y = "Cobalt concentration (µM)",
    fill = "Preservation Method"
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, colour = "black", size = 18),
    axis.text.y = element_text(colour = "black", size = 18),
    axis.title.x = element_text(margin = margin(t = 16), size = 20),
    axis.title.y = element_text(margin = margin(r = 16), size = 20),
    panel.grid = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5),
    strip.background = element_rect(fill = "grey85", colour = "grey50"),
    strip.text = element_text(face = "bold", size = 22),
    legend.position = "none"
  )

Cobalt_boxplot_free
Cobalt_boxplot_free
ggsave("FigS7_Cobalt_recovered_sample_preservation_method_comparison_20260820.pdf", dpi = 300, height=6, width=22, device= pdf)

Cobalt_boxplot_free_outlier_removed <- Cobalt_data_treatments_added %>%
                                       dplyr::filter(!Sample_name == "2026-05-01_48hr_10%inoc+2µMCo2+_1_F-A", !Sample_name == "2026-05-01_48hr_10%inoc+2µMCo2+_1_A-F") %>% #remove outlier
  ggplot(aes(x = Sample_preservation,
                       y = Cobalt_concentration_uM,
                       fill = Sample_preservation)) +
  geom_boxplot(width = 0.6, outlier.shape = NA, alpha = 0.8) +
  
  geom_jitter(
    shape = 21,           # filled circle
    colour = "black",     # optional outline
    width = 0.15,
    height = 0,
    alpha = 0.6,
    size = 1.8
  ) +
  
  facet_wrap(
    ~ Spike_concentration,
    scales = "free_y",
    nrow = 1,
    labeller = labeller(
      Spike_concentration = function(x) paste0(x, " µM")
    )
  ) +
  labs(
    x = "Preservation Method",
    y = "Cobalt concentration (µM)",
    fill = "Preservation Method"
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, colour = "black", size = 18),
    axis.text.y = element_text(colour = "black", size = 18),
    axis.title.x = element_text(margin = margin(t = 16), size = 20),
    axis.title.y = element_text(margin = margin(r = 16), size = 20),
    panel.grid = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5),
    strip.background = element_rect(fill = "grey85", colour = "grey50"),
    strip.text = element_text(face = "bold", size = 22),
    legend.position = "none"
  )

Cobalt_boxplot_free_outlier_removed
ggsave("FigS10_Cobalt_recovered_sample_preservation_method_comparison_outlier_removed_20260820.pdf", dpi = 300, height=6, width=22, device= pdf)
```


# Figure 3: Boxplot of [Co(2+)] at 168 hours for oxic vs anoxic wild-type live cells, wild-type spent medium, and sterile medium (with and without supplied nitrate)

1. Load the data frame from the compiled data. 
```{r}
All_cobalt_data_combined_revised <- fread("All_cobalt_ICPMS_data_cleaned_formatted_revised_mass_balance_revised_20260615.txt", sep = "\t", header = "auto")


Oxic_vs_anoxic_data <- All_cobalt_data_combined_revised %>%
                       filter(DMSO_treatment == "No DMSO") %>% #use this to only focus on experiments that did not receive DMSO
                       filter(Sample_preservation == "Acidified, filtered") %>%
                       filter(Metabolite_treatment == "Nothing added") %>%
                       filter(Time == 168 & Wire_treatment == "Wire included") %>%
                       filter(Abiotic_vs_biotic == "Wild type" |
                              Abiotic_vs_biotic == "Wild type, anoxic" |
                              Abiotic_vs_biotic == "Wild type, anoxic, nitrate" |
                              Abiotic_vs_biotic == "Spent medium" |
                              Abiotic_vs_biotic == "Spent medium from WT, anoxic" |
                              Abiotic_vs_biotic == "Spent medium from WT, anoxic, nitrate" |
                              Abiotic_vs_biotic == "Sterile medium" |
                              Abiotic_vs_biotic == "Sterile medium, anoxic" |
                              Abiotic_vs_biotic == "Sterile medium, anoxic, nitrate")
```

2. Generate the linear models and verify assumptions for the ANOVA using type III sum of squares.
```{r}
Oxic_vs_anoxic_model_log10 <- lm(log10(Cobalt_concentration_uM) ~ Abiotic_vs_biotic, data = Oxic_vs_anoxic_data, contrasts = list(Abiotic_vs_biotic = contr.sum), type =3)

Oxic_vs_anoxic_model_log10
par(mfrow = c(2,2))
plot(Oxic_vs_anoxic_model_log10)

#Run Breusch Pagan test for homoscedasticity
bptest(Oxic_vs_anoxic_model_log10)

#Run Shapiro-Wilk normality test
shapiro.test(Oxic_vs_anoxic_model_log10$residuals)

#Run Durbin-Watson test for autocorrelation
dwtest(Oxic_vs_anoxic_model_log10)
```

### Interpretation
- Homoscedasticity looks more even across cobalt concentrations
- Normal distribution of residuals is much more linear, looks good
- BP test returns a p-value of *0.4949* --> we cannot reject null hypothesis of homoscedasticity.
- SW test returns a p-value of *0.8346* --> we cannot reject the null hypothesis of normally distributed residuals.
- DW test returns a p-value of *0.2502* --> we cannot reject the null hypothesis that there is no autocorrelation between data points.

**All assumptions met, okay to proceed with ANOVA**

3. Run the ANOVA.
```{r}
Oxic_anoxic_type_III_ANOVA_results <- Anova(Oxic_vs_anoxic_model_log10, type = 3)
Oxic_anoxic_type_III_ANOVA_results
```
- F-value = 111.53
- p-value = < 2.2e-16 ***
- There is a *significant* effect of the different oxic vs anoxic treatments

4. Run a Tukey post-hoc test to figure out where the significant differences are.
```{r}
Oxic_vs_anoxic_model_log10_converted <- aov(Oxic_vs_anoxic_model_log10)
Tukey_comparison_oxic_vs_anoxic <- TukeyHSD(Oxic_vs_anoxic_model_log10_converted, conf.level = 0.95)
Tukey_comparison_oxic_vs_anoxic
Tukey_comparison_oxic_vs_anoxic_data_frame <- as.data.frame(Tukey_comparison_oxic_vs_anoxic$Abiotic_vs_biotic)

letters_OvA <- multcompLetters4(
  Oxic_vs_anoxic_model_log10_converted,
  Tukey_comparison_oxic_vs_anoxic)
letters_OvA
```

5. Generate boxplot.
```{r}
Oxic_vs_anoxic_data <- Oxic_vs_anoxic_data %>%
  mutate(Abiotic_vs_biotic_ordered = fct_recode(Abiotic_vs_biotic,
    "Sterile medium,\nanoxic" = "Sterile medium, anoxic",
    "Sterile medium,\nanoxic, nitrate" = "Sterile medium, anoxic, nitrate",
    "Spent medium,\nanoxic" = "Spent medium from WT, anoxic",
    "Spent medium,\nanoxic, nitrate" = "Spent medium from WT, anoxic, nitrate",
    "Wild type,\nanoxic" = "Wild type, anoxic",
    "Wild type,\nanoxic, nitrate" = "Wild type, anoxic, nitrate"
  )) %>%
  mutate(
    Abiotic_vs_biotic_ordered = factor(
      Abiotic_vs_biotic_ordered,
      levels = c(
        "Wild type",
        "Wild type,\nanoxic",
        "Wild type,\nanoxic, nitrate",
        "Spent medium",
        "Spent medium,\nanoxic",
        "Spent medium,\nanoxic, nitrate",
        "Sterile medium",
        "Sterile medium,\nanoxic",
        "Sterile medium,\nanoxic, nitrate"
      )
    )
  )

theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 14, angle = 60, hjust = 0.5, vjust = 0.55, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 14, angle = 0, hjust = 0.5, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 18, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
        axis.title.y = element_text(color = "grey20", size = 18, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 20, colour = "grey20"),
        strip.text.y = element_text(size = 20, colour = "grey20"),
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5, vjust = 1),
        legend.title=element_text(size=18), 
        legend.text=element_text(size=18),
        legend.position = "right",
        panel.border = element_rect(colour = "black", fill=NA, size=1))

Oxic_vs_anoxic_boxplot <- Oxic_vs_anoxic_data %>%
                          mutate(Abiotic_vs_biotic_ordered = fct_recode(Abiotic_vs_biotic_ordered, "Sterile medium,\nanoxic" = "Sterile medium, anoxic")) %>%
                          mutate(Abiotic_vs_biotic_ordered = fct_recode(Abiotic_vs_biotic_ordered, "Sterile medium,\nanoxic, nitrate" = "Sterile medium, anoxic, nitrate")) %>%
                          mutate(Abiotic_vs_biotic_ordered = fct_recode(Abiotic_vs_biotic_ordered, "Spent medium,\nanoxic" = "Spent medium from WT, anoxic")) %>%
                          mutate(Abiotic_vs_biotic_ordered = fct_recode(Abiotic_vs_biotic_ordered, "Spent medium,\nanoxic, nitrate" = "Spent medium from WT, anoxic, nitrate")) %>%
                          mutate(Abiotic_vs_biotic_ordered = fct_recode(Abiotic_vs_biotic_ordered, "Wild type,\nanoxic" = "Wild type, anoxic")) %>%
                          mutate(Abiotic_vs_biotic_ordered = fct_recode(Abiotic_vs_biotic_ordered, "Wild type,\nanoxic, nitrate" = "Wild type, anoxic, nitrate")) %>% 
                          ggplot(aes(x=Abiotic_vs_biotic_ordered, 
                                     y=log10(Cobalt_concentration_uM_corrected))) + 
                          geom_boxplot() +
                          geom_jitter(fill = "grey", colour = "black", width = 0.1, pch =21, stroke =1)+
                          #ylim(0, 400) +
                          labs(title = "" ,
                          y = "log10[Co(2+) concentration (µM)]\nrecoveries applied", 
                          x = "Oxic vs anoxic treatments") +
                          theme +
                          theme(plot.margin = unit(c(1,1,1,1), "cm"),
                                axis.title.x = element_text(vjust = -2),
                                axis.title.y = element_text(vjust = 2))
Oxic_vs_anoxic_boxplot

```

**Results**

 Tukey multiple comparisons of means
    95% family-wise confidence level

Fit: aov(formula = Oxic_anoxic_model)

**Significant results**
                                                                               upr     p adj
Sterile medium-Spent medium                                           -0.148070641 0.0002454
Sterile medium, anoxic-Spent medium                                   -0.531948935 0.0000000
Sterile medium, anoxic, nitrate-Spent medium                          -0.347754657 0.0000024
Wild type-Spent medium                                                 0.804178670 0.0000023
Wild type, anoxic-Spent medium                                        -1.448277458 0.0000000
Wild type, anoxic, nitrate-Spent medium                               -0.085516555 0.0049383
Sterile medium, anoxic-Spent medium from WT, anoxic                   -0.336803387 0.0000060
Sterile medium, anoxic, nitrate-Spent medium from WT, anoxic          -0.152609109 0.0009380
Wild type-Spent medium from WT, anoxic                                 1.002825534 0.0000001
Wild type, anoxic-Spent medium from WT, anoxic                        -1.253131910 0.0000000
Sterile medium, anoxic-Spent medium from WT, anoxic, nitrate          -0.295160228 0.0000185
Sterile medium, anoxic, nitrate-Spent medium from WT, anoxic, nitrate -0.110965950 0.0029216
Wild type-Spent medium from WT, anoxic, nitrate                        1.044468693 0.0000000
Wild type, anoxic-Spent medium from WT, anoxic, nitrate               -1.211488751 0.0000000
Sterile medium, anoxic-Sterile medium                                 -0.185123590 0.0001211
Sterile medium, anoxic, nitrate-Sterile medium                        -0.000929312 0.0486799
Wild type-Sterile medium                                               1.142579720 0.0000000
Wild type, anoxic-Sterile medium                                      -1.101452113 0.0000000
Wild type-Sterile medium, anoxic                                       1.661727195 0.0000000
Wild type, anoxic-Sterile medium, anoxic                              -0.594230250 0.0000000
Wild type, anoxic, nitrate-Sterile medium, anoxic                      0.768530654 0.0020332
Wild type-Sterile medium, anoxic, nitrate                              1.477532916 0.0000000
Wild type, anoxic-Sterile medium, anoxic, nitrate                     -0.778424528 0.0000000
Wild type, anoxic-Wild type                                           -2.020165143 0.0000000
Wild type, anoxic, nitrate-Wild type                                  -0.657404240 0.0000000
Wild type, anoxic, nitrate-Wild type, anoxic                           1.684859177 0.0000000


**Non-significant results**
Spent medium from WT, anoxic-Spent medium                              0.126952725 0.5930314
Spent medium from WT, anoxic, nitrate-Spent medium                     0.085309566 0.3197846
Spent medium from WT, anoxic, nitrate-Spent medium from WT, anoxic     0.280455114 0.9999528
Sterile medium-Spent medium from WT, anoxic                            0.052206208 0.1970318
Wild type, anoxic, nitrate-Spent medium from WT, anoxic                0.109628993 0.4247827
Sterile medium-Spent medium from WT, anoxic, nitrate                   0.093849367 0.4577170
Wild type, anoxic, nitrate-Spent medium from WT, anoxic, nitrate       0.151272152 0.6944036
Wild type, anoxic, nitrate-Sterile medium                              0.261308790 1.0000000
Sterile medium, anoxic, nitrate-Sterile medium, anoxic                 0.506292552 0.6074848
Wild type, anoxic, nitrate-Sterile medium, anoxic, nitrate             0.584336376 0.1821758

## Interpretation
- Wild type live cells was significantly higher than its anoxic counterpart, spent medium (oxic or anoxic), and sterile medium (oxic or anoxic). This is as expected.
- Anoxic WT spent medium was not significantly different from oxic WT spent medium or sterile medium.
- Spent medium generated with the ∆phz mutant under anoxia could not be statistically distinguished from the sterile medium under anoxia
- Sterile medium with oxygen was significantly higher than sterile medium without oxygen
- Sterile medium with oxygen was also significantly higher than anoxic sterile medium (not too sure why).
- Honestly a little confused about why the anoxic spent medium, anoxic sterile medium, and anoxic WT are all significantly different for each other. If we plot the unmanipulated data (not log transformed), the differences look better, but then the assumptions are not met and the residuals are off.

**Letters to add to plot**
- Wild type = a
- Wild type, anoxic = b
- Wild type, anoxic, nitrate = cf
- Spent medium = d
- Spent medium, anoxic = cd
- Spent medium, anoxic, nitrate = cd
- Sterile medium = c
- Sterile medium, anoxic = e
- Sterile medium, anoxic, nitrate = ef

Add in letters to plot:
```{r}
#Create vector of letters in order to add to boxplot:
letters_vec <- c("a","b","cf","d","cd","cd","c","e","ef")

#Make dataframe to integrate into plot:
annotation_df <- data.frame(Abiotic_vs_biotic_ordered = levels(Oxic_vs_anoxic_data$Abiotic_vs_biotic_ordered),
                       letter = letters_vec)

#Integrate into boxplot:
Oxic_vs_anoxic_boxplot + 
  coord_cartesian(clip = "off") +
  geom_text(
    data = annotation_df,
    aes(x = Abiotic_vs_biotic_ordered, y = Inf, label = letter),
    vjust = -1.5,
    size = 6, fontface = "bold",
    inherit.aes = FALSE
  ) +
  theme(plot.margin = margin(t = 40, r = 10, b = 10, l = 10), 
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.4))

ggsave("Fig3_Cobalt_ICPMS_data_oxic_vs_anoxic_comparison_boxplot_20260820.jpg", dpi = 300, height=6, width=11, device=jpeg)
ggsave("Fig3_Cobalt_ICPMS_data_oxic_vs_anoxic_comparison_boxplot_20260820.tiff", dpi = 300, height=6, width=11, device=tiff)
ggsave("Fig3_Cobalt_ICPMS_data_oxic_vs_anoxic_comparison_boxplot_20260820.pdf", dpi = 300, height=6, width=11, device=cairo_pdf)

```


#Figure S12: Comparison of oxic vs anoxic for wild-type spent medium and ∆phz spent medium (no nitrate)

1. Load the data frame from the compiled data. 
```{r}
Oxic_vs_anoxic_spent <- All_cobalt_data_combined_revised %>%
                       filter(DMSO_treatment == "No DMSO") %>% #use this to only focus on experiments that did not receive DMSO
                       filter(Sample_preservation == "Acidified, filtered") %>%
                       filter(Metabolite_treatment == "Nothing added") %>%
                       filter(Time == 168 & Wire_treatment == "Wire included") %>%
                       filter(Abiotic_vs_biotic == "Spent medium" |
                              Abiotic_vs_biotic == "Spent medium from WT, anoxic" |
                              Abiotic_vs_biotic == "Spent medium from ∆phz" |
                              Abiotic_vs_biotic == "Spent medium from ∆phz, anoxic")
```

2. Generate the linear models and verify assumptions for the ANOVA using type III sum of squares.
```{r}
Oxic_vs_anoxic_spent_model_log10 <- lm(log10(Cobalt_concentration_uM) ~ Abiotic_vs_biotic, data = Oxic_vs_anoxic_spent, contrasts = list(Abiotic_vs_biotic = contr.sum), type =3)

Oxic_vs_anoxic_spent_model_log10
par(mfrow = c(2,2))
plot(Oxic_vs_anoxic_spent_model_log10)

#Run Breusch Pagan test for homoscedasticity
bptest(Oxic_vs_anoxic_spent_model_log10)

#Run Shapiro-Wilk normality test
shapiro.test(Oxic_vs_anoxic_spent_model_log10$residuals)

#Run Durbin-Watson test for autocorrelation
dwtest(Oxic_vs_anoxic_spent_model_log10)
```

### Interpretation
- Homoscedasticity looks more even across cobalt concentrations
- Normal distribution of residuals is much more linear, looks good
- BP test returns a p-value of *0.08972* --> we cannot reject null hypothesis of homoscedasticity.
- SW test returns a p-value of *0.2968* --> we cannot reject the null hypothesis of normally distributed residuals.
- DW test returns a p-value of *0.3568* --> we cannot reject the null hypothesis that there is no autocorrelation between data points.

**All assumptions met, okay to proceed with ANOVA**

3. Run the ANOVA.
```{r}
Oxic_anoxic_spent_type_III_ANOVA_results <- Anova(Oxic_vs_anoxic_spent_model_log10, type = 3)
Oxic_anoxic_spent_type_III_ANOVA_results
```
- F-value = 59.252
- p-value = 3.002e-06 ***
- There is a *significant* effect of the different oxic vs anoxic treatments

4. Run a Tukey post-hoc test to figure out where the significant differences are.
```{r}
Oxic_vs_anoxic_spent_model_log10_converted <- aov(Oxic_vs_anoxic_spent_model_log10)
Tukey_comparison_oxic_vs_anoxic_spent <- TukeyHSD(Oxic_vs_anoxic_spent_model_log10_converted, 
                                                  conf.level = 0.95)
Tukey_comparison_oxic_vs_anoxic_spent
Tukey_comparison_oxic_vs_anoxic_spent_dataframe <- as.data.frame(Tukey_comparison_oxic_vs_anoxic_spent$Abiotic_vs_biotic)
```

**Results**

 Tukey multiple comparisons of means
    95% family-wise confidence level

Fit: aov(formula = Oxic_anoxic_model)

**Significant results**
                                                                p adj
Spent medium from ∆phz-Spent medium                         0.0316047
Spent medium from ∆phz, anoxic-Spent medium                 0.0000150
Spent medium from ∆phz, anoxic-Spent medium from ∆phz       0.0000023
Spent medium from WT, anoxic-Spent medium from ∆phz         0.0027098
Spent medium from WT, anoxic-Spent medium from ∆phz, anoxic 0.0001658

**Non-significant results**
Spent medium from WT, anoxic-Spent medium                   0.2261245

**Letters to add to plot**
- Spent medium WT = a
- Spent medium WT, anoxic = a
- Spent medium ∆phz = b
- Spent medium ∆phz, anoxic = c

5. Generate boxplot.
```{r}
Oxic_vs_anoxic_spent <- Oxic_vs_anoxic_spent %>%
  mutate(Abiotic_vs_biotic_ordered = fct_recode(Abiotic_vs_biotic,
    "Spent medium\nWT" = "Spent medium",                                            
    "Spent medium\nWT, anoxic" = "Spent medium from WT, anoxic",
    "Spent medium\n∆phz" = "Spent medium from ∆phz", 
    "Spent medium\n∆phz, anoxic" = "Spent medium from ∆phz, anoxic")) %>%
  mutate(
    Abiotic_vs_biotic_ordered = factor(
      Abiotic_vs_biotic_ordered,
      levels = c(
        "Spent medium\nWT",
        "Spent medium\nWT, anoxic",
        "Spent medium\n∆phz", 
        "Spent medium\n∆phz, anoxic")))

theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 14, angle = 60, hjust = 0.5, vjust = 0.55, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 14, angle = 0, hjust = 0.5, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 18, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
        axis.title.y = element_text(color = "grey20", size = 18, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 20, colour = "grey20"),
        strip.text.y = element_text(size = 20, colour = "grey20"),
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5, vjust = 1),
        legend.title=element_text(size=18), 
        legend.text=element_text(size=18),
        legend.position = "right",
        panel.border = element_rect(colour = "black", fill=NA, size=1))

#Create vector of letters in order to add to boxplot:
letters_vec_spent <- c("a","a","b","c")

#Make dataframe to integrate into plot:
annotation_df_spent <- data.frame(Abiotic_vs_biotic_ordered = levels(Oxic_vs_anoxic_spent$Abiotic_vs_biotic_ordered),
                       letter = letters_vec_spent)

Oxic_vs_anoxic_spent_boxplot <- Oxic_vs_anoxic_spent %>% 
                          ggplot(aes(x=Abiotic_vs_biotic_ordered, 
                                     y=log10(Cobalt_concentration_uM_corrected))) + 
                          geom_boxplot() +
                          geom_jitter(fill = "grey", colour = "black", width = 0.1, pch =21, stroke =1)+
                          #ylim(0, 400) +
                          coord_cartesian(clip = "off") +
                          geom_text(data = annotation_df_spent,
                                    aes(x = Abiotic_vs_biotic_ordered, y = Inf, label = letter),
                                    vjust = -1.5,
                                    size = 6, fontface = "bold", 
                                    inherit.aes = FALSE) +
                          labs(title = "" ,
                          y = "log10[Co(2+) concentration (µM)]\ncorrected for recovery", 
                          x = "Oxic vs anoxic treatments") +
                          theme +
                          theme(plot.margin = margin(t = 40, r = 10, b = 10, l = 10), ,
                                axis.title.x = element_text(vjust = -2),
                                axis.title.y = element_text(vjust = 2), 
                                panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.4))
Oxic_vs_anoxic_spent_boxplot

ggsave("FigS12_Cobalt_ICPMS_data_oxic_vs_anoxic_spentmedium_boxplot_20260820.jpg", dpi = 300, height=6, width=11, device=jpeg)
ggsave("FigS12_Cobalt_ICPMS_data_oxic_vs_anoxic_spentmedium_boxplot_20260820.tiff", dpi = 300, height=6, width=11, device=tiff)
ggsave("FigS12_Cobalt_ICPMS_data_oxic_vs_anoxic_spentmedium_boxplot_20260820.pdf", dpi = 300, height=6, width=11, device=cairo_pdf)

```


#Figure S13: Boxplot of [Co(2+)] at 168 hours comparing all treatments

1. Load the data frame from the compiled data. 
```{r}
All_data <- All_cobalt_data_combined_revised %>%
                       filter(DMSO_treatment == "No DMSO") %>% #use this to only focus on experiments that did not receive DMSO
                       filter(Sample_preservation == "Acidified, filtered") %>%
                       filter(Metabolite_treatment == "Nothing added") %>%
                       filter(Time == 168 & Wire_treatment == "Wire included")
```

2. Generate the linear models and verify assumptions for the ANOVA using type III sum of squares.
```{r}
All_data_model_log10 <- lm(log10(Cobalt_concentration_uM) ~ Abiotic_vs_biotic, data = All_data, contrasts = list(Abiotic_vs_biotic = contr.sum), type =3)

All_data_model_log10
par(mfrow = c(2,2))
plot(All_data_model_log10)

#Run Breusch Pagan test for homoscedasticity
bptest(All_data_model_log10)

#Run Shapiro-Wilk normality test
shapiro.test(All_data_model_log10$residuals)

#Run Durbin-Watson test for autocorrelation
dwtest(All_data_model_log10)
```

### Interpretation
- Homoscedasticity looks more even across cobalt concentrations
- Normal distribution of residuals is much more linear, looks good
- BP test returns a p-value of *0.2965* --> we cannot reject null hypothesis of homoscedasticity.
- SW test returns a p-value of *0.3905* --> we cannot reject the null hypothesis of normally distributed residuals.
- DW test returns a p-value of *0.03467* --> we can reject the null hypothesis that there is no autocorrelation between data points.

**All assumptions met, okay to proceed with ANOVA**

3. Run the ANOVA.
```{r}
All_data_type_III_ANOVA_results <- Anova(All_data_model_log10, type = 3)
All_data_type_III_ANOVA_results
```
- F-value = 142.37
- p-value = < 2.2e-16 ***
- There is a *significant* effect of the different abiotic vs biotic treatments

4. Run a Tukey post-hoc test to figure out where the significant differences are.
```{r}
All_data_model_log10_converted <- aov(All_data_model_log10)
Tukey_comparison_all_data <- TukeyHSD(All_data_model_log10_converted, 
                                                  conf.level = 0.95)
Tukey_comparison_all_data
Tukey_comparison_all_data_dataframe <- as.data.frame(Tukey_comparison_all_data$Abiotic_vs_biotic)


# Generate letters from Tukey results
letters_All <- multcompLetters4(All_data_model_log10_converted, Tukey_comparison_all_data)
letters_All
```

**Results**

 Tukey multiple comparisons of means
    95% family-wise confidence level

Fit: aov(formula = Oxic_anoxic_model)

**Significant results**
                                                                                upr     p adj
∆phz-∆hcn                                                              0.5936886647 0.0162209
∆phz∆hcn-∆hcn                                                          0.8551418909 0.0000004
Heat kill-∆hcn                                                        -0.3701536398 0.0000000
Spent medium-∆hcn                                                     -0.0913340553 0.0017861
Spent medium from ∆phz, anoxic-∆hcn                                   -0.9161596451 0.0000000
Spent medium from WT, anoxic-∆hcn                                     -0.2432133678 0.0000070
Spent medium from WT, anoxic, nitrate-∆hcn                            -0.2848565269 0.0000014
Sterile medium-∆hcn                                                   -0.5185622306 0.0000000
Sterile medium, anoxic-∆hcn                                           -0.9021150285 0.0000000
Sterile medium, anoxic, nitrate-∆hcn                                  -0.7179207501 0.0000000
Wild type, anoxic-∆hcn                                                -1.8184435514 0.0000000
Wild type, anoxic, nitrate-∆hcn                                       -0.4556826479 0.0000000
Heat kill-∆phz                                                        -0.6834835800 0.0000000
Spent medium-∆phz                                                     -0.4046639955 0.0000000
Spent medium from ∆phz-∆phz                                           -0.0967304548 0.0019161
Spent medium from ∆phz, anoxic-∆phz                                   -1.2294895853 0.0000000
Spent medium from WT, anoxic-∆phz                                     -0.5565433079 0.0000000
Spent medium from WT, anoxic, nitrate-∆phz                            -0.5981864671 0.0000000
Sterile medium-∆phz                                                   -0.8318921708 0.0000000
Sterile medium, anoxic-∆phz                                           -1.2154449687 0.0000000
Sterile medium, anoxic, nitrate-∆phz                                  -1.0312506903 0.0000000
Wild type, anoxic-∆phz                                                -2.1317734916 0.0000000
Wild type, anoxic, nitrate-∆phz                                       -0.7690125880 0.0000000
Heat kill-∆phz∆hcn                                                    -0.9449368062 0.0000000
Spent medium-∆phz∆hcn                                                 -0.6661172217 0.0000000
Spent medium from ∆phz-∆phz∆hcn                                       -0.3581836810 0.0000001
Spent medium from ∆phz, anoxic-∆phz∆hcn                               -1.4909428115 0.0000000
Spent medium from WT, anoxic-∆phz∆hcn                                 -0.8179965341 0.0000000
Spent medium from WT, anoxic, nitrate-∆phz∆hcn                        -0.8596396933 0.0000000
Sterile medium-∆phz∆hcn                                               -1.0933453970 0.0000000
Sterile medium, anoxic-∆phz∆hcn                                       -1.4768981949 0.0000000
Sterile medium, anoxic, nitrate-∆phz∆hcn                              -1.2927039165 0.0000000
Wild type-∆phz∆hcn                                                    -0.1410069908 0.0001368
Wild type, anoxic-∆phz∆hcn                                            -2.3932267178 0.0000000
Wild type, anoxic, nitrate-∆phz∆hcn                                   -1.0304658142 0.0000000
Spent medium from ∆phz-Heat kill                                       0.9104834953 0.0000062
Spent medium from ∆phz, anoxic-Heat kill                              -0.2222756352 0.0000278
Sterile medium, anoxic-Heat kill                                      -0.2082310186 0.0000463
Sterile medium, anoxic, nitrate-Heat kill                             -0.0240367402 0.0248736
Wild type-Heat kill                                                    1.1311792432 0.0000000
Wild type, anoxic-Heat kill                                           -1.1245595415 0.0000000
Spent medium from ∆phz, anoxic-Spent medium                           -0.5444668653 0.0000000
Sterile medium-Spent medium                                           -0.1468694508 0.0000635
Sterile medium, anoxic-Spent medium                                   -0.5304222486 0.0000000
Sterile medium, anoxic, nitrate-Spent medium                          -0.3462279703 0.0000001
Wild type-Spent medium                                                 0.8054689554 0.0000001
Wild type, anoxic-Spent medium                                        -1.4467507716 0.0000000
Wild type, anoxic, nitrate-Spent medium                               -0.0839898680 0.0030401
Spent medium from ∆phz, anoxic-Spent medium from ∆phz                 -0.8090287604 0.0000000
Spent medium from WT, anoxic-Spent medium from ∆phz                   -0.1360824830 0.0006169
Spent medium from WT, anoxic, nitrate-Spent medium from ∆phz          -0.1777256422 0.0001399
Sterile medium-Spent medium from ∆phz                                 -0.4062740442 0.0000000
Sterile medium, anoxic-Spent medium from ∆phz                         -0.7949841438 0.0000000
Sterile medium, anoxic, nitrate-Spent medium from ∆phz                -0.6107898654 0.0000000
Wild type, anoxic-Spent medium from ∆phz                              -1.7113126667 0.0000000
Wild type, anoxic, nitrate-Spent medium from ∆phz                     -0.3485517631 0.0000003
Spent medium from WT, anoxic-Spent medium from ∆phz, anoxic            0.9966766475 0.0000003
Spent medium from WT, anoxic, nitrate-Spent medium from ∆phz, anoxic   0.9550334884 0.0000012
Sterile medium-Spent medium from ∆phz, anoxic                          0.7264850863 0.0000133
Wild type-Spent medium from ∆phz, anoxic                               1.6771852486 0.0000000
Wild type, anoxic-Spent medium from ∆phz, anoxic                      -0.5785535362 0.0000000
Wild type, anoxic, nitrate-Spent medium from ∆phz, anoxic              0.7842073674 0.0006027
Sterile medium, anoxic-Spent medium from WT, anoxic                   -0.3351712906 0.0000004
Sterile medium, anoxic, nitrate-Spent medium from WT, anoxic          -0.1509770122 0.0003644
Wild type-Spent medium from WT, anoxic                                 1.0042389712 0.0000000
Wild type, anoxic-Spent medium from WT, anoxic                        -1.2514998135 0.0000000
Sterile medium, anoxic-Spent medium from WT, anoxic, nitrate          -0.2935281315 0.0000020
Sterile medium, anoxic, nitrate-Spent medium from WT, anoxic, nitrate -0.1093338531 0.0015641
Wild type-Spent medium from WT, anoxic, nitrate                        1.0458821303 0.0000000
Wild type, anoxic-Spent medium from WT, anoxic, nitrate               -1.2098566544 0.0000000
Sterile medium, anoxic-Sterile medium                                 -0.1837909890 0.0000250
Wild type-Sterile medium                                               1.1436332334 0.0000000
Wild type, anoxic-Sterile medium                                      -1.1001195119 0.0000000
Wild type-Sterile medium, anoxic                                       1.6631406319 0.0000000
Wild type, anoxic-Sterile medium, anoxic                              -0.5925981528 0.0000000
Wild type, anoxic, nitrate-Sterile medium, anoxic                      0.7701627508 0.0009852
Wild type-Sterile medium, anoxic, nitrate                              1.4789463536 0.0000000
Wild type, anoxic-Sterile medium, anoxic, nitrate                     -0.7767924312 0.0000000
Wild type, anoxic-Wild type                                           -2.0187517058 0.0000000
Wild type, anoxic, nitrate-Wild type                                  -0.6559908023 0.0000000
Wild type, anoxic, nitrate-Wild type, anoxic                           1.6864912737 0.0000000


**Non-significant results**
Spent medium from ∆phz-∆hcn                                            0.2165994854 0.9992144
Wild type-∆hcn                                                         0.4337761756 0.4627853
∆phz∆hcn-∆phz                                                          0.5418119507 0.0902597
Wild type-∆phz                                                         0.1204462354 0.8387728
Spent medium-Heat kill                                                 0.6041049253 0.0523567
Spent medium from WT, anoxic-Heat kill                                 0.4506706422 0.9818951
Spent medium from WT, anoxic, nitrate-Heat kill                        0.4090274830 0.9996660
Sterile medium-Heat kill                                               0.1804790810 0.9975546
Wild type, anoxic, nitrate-Heat kill                                   0.2382013621 0.9996557
Spent medium from ∆phz-Spent medium                                    0.5882922653 0.0828367
Spent medium from WT, anoxic-Spent medium                              0.1284794121 0.7463806
Spent medium from WT, anoxic, nitrate-Spent medium                     0.0868362530 0.4204130
Wild type-Spent medium from ∆phz                                       0.5444261180 0.0834092
Sterile medium, anoxic-Spent medium from ∆phz, anoxic                  0.3377749868 1.0000000
Sterile medium, anoxic, nitrate-Spent medium from ∆phz, anoxic         0.5219692651 0.6623214
Spent medium from WT, anoxic, nitrate-Spent medium from WT, anoxic     0.2820872110 1.0000000
Sterile medium-Spent medium from WT, anoxic                            0.0535388090 0.2522437
Wild type, anoxic, nitrate-Spent medium from WT, anoxic                0.1112610900 0.5555962
Sterile medium-Spent medium from WT, anoxic, nitrate                   0.0951819681 0.5957059
Wild type, anoxic, nitrate-Spent medium from WT, anoxic, nitrate       0.1529042492 0.8417775
Sterile medium, anoxic, nitrate-Sterile medium                         0.0004032894 0.0506973
Wild type, anoxic, nitrate-Sterile medium                              0.2626413916 1.0000000
Sterile medium, anoxic, nitrate-Sterile medium, anoxic                 0.5079246485 0.7609747
Wild type, anoxic, nitrate-Sterile medium, anoxic, nitrate             0.5859684724 0.2315235

**Letters to add to plot**
- Wild type = ac
- ∆phz = ab
- ∆hcn = c
- ∆phz∆hcn = d
- Spent medium = ef
- Spent medium ∆phz = ce
- Heat kill = fg
- Sterile medium = gh
- Wild type, anoxic = i
- Wild type, anoxic, nitrate = gh
- Spent medium from WT, anoxic = fg
- Spent medium from WT, anoxic, nitrate = fg
- Spent medium ∆phz, anoxic = j
- Sterile medium, anoxic = j
- Sterile medium, anoxic, nitrate = hj

5. Generate boxplot.
```{r}
All_data <- All_data %>%
  mutate(Abiotic_vs_biotic_ordered = fct_recode(Abiotic_vs_biotic,
    "Spent medium\nWT" = "Spent medium",                                            
    "Spent medium\nWT, anoxic" = "Spent medium from WT, anoxic",
    "Spent medium\nWT, anoxic, nitrate" = "Spent medium from WT, anoxic, nitrate", 
    "Spent medium\n∆phz" = "Spent medium from ∆phz", 
    "Spent medium\n∆phz, anoxic" = "Spent medium from ∆phz, anoxic", 
    "Sterile medium,\nanoxic" = "Sterile medium, anoxic", 
    "Sterile medium,\nanoxic, nitrate" = "Sterile medium, anoxic, nitrate", 
    "Wild type,\nanoxic" = "Wild type, anoxic", 
    "Wild type,\nanoxic, nitrate" = "Wild type, anoxic, nitrate")) %>%
  mutate(Abiotic_vs_biotic_ordered = factor(Abiotic_vs_biotic_ordered, levels = c(
        "Wild type",
        "∆phz", 
        "∆hcn", 
        "∆phz∆hcn", 
        "Spent medium\nWT", 
        "Spent medium\n∆phz",
        "Heat kill", 
        "Sterile medium",
        "Wild type,\nanoxic",
        "Wild type,\nanoxic, nitrate",
        "Spent medium\nWT, anoxic", 
        "Spent medium\nWT, anoxic, nitrate", 
        "Spent medium\n∆phz, anoxic",
        "Sterile medium,\nanoxic", 
        "Sterile medium,\nanoxic, nitrate")))

theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 14, angle = 70, hjust = 0.5, vjust = 0.55, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 14, angle = 0, hjust = 0.5, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 18, angle = 0, hjust = 0.5, vjust = -1, face = "bold"),
        axis.title.y = element_text(color = "grey20", size = 18, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 20, colour = "grey20"),
        strip.text.y = element_text(size = 20, colour = "grey20"),
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5, vjust = 1),
        legend.title=element_text(size=18), 
        legend.text=element_text(size=18),
        legend.position = "right",
        panel.border = element_rect(colour = "black", fill=NA, size=1))

#Create vector of letters in order to add to boxplot:
letters_vec_all <- c("ac","ab","c", "d", "ef", "ce", "fg", "gh", "i", "gh", "fg", "fg", "j", "j", "hj")

#Make dataframe to integrate into plot:
annotation_df_all <- data.frame(Abiotic_vs_biotic_ordered = levels(All_data$Abiotic_vs_biotic_ordered),
                       letter = letters_vec_all)

All_data_boxplot <- All_data %>% 
                          ggplot(aes(x=Abiotic_vs_biotic_ordered, 
                                     y=log10(Cobalt_concentration_uM_corrected))) + 
                          geom_boxplot() +
                          geom_jitter(fill = "grey", colour = "black", width = 0.1, pch =21, stroke =1)+
                          #ylim(0, 400) +
                          coord_cartesian(clip = "off") +
                          geom_text(data = annotation_df_all,
                                    aes(x = Abiotic_vs_biotic_ordered, y = Inf, label = letter),
                                    vjust = -1.5,
                                    size = 6, fontface = "bold", 
                                    inherit.aes = FALSE) +
                          labs(title = "" ,
                          y = "log10[Co(2+) concentration (µM)]\nrecoveries applied", 
                          x = "Oxic vs anoxic treatments") +
                          theme +
                          theme(plot.margin = margin(t = 40, r = 10, b = 10, l = 10), ,
                                axis.title.x = element_text(vjust = -2),
                                axis.title.y = element_text(vjust = 2), 
                                panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.4))
All_data_boxplot

ggsave("FigS13_Cobalt_ICPMS_data_all_comparison_boxplot_20260820.jpg", dpi = 300, height=6, width=11, device=jpeg)
ggsave("FigS13_Cobalt_ICPMS_data_all_comparison_boxplot_20260820.tiff", dpi = 300, height=6, width=11, device=tiff)
ggsave("FigS13_Cobalt_ICPMS_data_all_comparison_boxplot_20260820.pdf", dpi = 300, height=6, width=11, device=cairo_pdf)

```


# Figure S14: Wild type growth in anoxic corrosion assays as measured by viable plate counts

1. Load the plate count data.
```{r}
Anoxic_plate_counts <- read_excel("Anoxic_plate_counts_combined_20260312.xlsx")
```

2. Add new variables to the data frame parsing the Comment variable that contains detailed informtion on the sample
```{r}
Anoxic_plate_counts_treatments_added <- Anoxic_plate_counts %>%
                                          #Add time as a numeric variable
                                          mutate(Time = case_when(grepl("0hr", Comment) ~ "0",
                                                                  grepl("24hr", Comment) ~ "24",
                                                                  grepl("72hr", Comment) ~ "72",
                                                                  grepl("168hr", Comment) ~ "168")) %>%
  
                                          mutate(Time = as.numeric(Time)) %>%
        
                                          #Add wire treatment as a dedicated variable and leave as character
                                          mutate(Wire_treatment = case_when(grepl("_Wire", Comment) ~ "Wire included",
                                                                            grepl("NoWire", Comment) ~ "No wire")) %>%
                                          #Add strain information
                                          mutate(Strain = case_when(grepl("+10%inoc_", Comment) ~ "Wild type")) %>%
  
                                          #Add NO3 treatment
                                          mutate(NO3_treatment = case_when(grepl("_NO3", Comment) ~ "NO3", 
                                                 !grepl("_NO3", Comment) ~ "No NO3"))
  

Anoxic_plate_counts_summary_statistics <- Anoxic_plate_counts_treatments_added %>%
                                            group_by(Wire_treatment, NO3_treatment, Flask_number, Time) %>%
                                            #filter(CFU_mL >0) %>% #remove any values that had 0 counts on the plate for graphing
            
                                            #Calculate plate counts using technical replicates for each flask at each time point
                                            mutate(Technical_mean_CFU_mL = mean(CFU_mL),
                                                   Technical_stdev_CFU_mL = sd(CFU_mL)) %>%
                              
                                             ungroup() %>%
                            
                                             #Calculate plate counts using average of technical replicates for all flasks (biological replicates)
                                             group_by(Wire_treatment, NO3_treatment, Time) %>%
                                             
                                             mutate(Biological_mean_CFU_mL = mean(Technical_mean_CFU_mL),
                                                    Biological_stdev_CFU_mL = sd(Technical_mean_CFU_mL))

write.table(Anoxic_plate_counts_summary_statistics, file = "Anoxic_plate_counts_summary_statistics_20260709.txt", sep = "\t", 
            row.names = FALSE, col.names = TRUE) 
```

3. Plot the growth data, summarizing the means and standard deviations prior to the ggplot command.
```{r}
#Adjust the presentation aesthetics
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
        size = 48, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", 
        size = 48, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
        size = 50, angle = 0, hjust = .5, vjust = 0, face = "bold"),
        axis.title.y = element_text(color = "grey20", 
        size = 50, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 48, colour = "grey20", face = "bold"),
        strip.text.y = element_text(size = 48, colour = "grey20"),
        plot.title = element_text(size = 48, face = "bold", hjust = .5),
        legend.title=element_text(size=48, face = "bold"), 
        legend.text=element_text(size=48),
        panel.border = element_rect(colour = "black", fill=NA, size=1))

dark2_colors <- brewer.pal(n = 2, name = "Dark2") #generate colour palette with desired levels to match ICP-MS data

Anoxic_growth_curves <- Anoxic_plate_counts_treatments_added %>%
  filter(Wire_treatment == "Wire included") %>%
  mutate(NO3_treatment = factor(NO3_treatment, levels = c("NO3", "No NO3"), labels = c("Nitrate", "No nitrate"))) %>%
                           group_by(Wire_treatment, NO3_treatment, Flask_number, Time) %>%
                           summarise(Technical_mean_CFU_mL = mean(CFU_mL),
                                     Technical_stdev_CFU_mL = sd(CFU_mL), 
                                     .groups = "drop") %>%
  group_by(Wire_treatment, NO3_treatment, Time) %>%
  summarise(
    Biological_mean_CFU_mL = mean(Technical_mean_CFU_mL),
    Biological_sd_CFU_mL = sd(Technical_mean_CFU_mL),
    n = n(),
    Biological_se_CFU_mL = Biological_sd_CFU_mL / sqrt(n),
    .groups = "drop") %>% 
                           #plotting script
                      
                           ggplot(aes(x=Time, y = Biological_mean_CFU_mL, colour = NO3_treatment, fill = NO3_treatment, group = NO3_treatment)) +
                           geom_line(linewidth = 1) +
                           geom_errorbar(aes(ymin = Biological_mean_CFU_mL - Biological_sd_CFU_mL,
                                             ymax = Biological_mean_CFU_mL + Biological_sd_CFU_mL), 
                                             width = 3, linewidth = 0.8) +
                           geom_point(shape = 21, colour = "black", size = 10, stroke = 1.2) +
                           facet_grid(~ Wire_treatment) +
                           scale_x_continuous(breaks = seq(0, 168, by = 24)) +
                           scale_y_continuous(labels = \(x) {sapply(x, function(y) {if (y == 0) return("0")
                             e <- floor(log10(y))
                             m <- signif(y / 10^e, 2)
                             if (m == 1) {paste0("10^", e)} else {paste0(m, "%*%10^", e)}}) |> parse(text = _)}) +
                           scale_colour_discrete(name = "Nitrate Treatment") + 
                           scale_fill_discrete(name = "Nitrate Treatment") + 
                           ylab("CFU/mL") +
                           xlab("Time elapsed (hours)") +
                           theme +
                           theme(plot.margin = unit(c(1,1,1.5,1), "cm"),
                                 axis.title.x = element_text(vjust = -2),
                                 axis.title.y = element_text(vjust = 2))
          unique(Anoxic_plate_counts_treatments_added$NO3_treatment)
Anoxic_growth_curves

ggsave("FigS14_anoxic_plate_count_scatterplot_20260820.jpg", dpi = 300, height=16, width=20, device=jpeg)
ggsave("FigS14_anoxic_plate_count_scatterplot_20260820.tiff", dpi = 300, height=16, width=20, device=tiff)
ggsave("FigS14_anoxic_plate_count_scatterplot_20260820.pdf", dpi = 300, height=16, width=20, device=cairo_pdf)
```


# Analysis for genetic basis of DET 

## Run Nanophase to assemble nanopore long read sequencing data on mutants

1. Make a directory to house the pre-assembled genomes and raw fasqt.gz files from the sequencing centre. Copy the files over to the target directories.
```{bash}
mkdir Pseudomonas_deletion_mutant_long_read_sequencing_data
mkdir Raw_fasta_files
mkdir Preassembled_genomes

#Copy pre-assembled genomes
scp daniel@134.117.183.97:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/FlowGenomics_data/Pca*.fasta ~/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Preassembled_genomes

#Copy raw fastq files
scp daniel@134.117.183.97:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/FlowGenomics_data/4074f4bc/fastqs/*.fastq.gz ~/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Raw_fasta_files
```

2. Start a screen, navigate to the proper directory, and activate nanophase to run the assembly in each raw fastq.gz file.
```{bash}
tmux new -s nanophase_run #you can change this name to whatever you want
cd ~/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Raw_fasta_files
conda activate nanophase

#Wild type
nanophase isolate -l Pca-WT.fastq.gz -t 40 -o ~/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_WT_nanophase_assembly

#∆phz deletion mutant
nanophase isolate -l Pca-phz.fastq.gz -t 40 -o ~/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_phz_nanophase_assembly

#∆hcn deletion mutant
nanophase isolate -l Pca-hcn.fastq.gz -t 40 -o ~/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_hcn_nanophase_assembly
```

3. Rename files (nanophase provides generic names). Export and copy over required files to computer.
```{bash}
cd /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_WT_nanophase_assembly/03-Polishing/Final-bins

mv bin.fasta Pca_WT_nanophase.fasta

mv nanophase.ont.genome.summary Pca_WT_nanophase.ont.genome.summary

cd /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_phz_nanophase_assembly/03-Polishing/Final-bins

mv bin.fasta Pca_phz_nanophase.fasta

mv nanophase.ont.genome.summary Pca_phz_nanophase.ont.genome.summary

cd /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_hcn_nanophase_assembly/03-Polishing/Final-bins

mv bin.fasta Pca_hcn_nanophase.fasta

mv nanophase.ont.genome.summary Pca_hcn_nanophase.ont.genome.summary

scp /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_WT_nanophase_assembly/03-Polishing/Pca_WT_nanophase.ont.genome.summary daniel@134.117.183.120:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/

scp /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_phz_nanophase_assembly/03-Polishing/Pca_phz_nanophase.ont.genome.summary daniel@134.117.183.120:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/

scp /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_hcn_nanophase_assembly/03-Polishing/Pca_hcn_nanophase.ont.genome.summary daniel@134.117.183.120:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/
```


## Re-run DRAM on the pre-assembled and nanophase assembled genomes

1. Copy all of the input .fasta files to a dedicated directory to facilitate running DRAM.
```{bash}
cd /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data

mkdir DRAM_analyses
mkdir DRAM_analyses/Input_genomes

cp /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Preassembled_genomes/*.fasta /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/DRAM_analyses/Input_genomes

cp /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_WT_nanophase_assembly/03-Polishing/Final-bins/*.fasta /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/DRAM_analyses/Input_genomes

cp /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_phz_nanophase_assembly/03-Polishing/Final-bins/*.fasta /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/DRAM_analyses/Input_genomes

cp /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_hcn_nanophase_assembly/03-Polishing/Final-bins/*.fasta /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/DRAM_analyses/Input_genomes
```

2. Create a screen to run DRAM in and activate the conda environment. Run DRAM with the --use_uniref flag.
```{bash}
cd /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/DRAM_analyses

tmux new -s Pseudomonas_DRAM
conda activate DRAM_conda
DRAM.py annotate -i 'Input_genomes/*.fasta' -o Pseudomonas_genome_DRAM_annotation_uniref90_20250916 --use_uniref --threads 40
```

3. Create the distillate to obtain all of the summary files.
```{bash}
DRAM.py distill -i Pseudomonas_genome_DRAM_annotation_uniref90_20250916/annotations.tsv -o Pseudomonas_genome_DRAM_annotation_uniref90_20250916/Pseudomonas_genome_DRAM_genome_summaries_20250916 --trna_path Pseudomonas_genome_DRAM_annotation_uniref90_20250916/trnas.tsv --rrna_path Pseudomonas_genome_DRAM_annotation_uniref90_20250916/rrnas.tsv
```

4. Copy the whole directory over to your computer.
```{bash}
scp -r /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/DRAM_analyses daniel@134.117.183.120:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/
```

## Verify long-read assemblies for phz and hcn genes.

1. Load the annotation file for all genomes output by DRAM.
```{r}
#Change the directory to load the data
setwd("/Users/daniel/Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/DRAM_analyses/Pseudomonas_genome_DRAM_annotation_uniref90_20250916")
Pseudomonas_long_read_genome_annotation <- fread("annotations.tsv", sep = "\t", header = "auto")
```

2. Separate each genome into a separate annotation data frame that includes the genome from the sequencing centre and the ones asembled with nanophase.
```{r}
Pca_WT_genomes <- Pseudomonas_long_read_genome_annotation %>% filter(fasta == "Pca-WT" | fasta == "Pca_WT_nanophase")
Pca_phz_genomes <- Pseudomonas_long_read_genome_annotation %>% filter(fasta == "Pca-phz" | fasta == "Pca_phz_nanophase")
Pca_hcn_genomes <- Pseudomonas_long_read_genome_annotation %>% filter(fasta == "Pca-hcn" | fasta == "Pca_hcn_nanophase")
```

3. Manually search the expected gene positions for the phz and hcn genes in Proksee (separate figures manually downloaded).

## Figure 4: Corrosion gene annotation

UniProt KB entries were manually collected by using the literature to identify genes directly involved in direct electron transfer in Geobacter metallireducens GS15, Geobacter sulfurreducens PCA (by way of gene deletion mutants using this as the parent strain), Methanosarcina acetivorans C2A (by way of gene deletion mutants using this as the parent strain), and Shewanella oneidensis MR-1. 

The UniProtKB accession numbers were then used to create a concatanated .fasta file that would be used as a custom annotation database with DRAM. This annotation run would apply to the Pseudomonas chlororaphis subspecieis aureofaciens genome obtained via long-read sequencing alongside the strains noted above to ensure that the genes that should be present in the parent genomes are indeed there. Genomes for the model organisms noted above were manually retrieved by searching for the organism names in KEGG and clicking through links to obtain the raw accession numbers for download via NCBI.

1. To create the gene database, use the UniProtKB website to do a batch download, copy and pasting the raw accession number collected in the associated spreadsheet that was used to collect these data. Download the associated metadata manually as well.

This provided *47 unique entries* which were exported as a .fasta file using the amino acid alphabet.

2. Create a dedicated directory to house all of the input genomes and the data base file. Copy over all source files required from your computer (code removed for privacy)
```{bash}
mkdir Model_corroder_genomes
scp daniel@134.117.183.226:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/Model_corroder_genomes/*.fna ~/Model_corroder_genomes

scp daniel@134.117.183.226:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/Model_corroder_genomes/*.fasta ~/Model_corroder_genomes

scp daniel@134.117.183.226:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/Model_corroder_genomes/Corrosion_gene_database.fasta /datastore/userdata/daniel/Custom_annotation_databases/

#Rename everything to .fasta extension
for f in *.fna; do
    mv "$f" "${f%.fna}.fasta"
done
```


3. Create a screen to run the process in the background and run DRAM calling uniref90 but also providing a path to the custom database. 
```{bash}
tmux new -s Corrosion_gene_annotation
conda activate DRAM_conda

DRAM.py annotate -i 'Model_corroder_genomes/*.fasta' -o Model_corroder_annotations_uniref90_20260108 --use_uniref --custom_fasta_loc /datastore/userdata/daniel/Custom_annotation_databases/Corrosion_gene_database.fasta --custom_db_name Corrosion_genes --threads 4
```

Started 2026-01-08 at 16:15

Got an error message at 20:45: 

/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/database_handler.py:218: UserWarning: No descriptions were found for your id's. Does this GT2_Glycos_transf_2 look like an id from dbcan_description           
  warnings.warn(
/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/database_handler.py:218: UserWarning: No descriptions were found for your id's. Does this GT2_Glyco_tranf_2_3 look like an id from dbcan_description           
  warnings.warn(
2026-01-08 18:05:53,932 - Getting hits from Corrosion_genes
2026-01-08 18:05:53,933 - Getting forward best hits from Corrosion_genes
2026-01-08 18:06:08,313 - Getting reverse best hits from Corrosion_genes
2026-01-08 18:06:08,549 - Getting descriptions of hits from Corrosion_genes
Traceback (most recent call last):
  File "/datastore/tools/gregoire/envs/DRAM_conda/bin/DRAM.py", line 211, in <module>
    args.func(**args_dict)
  File "/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/annotate_bins.py", line 1531, in annotate_bins                                                                                                         
    all_annotations = annotate_fastas(
  File "/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/annotate_bins.py", line 1410, in annotate_fastas                                                                                                       
    annotate_fasta(
  File "/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/annotate_bins.py", line 1255, in annotate_fasta                                                                                                        
    annotations = annotate_orfs(
  File "/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/annotate_bins.py", line 1159, in annotate_orfs                                                                                                         
    do_blast_style_search(
  File "/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/annotate_bins.py", line 859, in do_blast_style_search                                                                                                  
    hits = formater(hits, header_dict)
  File "/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/annotate_bins.py", line 181, in get_basic_description                                                                                                  
    header = header_dict[hit]
KeyError: 'Q8EBS0'

May need to clean up the leading text in the headers to facilitate look up.

4. Tidy up headers in a new file and try again.
```{bash}
#Clean up sp and tr leading the headers
sed -E 's/^>(sp|tr)\|/>/' Corrosion_gene_database.fasta > Corrosion_gene_database_cleaned.fasta

DRAM.py annotate -i 'Model_corroder_genomes/*.fasta' -o Model_corroder_annotations_uniref90_20260108 --use_uniref --custom_fasta_loc /datastore/userdata/daniel/Custom_annotation_databases/Corrosion_gene_database_cleaned.fasta --custom_db_name Corrosion_genes --threads 4
```

5. Run the distillation command for an additional summary (this file won't be used in the final analyses, but is helpful for other applications).
```{bash}
DRAM.py distill -i ~/Model_corroder_annotations_uniref90_20260108/annotations.tsv -o ~/Model_corroder_annotations_uniref90_20260108/genome_summaries --trna_path ~/Model_corroder_annotations_uniref90_20260108/trnas.tsv --rrna_path ~/Model_corroder_annotations_uniref90_20260108/rrnas.tsv
```

6. Copy the files over to your computer to begin inspecting in R.
```{bash}
scp -r ~/Model_corroder_annotations_uniref90_20260108 daniel@134.117.183.34:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/Model_corroder_genomes/
```

7. Import the annotation data into R and trim the file to the relevant genomes and columns of interest.
```{r}
#Load the data frame from scratch to test
Pseudomonas_annotations <- fread("Supporting_data_1A_corrosion_gene_annotations.tsv", sep = "\t", header = "auto") %>% 
                           filter(fasta == "Pseudomonas_chlororaphis_aureofaciens_WT") 

write.table(Pseudomonas_annotations, file = "Supporting_data_1_Pseudomonas_chlororaphis_subsp_aureofaciens_genome_annotation_20260129.txt", sep = "\t", 
            row.names = FALSE, col.names = TRUE) 

Corrosion_gene_annotations <- fread("Supporting_data_1A_corrosion_gene_annotations.tsv", sep = "\t", header = "auto") %>% 
                              filter(!fasta == "Cupriavidus_metallidurans_CH34") %>%
                              dplyr::select(c(fasta:uniref_eVal, Corrosion_genes_id:heme_regulatory_motif_count)) %>%
                              filter(!Corrosion_genes_id == "") %>%
                              mutate(Gene_name = str_extract(Corrosion_genes_hit, "(?<=GN=).*?(?=\\sPE=)")) %>%
                              relocate(Gene_name, .after = uniref_eVal)

write.table(Corrosion_gene_annotations, file = "Supporting_data_1A_Corrosion_genome_annotations_20260129.txt", sep = "\t", 
            row.names = FALSE, col.names = TRUE)

#Quickly inspect density plots for bit score and percent identity

ggplot(Corrosion_gene_annotations, aes(x = Corrosion_genes_identity)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  labs(
    title = "Density Plot of Corrosion_genes_identity",
    x = "Corrosion genes identity",
    y = "Density"
  ) +
  theme_minimal()

ggplot(Corrosion_gene_annotations, aes(x = log10(Corrosion_genes_bitScore))) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  labs(
    title = "Density Plot of Corrosion_genes_identity",
    x = "Corrosion genes identity",
    y = "Density"
  ) +
  theme_minimal()
```
This looks like setting a simple bitScore cutoff of 100 would get rid of a lot of the hits that align over a small portion of the gene (i.e., < 50%)

8. Filter the data to group by strain, and then corrosion gene id, and then take the highest similarity scores for each one.
```{r}
Corrosion_gene_annotations_max_similarity <- Corrosion_gene_annotations %>%
                                             group_by(fasta, Corrosion_genes_id) %>%
                                             slice_max(Corrosion_genes_identity, n = 1, with_ties = FALSE) %>%
                                             ungroup() #87 observations

ggplot(Corrosion_gene_annotations_max_similarity, aes(x = Corrosion_genes_identity)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  labs(
    title = "Density Plot of Corrosion_genes_identity",
    x = "Corrosion genes identity",
    y = "Density"
  ) +
  theme_minimal()

ggplot(Corrosion_gene_annotations_max_similarity, aes(x = log10(Corrosion_genes_bitScore))) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  labs(
    title = "Density Plot of Corrosion_genes_identity",
    x = "Corrosion genes identity",
    y = "Density"
  ) +
  theme_minimal()
```

9. Plot a heat map that uses a custom order for the genomes and shows the percent homology.
```{r}
custom_order <- c(
  "*Shewanella oneidensis MR1*",
  "*Methanosarcina acetivorans C2A*",
  "*Geobacter sulfurreducens PCA*",
  "*Geobacter metallireducens GS15*",
  "*Pseudomonas chlororaphis aureofaciens WT*"
  )


Corrosion_gene_heatmap <- Corrosion_gene_annotations_max_similarity %>%
                          filter(!Gene_name == "recA") %>%
                          mutate(Corrosion_genes_identity_percent = Corrosion_genes_identity*100) %>%
                          mutate(
                          Genome = paste0(
                            "*",
                            gsub("_", " ", fasta),
                            "*"
                          )
                        ) %>%
                        mutate(Genome_ordered = factor(Genome, levels = custom_order)) %>%
                          ggplot(aes(x = Gene_name, y = Genome_ordered, fill = Corrosion_genes_identity_percent)) +
                          geom_tile(color = "grey80") +
                          scale_fill_viridis_c(
                            option = "viridis",
                            na.value = "white"
                          ) +
                          theme_minimal() +
                          theme(
                            panel.grid = element_blank(),            # remove grid lines
                            axis.text.x = element_text(
                              angle = 90,
                              vjust = 0.5,
                              hjust = 1,
                              size = 20
                            ),
                            axis.text.y = element_markdown(size = 20),
                            legend.text = element_text(size = 22),
                            legend.title = element_text(size = 24, margin = margin(b = 20))
                          ) +
                          labs(
                            x = NULL,
                            y = NULL,
                            fill = "% similarity"
                          )

Corrosion_gene_heatmap
ggsave("Fig4_Corrosion_gene_annotation_heatmap_20260820.jpg", 
       dpi = 300, height=6, width=22, device= jpeg)
ggsave("Fig4_Corrosion_gene_annotation_heatmap_20260820.tiff", 
       dpi = 300, height=6, width=22, device= tiff)
ggsave("Fig4_Corrosion_gene_annotation_heatmap_20260820.pdf", 
       dpi = 300, height=6, width=22, device= cairo_pdf)
```


## Electrochemical measurements

#Figure 7: Cyclic voltammetry for abiotic and biotic treatments over 72 hours

1. Load the CV data and inspect.
```{r}
CV_data <- read_excel("Compiled_electrochem_data_All_2026-06-15.xlsx", sheet = "CV")
```

2. Plot the data using faceting. In this case, we want to try and colour code the time points on the facet of an individual replicate so that we're seeing all the same traces in the same area. 
```{r}
CV_plot <- CV_data %>%
           mutate(Treatment = factor(Treatment, levels = c("Sterile", "PCA", "Spent", "Wild type"))) %>% #reorder the variables
            ggplot(
              aes(
                x = Voltage_mV,
                y = Current_µA,
                colour =Time_point_hours)
              ) +
            geom_point(size = 0.5,
                      ) +
            facet_grid(Treatment ~ Replicate) +
            scale_colour_gradient(low = "blue",
                                  high = "red"
                                ) +
            labs(title = "Cyclic voltammetry measurements in abiotic vs biotic treatments",
                 x = "Voltage (mV)",
                 y = "Current (µA)",
                 colour = "Time (hours)") +
            theme_classic(base_size = 15) +  # 11 + 4
            theme(
              panel.border = element_rect(
                colour = "black",
                fill = NA,
                linewidth = 1
              )
            )

CV_plot

ggsave("CV_plot_zoomed_out.pdf", width = 10, height = 8, device = pdf)
```

3. Make the CV plot but this time zoom into the relevant potential range for the Co(2+) + 2e- --> Co(0)(s) reaction voltage on the x-axis.
```{r}
CV_plot_zoomed_in <- CV_data %>%
                     filter(between(Voltage_mV, -0.7, 0)) %>%
                     filter(Time_point_hours == "0" | Time_point_hours == "24" | Time_point_hours == "72") %>%
           mutate(Treatment = factor(Treatment, levels = c("Sterile", "PCA", "Spent", "Wild type"))) %>% #reorder the variables
            ggplot(
              aes(
                x = Voltage_mV,
                y = Current_µA,
                colour =as.factor(Time_point_hours))
              ) +
            geom_point(size = 0.5) +
            ##Dashed line for SHE
            geom_vline(
                xintercept = -0.282,
                linetype = "dashed",
                linewidth = 0.8,
                colour = "black",
                linetype = "SHE"
              ) +
  
            ##Dashed line for Ag electrode
            geom_vline(
                xintercept = -0.282 - 0.197, #correct for saturated silver electrode vs SHE
                linetype = "dashed",
                linewidth = 0.8,
                colour = "darkgrey",
                linetype = "Ag/AgCl"
              ) +
            facet_grid(Treatment ~ Replicate) +
            scale_colour_manual(
                values = c("0" = "blue", "24" = "purple", "72" = "red"),
                guide = guide_legend(override.aes = list(size = 4))) +
  
            scale_linetype_manual(
                name = "Reference electrode",
                values = c(
                  "SHE" = "dashed",
                  "Ag/AgCl" = "dashed"
                )
              ) +
            labs(#title = "Cyclic voltammetry measurements in abiotic vs biotic treatments",
                 x = "Voltage (V)",
                 y = "Current (µA)",
                 colour = "Time (hours)") +
            theme_classic(base_size = 15) +  # 11 + 4
            theme(
              panel.border = element_rect(
                colour = "black",
                fill = NA,
                linewidth = 1
              )
            ) +
            annotate("text", x = -0.282, y = Inf, label = "SHE", vjust = 2, hjust = -0.5) +
            annotate("text", x = -0.497, y = Inf, label = "Ag/AgCl", vjust = 2, hjust = -0.25, color = "darkgrey")

CV_plot_zoomed_in
ggsave("Fig7_CV_plot_zoomed_in_20260820.pdf", width = 14, height = 8, device = pdf)
ggsave("Fig7_CV_plot_zoomed_in_20260820.jpg", width = 14, height = 8, device = jpeg)
ggsave("Fig7_CV_plot_zoomed_in_20260820.tiff", width = 14, height = 8, device = tiff)

```


#Figure S20: Open circuit potential for abiotic and biotic treatments over 72 hours

1. Read in the sheet on the excel file containing open circuit potential data.
```{r}
OCP_data <- read_excel("Compiled_electrochem_data_All_2026-06-15.xlsx", sheet = "OCP")
```

2. Generate a plot that shows all of the replicates for each treatment on the same facet. Generate bars that represent the min and max values for each 10 min interval that was scanned and use the median to present a single point to connect the data series.
```{r}
OCP_plot <- OCP_data %>%
            group_by(Treatment, Replicate, Time_point_hours) %>%
            summarise(
              Median_OCP_mV = median(OCP_mV),
              Min_OCP_mV = min(OCP_mV),
              Max_OCP_mV = max(OCP_mV),
              .groups = "drop"
            ) %>%
            mutate(Treatment = factor(Treatment, levels = c("Sterile", "PCA", "Spent", "Wild type"))) %>% #reorder the variables 
            ggplot(
              aes(
                x = Time_point_hours,
                y = Median_OCP_mV,
                colour = as.factor(Replicate),
                group = Replicate
              )
            ) +
            geom_errorbar(
              aes(ymin = Min_OCP_mV, ymax = Max_OCP_mV),
              width = 0.1
            ) +
            geom_line(linewidth = 1) +
            geom_point(
              aes(fill = as.factor(Replicate)),
              shape = 21,
              size = 4,
              colour = "black",
              stroke = 1
            ) +
            facet_grid(~Treatment) +
            labs(x = "Time (hours)",
                 y = "Open circuit potential (mV)",
                 fill = "Replicate") +
            theme_classic(base_size = 15) +  # 11 + 4
            theme(
              panel.border = element_rect(
                colour = "black",
                fill = NA,
                linewidth = 1
              )
            ) + guides(colour = "none")

OCP_plot
```

3. Adjust the code slightly to facet the graphs to show each individual replicate to tidy things up.
```{r}
OCP_plot_facet <- OCP_data %>%
            group_by(Treatment, Replicate, Time_point_hours) %>%
            summarise(
              Median_OCP_mV = median(OCP_mV),
              Min_OCP_mV = min(OCP_mV),
              Max_OCP_mV = max(OCP_mV),
              .groups = "drop"
            ) %>%
            mutate(Treatment = factor(Treatment, levels = c("Sterile", "PCA", "Spent", "Wild type"))) %>% #reorder the variables
            ggplot(
              aes(
                x = Time_point_hours,
                y = Median_OCP_mV,
                group = Replicate
              )
            ) +
            geom_errorbar(
              aes(ymin = Min_OCP_mV, ymax = Max_OCP_mV),
              width = 0.1
            ) +
            geom_line(linewidth = 1) +
            geom_point(shape = 21,
                       size = 4,
                       fill = "white",
                       colour = "black",
                       stroke = 1
                      ) +
            facet_grid(Treatment ~ Replicate) +
            labs(x = "Time (hours)",
                 y = "Open circuit potential (mV)",
                 colour = "Replicate") +
            theme_classic(base_size = 15) +  # 11 + 4
            theme(
              panel.border = element_rect(
                colour = "black",
                fill = NA,
                linewidth = 1
              )
            )

OCP_plot_facet
ggsave("FigS20_OCP_plot_individual_replicates_20260820.jpg", dpi = 300, width = 8, height = 8, device= jpeg)
ggsave("FigS20_OCP_plot_individual_replicates_20260820.tiff", dpi = 300, width = 8, height = 8, device= tiff)
ggsave("FigS20_OCP_plot_individual_replicates_20260820.pdf", dpi = 300, width = 8, height = 8, device = "pdf")
```

#Figure S21: [Co(2+)] over time in abiotic and biotic treatments over 72 hours

1. Use the code below to load in source data.
```{r}
#Load the data frame from scratch to test
Electrochem_cobalt_data_combined_revised <- fread("Electrochem_cobalt_ICPMS_data_cleaned_formatted_revised_mass_balance_revised_20260710.txt", sep = "\t", header = "auto")
```

2. Use the graphing script and call the *corrected* cobalt concentration in this case and save it as new dedicated files.
```{r}
#Adjust the presentation aesthetics
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
        size = 34, angle = 0, hjust = .5, vjust = 0, face = "bold"),
        axis.title.y = element_text(color = "grey20", 
        size = 34, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 32, colour = "grey20", face = "bold"),
        strip.text.y = element_text(size = 32, colour = "grey20"),
        plot.title = element_text(size = 30, face = "bold", hjust = .5),
        legend.title=element_text(size=32, face = "bold"), 
        legend.text=element_text(size=32),
        panel.border = element_rect(colour = "black", fill=NA, size=1))

dark2_colors <- brewer.pal(n = 4, name = "Dark2") #make palette with desired levels

#Create custom display orders
Electrochem_cobalt_data_combined_revised$Abiotic_vs_biotic_ordered <- factor(Electrochem_cobalt_data_combined_revised$Abiotic_vs_biotic,
                                                             levels = c("Wild type",
                                                                        "Spent medium",
                                                                        "PCA", 
                                                                        "Sterile medium"))

Electrochem_cobalt_data_scatterplot_recoveries <- Electrochem_cobalt_data_combined_revised %>%
                                           ggplot(aes(x = Time, 
                                                      y = Cobalt_concentration_uM_corrected, 
                                                      fill = Abiotic_vs_biotic_ordered)) +
                                                  geom_point(shape = 21, color = "black", stroke = 0.8, size = 6) +
                                                  geom_line(aes(y = Cobalt_concentration_uM_corrected, 
                                                              group = interaction(Abiotic_vs_biotic_ordered, Replicate),
                                                              color = Abiotic_vs_biotic_ordered)) +
                                                  scale_x_continuous(breaks = seq(0, 168, by = 24)) +
                                                  ylab("Co(2+) concentration (µM)\nrecoveries applied") +
                                                  xlab("Time elapsed (hours)") + 
                                                  scale_fill_manual(values = dark2_colors,
                                                                    name = "Treatment", 
                                                                    labels = c("Wild type",
                                                                        "Spent medium",
                                                                        "PCA", 
                                                                        "Sterile medium")) +
  
                                                     scale_colour_manual(values = dark2_colors,
                                                                    name = "Treatment", 
                                                                    labels = c("Wild type",
                                                                        "Spent medium",
                                                                        "PCA", 
                                                                        "Sterile medium")) +
                                                  theme +
                                                  theme(plot.margin = unit(c(1,1,1,1), "cm"),
                                                        axis.title.x = element_text(vjust = -2),
                                                        axis.title.y = element_text(vjust = 2))

#Save in different image formats at 300 dpi
ggsave("FigSX_Cobalt_ICPMS_data_electrochem_scatterplot_angled_recoveries_applied_20260820.jpg", 
       dpi = 300, height=10, width=20, device= jpeg)
ggsave("FigSX_Cobalt_ICPMS_data_electrochem_scatterplot_angled_recoveries_applied_20260820.tiff", 
       dpi = 300, height=10, width=20, device= tiff)
ggsave("FigSX_Cobalt_ICPMS_data_electrochem_scatterplot_angled_recoveries_applied_20260820.pdf", 
       dpi = 300, height=8, width=22, device= cairo_pdf)

Electrochem_cobalt_data_combined_revised$Abiotic_vs_biotic_ordered <- factor(Electrochem_cobalt_data_combined_revised$Abiotic_vs_biotic,
                                                             levels = c("Sterile medium",
                                                                        "PCA",
                                                                        "Spent medium", 
                                                                        "Wild type"))

Electrochem_cobalt_data_scatterplot_recoveries <- Electrochem_cobalt_data_combined_revised %>%
                                           mutate(Replicate = factor(Replicate)) %>% 
                                           ggplot(aes(x = Time, 
                                                      y = Cobalt_concentration_uM_corrected, 
                                                      #shape = factor(Replicate), 
                                                      fill = Replicate)) +
                                                  geom_line(aes(group = interaction(Abiotic_vs_biotic_ordered, 
                                                                                    Replicate), 
                                                                colour = Replicate)) +
                                                  geom_point(shape = 21, colour = "black", 
                                                             stroke = 0.8, size = 6) +
                                                  #scale_shape_manual(values = c("1" = 21, "2" = 22, "3" = 24),
                                                                     #name = "Replicate") +
                                                  #scale_fill_manual(values = c("1" = "#1B9E77", 
                                                                               #"2" = "#D95F02",
                                                                               #"3" = "#7570B3"), name = "Replicate") +
                                                  #scale_colour_manual(values = c("1" = "#1B9E77",
                                                                                 #"2" = "#D95F02",
                                                                                 #"3" = "#7570B3"), name = "Replicate") +
                                                  scale_x_continuous(breaks = seq(0, 168, by = 24)) +
                                                  facet_grid(~ Abiotic_vs_biotic_ordered) +
                                                  ylab("Co(2+) concentration (µM)\nrecoveries applied") +
                                                  xlab("Time elapsed (hours)") +
                                                  theme +
                                                  theme(plot.margin = unit(c(1,1,1,1), "cm"),
                                                        axis.title.x = element_text(vjust = -2),
                                                        axis.title.y = element_text(vjust = 2))

#Save in different image formats at 300 dpi
ggsave("FigSX_Cobalt_ICPMS_data_electrochem_scatterplot_replicates_angled_recoveries_applied_20260820.jpg", 
       dpi = 300, height=10, width=20, device= jpeg)
ggsave("FigSX_Cobalt_ICPMS_data_electrochem_scatterplot_replicates_angled_recoveries_applied_20260820.tiff", 
       dpi = 300, height=10, width=20, device= tiff)
ggsave("FigSX_Cobalt_ICPMS_data_electrochem_scatterplot_replicates_angled_recoveries_applied_20260820.pdf", 
       dpi = 300, height=8, width=22, device= cairo_pdf)
```



# Figure S30: Stacked barplots for phenazine measurements via LC-HRMS

1. Load the formatted LC-HRMS data.
```{r}
LC_MS_data_combined_labelled_long <- fread("LC_MS_data_combined_20250419.txt", sep = "\t", header = "auto")
```

2. Plot the stacked bar graphs using variables with a custom order.
```{r}
barplot_theme<-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
                 panel.background = element_blank(), axis.line = element_line(colour = "black"),
                 axis.text.x = element_text(color = "black", size = 32),
                 axis.text.y = element_text(color = "black", size = 32),
                 axis.title.x = element_text(color = "black", size = 34, vjust = -3, face = "bold"),
                 axis.title.y = element_text(color = "black", size = 34, vjust = 6, face = "bold"),
                 strip.text.x = element_text(color = "black", size = 32, face = "bold"),
                 strip.text.y = element_text(color = "black", size = 32, face = "bold"),
                 legend.title=element_text(size=40, vjust = 2, face = "bold"), 
                 legend.text=element_text(size=34),
                 legend.key = element_rect(fill = "white", color = NA),
                 plot.title = element_text(size = 68, face = "bold", hjust = 0.5),
                 panel.border = element_rect(colour = "black", fill=NA, size=1),
                 plot.margin = unit(c(2, 2, 2, 2), "cm"))

LC_MS_data_combined_labelled_long$Abiotic_vs_biotic_ordered <- factor(LC_MS_data_combined_labelled_long$Abiotic_vs_biotic, 
                                                                      levels = c("Wild type",
                                                                            "Spent medium",
                                                                            "∆phz"))

LC_MS_data_combined_labelled_long$Metabolite_ordered <- factor(LC_MS_data_combined_labelled_long$Metabolite, 
                                                                      levels = c("2-OHPCA",
                                                                            "PCA",
                                                                            "2-OHPHZ"))


Phenazine_barplot <-   LC_MS_data_combined_labelled_long %>%
                       group_by(Comment) %>%
                       mutate(Mean_concentration_ug_mL = mean(Concentration_ug_mL)) %>%
                       ggplot(aes(x= Time, y = Concentration_ug_mL, fill = Metabolite_ordered)) +
                       geom_bar(aes(group = Replicate), position="stack", stat="identity") +
                       facet_nested(Replicate~Wire_treatment + Abiotic_vs_biotic_ordered)+
                       scale_x_continuous(breaks = seq(0, 168, by = 24)) +
                       scale_fill_manual(name = "Metabolite", 
                                         values = c("2-OHPCA" = "#FA1F1E",   # Red
                                                    "PCA" = "#EBA602",       # Gold
                                                    "2-OHPHZ" = "#027AB0"    # Blue
                                                    ))+
                       #scale_fill_brewer(name = "Metabolite", palette = "YlOrRd")+ 
                       ylim(0,15) +
                       theme(axis.text.x=element_text(angle=0, hjust=0.5),
                       axis.title.y = element_text(color = "black", size = 24, vjust = 0))+
                       labs(x = "Time elapsed (hours)" , y = "Concentration (µg/mL)") +
                       barplot_theme + 
                       theme(legend.key.size = unit(2, 'cm'))

Phenazine_barplot

Phenazine_barplot_angled <- Phenazine_barplot + theme(axis.text.x = element_text(angle = 45, vjust = 0.5))

#Save in different image formats
ggsave("FigS31_phenazine_LCMS_data_stacked_barplot_faceted_20260820.jpg", dpi = 300, height=12, width=40, device=jpeg)
ggsave("FigS31_phenazine_LCMS_data_stacked_barplot_faceted_20260820.tiff", dpi = 300, height=12, width=40, device=tiff)
ggsave("FigS31_phenazine_LCMS_data_stacked_barplot_faceted_20260820.pdf", dpi = 300, height=12, width=40, device=cairo_pdf)
```