---
title: "Cobalt corrosion Github codebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Table of contents

1) Additional ICP-MS data, mass balances, and statistical tests conducted

- Figure 1A: Cobalt corrosion data for wild type cells, deletion mutants, spent medium, and abiotic controls in sterile medium
- Figure 1B: Boxplot of log10 [Co(2+)] at 168 hours across abiotic and biotic treatments
- Table S1 (partial): Mass balance recoveries
- Figure S1: Contribution of background Co(2+) in medium from "no wire" controls
- Figure S2: Boxplot of [Co(2+)] at 168 hours for abiotic metabolite treatments

2) Biosynthetic gene cluster analyses

- Bioinformatics analyses of genome annotations generated with DRAM (used to make Figure S3, Table S2, and Table S3)

3) Cell-associated Co(2+) and toxicity measurements 

- Figure 2: Wild type vs deletion cell growth in corrosion assays as measured by viable plate counts
- Figure S4: Comparison of [Co(2+)] between A-F and F-A preservation for wild type cells
- Figure S5: log(µmol Co(2+)/CFU) over Co(2+) exposure from spike concentrations in corrosion assays compared to dose-response experiments (processed in Excel)
- Figure S6. Cobalt dose-response curve reporting growth rate of P. chlororaphis subsp. aureofaciens in medium amended with CoCl2
- Figure S7: Boxplots of [Co(2+)] between A-F and F-A preservation in a Co(2+) dose-response curve with wild type cells.
-	Figure S8: Comparison of [Co(2+)] between acidify-filter and filter-acidify preservation with sterile medium

4) Additional SEM images, wire diameters, and EDX data 

- These are largely image files used in Figure 4, Table S4, FigureS9 to S19, and Table S5 that were not analyzed in R. 

5) Statistical analyses of ICP-MS data and SEM of anoxic and DMSO treatments 

- Figure 4B: Boxplot of [Co(2+)] at 168 hours comparing oxic and anoxic treatments 
-	Figure S20: Boxplot of [Co(2+)] at 168 hours across DMSO biotic treatments

6) Redox reaction comparison between Fe & Co 

- These are theoretical calculations summarized in Table S6 that were not performed in R

7) Long-read sequence genome validation

- Example bash code used to assemble long read data using Nanophase
- Example bash code used to anontate genomes with DRAM
- Fig S22: Corrosion gene annotation

8) Genetic manipulations and validation 

- Examples of nanophase code used in genome assembly are included here
- All other figures and alignment files were performed outside of R through other open source software

9) LC-HRMS QA/QC 

- Example chromatograms, m/z values, and retetion times that provided calibration curves were processed manually rather than in R
- Figure S31: Stacked barplots for phenazine measurements via LC-HRMS

# Packages required

Load all of these packages before running any of the chunks of code in the subsections below.
```{r}
library(tidyverse)
library(readxl)
library(data.table)
library(plotly)
library(Cairo)
library(lmtest)
library(car)
library(wesanderson)
library(ggbreak)
library(scales)
library(ggh4x)
library(ggtext)
library(drc)
library(RColorBrewer)
library(forcats)
library(markdown)
library(patchwork)
```

# Figure 1A: Cobalt corrosion data for wild type cells, deletion mutants, spent medium, and abiotic controls in sterile medium

1. Use the code below to load in source data.
```{r}
#Load the data frame from scratch to test
All_cobalt_data_combined_revised <- fread("All_cobalt_ICPMS_data_cleaned_formatted_revised_mass_balance_summaries_included_20251001.txt", sep = "\t", header = "auto")
```

2. Summarize the mass balance data.
```{r}
#Summarize the mass balance data
Mass_balance_data_summary <- All_cobalt_data_combined_revised %>%
                             filter(Wire_treatment == "Wire included") %>% #only include wire treatments
                             group_by(Abiotic_vs_biotic, Wire_treatment, DMSO_treatment, Metabolite_treatment, Sample_preservation, Time) %>%
                             mutate(Mean_recovery = mean(Recovery),
                                    Median_recovery = median(Recovery),
                                    Stdev_recovery = sd(Recovery),
                                    Min_recovery = min(Recovery),
                                    Max_recovery = max(Recovery)) %>%
                             filter(Wire_treatment == "Wire included" & Time == 168 & Replicate == 1) %>%
                             dplyr::select(c(Abiotic_vs_biotic, DMSO_treatment, Metabolite_treatment, Sample_preservation, Mean_recovery:Max_recovery))

write.table(Mass_balance_data_summary, file = "TableS1_Mass_balance_data_summary_20251009.txt", sep = "\t", 
            row.names = FALSE, col.names = TRUE) 

Mass_balance_data_summary_count <- All_cobalt_data_combined_revised %>%
                                   count(Abiotic_vs_biotic, Wire_treatment, DMSO_treatment, Metabolite_treatment, Sample_preservation, Time)
```

3. Use the graphing script and call the *corrected* cobalt concentration in this case and save it as new dedicated files.
```{r}
#Adjust the presentation aesthetics
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
        size = 34, angle = 0, hjust = .5, vjust = 0, face = "bold"),
        axis.title.y = element_text(color = "grey20", 
        size = 34, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 32, colour = "grey20", face = "bold"),
        strip.text.y = element_text(size = 32, colour = "grey20"),
        plot.title = element_text(size = 30, face = "bold", hjust = .5),
        legend.title=element_text(size=32, face = "bold"), 
        legend.text=element_text(size=32),
        panel.border = element_rect(colour = "black", fill=NA, size=1))

dark2_colors <- brewer.pal(n = 7, name = "Dark2") #make palette with desired levels

#Create custom display orders
All_cobalt_data_combined_revised$Abiotic_vs_biotic_ordered <- factor(All_cobalt_data_combined_revised$Abiotic_vs_biotic,
                                                             levels = c("Wild type",
                                                                        "∆phz",
                                                                        "∆hcn",
                                                                        "∆phz∆hcn",
                                                                        "Spent medium",
                                                                        "Spent medium from ∆phz",
                                                                        "Sterile medium"))

All_cobalt_data_combined_revised$Metabolite_treatment_ordered <- factor(All_cobalt_data_combined_revised$Metabolite_treatment,
                                                                 levels = c("Nothing added",
                                                                            "PCA",
                                                                            "MeOH"))

Abiotic_vs_biotic_treatment_scatterplot_recoveries <- All_cobalt_data_combined_revised %>%
                                           filter(Wire_treatment == "Wire included") %>% #filter to wire only experiments
                                           filter(DMSO_treatment == "No DMSO") %>% #use this to only focus on experiments that did not receive DMSO
                                           filter(Sample_preservation == "Acidified, filtered") %>% #216 observations
                                           filter(!Abiotic_vs_biotic == "Sterile medium, anoxic" & !Abiotic_vs_biotic == "Spent medium from ∆phz, anoxic") %>%
                                           filter(Metabolite_treatment_ordered == "Nothing added") %>% #remove all of the other metabolite treatments

                                          #plotting script
            
                                           ggplot(aes(x=Time, 
                                                      y = Mean_cobalt_concentration_uM_corrected, 
                                                      fill = Abiotic_vs_biotic_ordered)) +
                                                  geom_errorbar(aes(ymin = Mean_cobalt_concentration_uM_corrected - Stdev_cobalt_concentration_uM_corrected,
                                                                    ymax = Mean_cobalt_concentration_uM_corrected + Stdev_cobalt_concentration_uM_corrected,
                                                                    colour = Abiotic_vs_biotic_ordered)) +
                                                  geom_point(shape = 21, color = "black", stroke = 0.8, size = 6) +
                                                  geom_line(aes(group = Bottle_number, color = Abiotic_vs_biotic_ordered))+
                                                  scale_x_continuous(breaks = seq(0, 168, by = 24)) +
                                                  ylim(0, 6500) +
                                                  facet_grid(~ Cell_presence) +
                                                  ylab("Co(2+) concentration (µM)\nrecoveries applied") +
                                                  xlab("Time elapsed (hours)") + 
                                                  scale_fill_manual(values = dark2_colors,
                                                                    name = "Abiotic vs biotic", 
                                                                    labels = c("Wild type", 
                                                                             expression(italic("∆phz")),
                                                                             expression(italic("∆hcn")),
                                                                             expression(italic("∆phz∆hcn")),
                                                                             "Spent medium",
                                                                             expression(plain("Spent medium from ") * italic("∆phz")),
                                                                             "Sterile medium"))+
  
                                                     scale_colour_manual(values = dark2_colors,
                                                                    name = "Abiotic vs biotic", 
                                                                    labels = c("Wild type", 
                                                                             expression(italic("∆phz")),
                                                                             expression(italic("∆hcn")),
                                                                             expression(italic("∆phz∆hcn")),
                                                                             "Spent medium",
                                                                             expression(plain("Spent medium from ") * italic("∆phz")),
                                                                             "Sterile medium"))+
                                                  theme +
                                                  theme(plot.margin = unit(c(1,1,1,1), "cm"),
                                                        axis.title.x = element_text(vjust = -2),
                                                        axis.title.y = element_text(vjust = 2))
          
#Add a layer to angle the text
Abiotic_vs_biotic_treatment_scatterplot_angled_recoveries <- Abiotic_vs_biotic_treatment_scatterplot_recoveries + theme(axis.text.x = element_text(angle = 45))

#Save in different image formats at 300 dpi
ggsave("Fig1A_Cobalt_ICPMS_data_abiotic_biotic_and_metabolite_treatments_no_DMSO_scatterplot_angled_recoveries_applied_20260129.jpg", 
       dpi = 300, height=8, width=22, device= jpeg)
ggsave("Fig1A_Cobalt_ICPMS_data_abiotic_biotic_and_metabolite_treatments_no_DMSO_scatterplot_angled_recoveries_applied_20260129.tiff", 
       dpi = 300, height=8, width=22, device= tiff)
ggsave("Fig1A_Cobalt_ICPMS_data_abiotic_biotic_and_metabolite_treatments_no_DMSO_scatterplot_angled_recoveries_applied_20260129.pdf", 
       dpi = 300, height=8, width=22, device= cairo_pdf)
```
# Figure 1B: Boxplot of log10 [Co(2+)] at 168 hours across abiotic and biotic treatments

1. Load in the data from the manually adjusted tab-delimited file that contains recoveries. Generate a new variable where the recovery has been applied.
```{r}
All_cobalt_data_combined_revised_recoveries <- fread("All_cobalt_ICPMS_data_cleaned_formatted_revised_mass_balance_summaries_included_20251001.txt", sep = "\t", header = "auto")
```

2. Run the model applying the same outlier removal and log-transformations used on the data that was not corrected for recovery.
```{r}
Abiotic_vs_biotic_wire_only_ANOVA_data_outlier_removed_recoveries <- All_cobalt_data_combined_revised_recoveries %>%
                                                                     filter(DMSO_treatment == "No DMSO") %>% 
                                                                     filter(Sample_preservation == "Acidified, filtered") %>% 
                                                                     filter(Time == 168) %>%
                                                                     filter(Metabolite_treatment == "Nothing added") %>%
                                                                     filter(Wire_treatment == "Wire included") %>%
                                                                     filter(!Abiotic_vs_biotic == "Sterile medium, anoxic" & 
                                                                            !Abiotic_vs_biotic == "Spent medium from ∆phz, anoxic") #remove anoxic

##Generate model
Abiotic_vs_biotic_model_log10_outlier_removed_recoveries <- lm(log10(Cobalt_concentration_uM_corrected) ~ Abiotic_vs_biotic, 
                                                            data = Abiotic_vs_biotic_wire_only_ANOVA_data_outlier_removed_recoveries, 
                                                            contrasts = list(Abiotic_vs_biotic = contr.sum), type =3)

Abiotic_vs_biotic_model_log10_outlier_removed_recoveries
par(mfrow = c(2,2))
plot(Abiotic_vs_biotic_model_log10_outlier_removed_recoveries)

#Run Breusch Pagan test for homoscedasticity
bptest(Abiotic_vs_biotic_model_log10_outlier_removed_recoveries)

#Run Shapiro-Wilk normality test
shapiro.test(Abiotic_vs_biotic_model_log10_outlier_removed_recoveries$residuals)

#Run Durbin-Watson test for autocorrelation
dwtest(Abiotic_vs_biotic_model_log10_outlier_removed_recoveries)
```

## Figure 1B ANOVA Interpretation
- Homoscedasticity looks okay as does the normal distribution of residuals.
- BP test returns a p-value of 0.2887, so we cannot reject the null hypothesis of homoscedasticity.
- SW test returns a p-value of 0.5932, so we cannot reject the null hypothesis that residuals are normally distributed.
- DW test returns a p-value of 0.1577, so we cannot reject the null hypothesis that there is no autocorrelation between data points.

**All linear model assumptions are met**

3. Generate the actual output of anova model generated with the data where the outlier removed and a log10 transformation was applied. This is returning a type 2 sum of squares because we are not longer considering the interaction having controlled for wire treatment
```{r}
Abiotic_vs_biotic_type_II_ANOVA_results_recoveries <- Anova(Abiotic_vs_biotic_model_log10_outlier_removed_recoveries, type = 3)
Abiotic_vs_biotic_type_II_ANOVA_results_recoveries
```
- F-value = 81.597
- p-value = 5.051e-16 ***

- There are significant differences in cobalt concentration at 168 hours that are attributable to the different abiotic vs biotic treatments even after correcting for recoveries.
- The p-values are different for the recovery corrected data, but both version of the data have p < 0.001.

4. Run a Tukey post-hoc test to figure out where the significant differences are.
```{r}
#First you need to convert the model to an aov object
Abiotic_vs_biotic_model_log10_outlier_removed_converted_recoveries <- aov(Abiotic_vs_biotic_model_log10_outlier_removed_recoveries)
Tukey_comparison_abiotic_vs_biotic_model_log10_outlier_removed_recoveries <- TukeyHSD(Abiotic_vs_biotic_model_log10_outlier_removed_converted_recoveries, conf.level = 0.95)
Tukey_comparison_abiotic_vs_biotic_model_log10_outlier_removed_recoveries
Tukey_comparison_abiotic_vs_biotic_data_frame_recoveries <- as.data.frame(Tukey_comparison_abiotic_vs_biotic_model_log10_outlier_removed_recoveries$Abiotic_vs_biotic)
```

## Figure 1B multiple comparison interpretation

**Significant results**
                                              diff         lwr          upr     p adj

∆phz∆hcn-∆hcn                          0.576930931  0.30784754  0.846014318 0.0000048
Spent medium-∆hcn                     -0.370403748 -0.63948714 -0.101320361 0.0027252

Sterile medium-∆hcn                   -0.768244888 -0.99692180 -0.539567981 0.0000000

∆phz∆hcn-∆phz                          0.378550119  0.10946673  0.647633506 0.0021272
Spent medium-∆phz                     -0.568784560 -0.83786795 -0.299701173 0.0000062
Spent medium from ∆phz-∆phz           -0.297767620 -0.58841090 -0.007124344 0.0419806
Sterile medium-∆phz                   -0.966625700 -1.19530261 -0.737948793 0.0000000

Spent medium-∆phz∆hcn                 -0.947334679 -1.21641807 -0.678251292 0.0000000
Spent medium from ∆phz-∆phz∆hcn       -0.676317740 -0.96696102 -0.385674463 0.0000011
Sterile medium-∆phz∆hcn               -1.345175819 -1.57385273 -1.116498912 0.0000000
Wild type-∆phz∆hcn                    -0.388218663 -0.63385706 -0.142580261 0.0005055

Sterile medium-Spent medium           -0.397841140 -0.62651805 -0.169164233 0.0001340
Wild type-Spent medium                 0.559116016  0.31347761  0.804754418 0.0000017
Sterile medium-Spent medium from ∆phz -0.668858079 -0.92255233 -0.415163829 0.0000001
Wild type-Spent medium from ∆phz       0.288099077  0.01901569  0.557182464 0.0299877
Wild type-Sterile medium               0.956957156  0.75639424  1.157520071 0.0000000

**Non-significant results**
∆phz-∆hcn                              0.198380811 -0.07070258  0.467464198 0.2601652
Spent medium from ∆phz-∆hcn           -0.099386809 -0.39003009  0.191256467 0.9263737
Wild type-∆hcn                         0.188712267 -0.05692613  0.434350669 0.2197960
Wild type-∆phz                        -0.009668544 -0.25530695  0.235969858 0.9999996
Spent medium from ∆phz-Spent medium    0.271016939 -0.01962634  0.561660216 0.0796790

5. Generate boxplots that will be used to overlay the Tukey annotations manually.
```{r}
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 16, angle = 45, hjust = 0.5, vjust = 0.55, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 16, angle = 0, hjust = 0.5, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 18, angle = 0, hjust = 0.5, vjust = -1, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 18, angle = 90, face = "plain"),
        strip.text.x = element_text(size = 20, colour = "grey20"),
        strip.text.y = element_text(size = 20, colour = "grey20"),
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5, vjust = 1),
        legend.title=element_text(size=18), 
        legend.text=element_text(size=18),
        legend.position = "right",
        panel.border = element_rect(colour = "black", fill=NA, size=1))

#Create custom order to display x-axis

All_cobalt_data_combined_revised_recoveries$Abiotic_vs_biotic_ordered <- factor(All_cobalt_data_combined_revised_recoveries$Abiotic_vs_biotic,
                                                                 levels = c("Wild type",
                                                                        "∆phz",
                                                                        "∆hcn",
                                                                        "∆phz∆hcn",
                                                                        "Spent medium",
                                                                        "Spent medium from ∆phz",
                                                                        "Sterile medium"))

Abiotic_vs_biotic_wire_only_boxplot_log10_recoveries <- All_cobalt_data_combined_revised_recoveries %>%
                                                        filter(DMSO_treatment == "No DMSO") %>% 
                                                        filter(Sample_preservation == "Acidified, filtered") %>% 
                                                        filter(Time == 168) %>%
                                                        filter(Metabolite_treatment == "Nothing added") %>%
                                                        filter(Wire_treatment == "Wire included") %>%
                                                        filter(!Abiotic_vs_biotic == "Sterile medium, anoxic" & !Abiotic_vs_biotic == "Spent medium from ∆phz, anoxic") %>%
                                                        mutate(Abiotic_vs_biotic_ordered = fct_recode(Abiotic_vs_biotic_ordered, 
                                                                                                  "Spent medium\nfrom ∆phz" = "Spent medium from ∆phz")) %>%
                                                        #Graphing script
                                                        ggplot(aes(x=Abiotic_vs_biotic_ordered, y=log10(Cobalt_concentration_uM_corrected))) + 
                                                        geom_boxplot() +
                                                        geom_jitter(fill = "grey", colour = "black", width = 0.1, pch =21, stroke =1)+
                                                        ylim(2, 4)+
                                                        labs(title = "" ,
                                                             y = "log10[Co(2+) concentration (µM)]\ncorrected for recovery", 
                                                             x = "Abiotic and biotic treatments") +
                                                        theme + #this calls the aesthetic theme variable
                                                        theme(plot.margin = unit(c(1,1,1,1), "cm"),
                                                              axis.title.x = element_text(vjust = -2),
                                                              axis.title.y = element_text(vjust = 2))

Abiotic_vs_biotic_wire_only_boxplot_log10_recoveries #annotate this manually in an image editor
ggsave("Fig1B_Cobalt_ICPMS_data_biotic_treatments_no_DMSO_boxplot_log10_recoveries_applied_20260129.jpg", dpi = 300, height=9, width=12, device=jpeg)
ggsave("Fig1B_Cobalt_ICPMS_data_biotic_treatments_no_DMSO_boxplot_log10_recoveries_applied_20260129.tiff", dpi = 300, height=9, width=12, device=tiff)
ggsave("Fig1B_Cobalt_ICPMS_data_biotic_treatments_no_DMSO_boxplot_log10_recoveries_applied_20260129.pdf", dpi = 300, height=9, width=12, device=cairo_pdf)
```

## Combining Figures 1A and 1B

1. Replot 1B with the same colour coding as 1A.
```{r}
dark2_colors <- brewer.pal(n = 7, name = "Dark2") #make palette with desired levels

theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
        size = 32, angle = 45, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
        size = 34, angle = 0, hjust = .5, vjust = 0, face = "bold"),
        axis.title.y = element_text(color = "grey20", 
        size = 34, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 32, colour = "grey20", face = "bold"),
        strip.text.y = element_text(size = 32, colour = "grey20"),
        plot.title = element_text(size = 30, face = "bold", hjust = .5),
        legend.title=element_text(size=32, face = "bold"), 
        legend.text=element_text(size=32),
        panel.border = element_rect(colour = "black", fill=NA, size=1))

Abiotic_vs_biotic_wire_only_boxplot_log10_recoveries_colour_coded <- All_cobalt_data_combined_revised_recoveries %>%
                                                        filter(DMSO_treatment == "No DMSO") %>% 
                                                        filter(Sample_preservation == "Acidified, filtered") %>% 
                                                        filter(Time == 168) %>%
                                                        filter(Metabolite_treatment == "Nothing added") %>%
                                                        filter(Wire_treatment == "Wire included") %>%
                                                        filter(!Abiotic_vs_biotic == "Sterile medium, anoxic" & !Abiotic_vs_biotic == "Spent medium from ∆phz, anoxic") %>%
                                                        mutate(Abiotic_vs_biotic_ordered = fct_recode(Abiotic_vs_biotic_ordered, 
                                                                                                  "Spent medium\nfrom ∆phz" = "Spent medium from ∆phz")) %>%
                                                        #Graphing script
                                                        ggplot(aes(x=Abiotic_vs_biotic_ordered, y=log10(Cobalt_concentration_uM_corrected), fill = Abiotic_vs_biotic_ordered)) + 
                                                        geom_boxplot() +
                                                        geom_jitter(fill = "grey", colour = "black", width = 0.1, pch =21, stroke =1)+
                                                        scale_fill_manual(values = dark2_colors) +
                                                        ylim(2, 4)+
                                                        labs(title = "" ,
                                                             y = "log10[Co(2+) concentration (µM)]\ncorrected for recovery", 
                                                             x = "Abiotic and biotic treatments") +
                                                        theme + #this calls the aesthetic theme variable
                                                        theme(plot.margin = unit(c(1,1,1,1), "cm"),
                                                              axis.title.x = element_text(vjust = -2),
                                                              axis.title.y = element_text(vjust = 2),
                                                              axis.text.x = element_text(angle = 45),
                                                              legend.position = "none",
                                                              guides(fill = "none"))

Abiotic_vs_biotic_wire_only_boxplot_log10_recoveries_colour_coded #annotate this manually in an image editor
ggsave("Fig1B_Cobalt_ICPMS_data_biotic_treatments_no_DMSO_boxplot_log10_recoveries_applied_colour_coded_20260129.jpg", dpi = 300, height=9, width=12, device=jpeg)
ggsave("Fig1B_Cobalt_ICPMS_data_biotic_treatments_no_DMSO_boxplot_log10_recoveries_applied_colour_coded_20260129.tiff", dpi = 300, height=9, width=12, device=tiff)
ggsave("Fig1B_Cobalt_ICPMS_data_biotic_treatments_no_DMSO_boxplot_log10_recoveries_applied_colour_coded_20260129.pdf", dpi = 300, height=9, width=12, device=cairo_pdf)
```

2. Use the patchwork package to combine, adjust dimensions, and aesthetics.
```{r}
Fig1A_Fig1B_combined <- (Abiotic_vs_biotic_treatment_scatterplot_angled_recoveries / Abiotic_vs_biotic_wire_only_boxplot_log10_recoveries_colour_coded) + 
                         plot_layout(ncol = 1)

Fig1A_Fig1B_combined

ggsave("Fig1AB_combined_colour_coded_20260129.jpg", dpi = 300, height=16, width=20, device=jpeg)
ggsave("Fig1AB_combined_colour_coded_20260129.tiff", dpi = 300, height=16, width=20, device=tiff)
ggsave("Fig1AB_combined_colour_coded_20260129.pdf", dpi = 300, height=16, width=20, device=cairo_pdf)
```

### Letters to add to Figure 1B

Add these letters manually in an image editor prior to upload to the supplementary information.

- Wild type = a
- ∆phz = a
- ∆hcn = ad
- ∆phz∆hcn = b
- Spent medium = c
- Spent medium from ∆phz = cd
- Sterile medium = e

# Figure S1: Contribution of background Co(2+) in medium from "no wire" controls

1. Reuse the combined data frame but filter it only to the no wire treatments. Facet the data so each Abiotic_vs_biotic treatment ordered is in its own panel, preserve the colour coding but eliminate the legend, and fix the y-axis to the same range as the main figure.
```{r}
All_cobalt_data_combined_revised <- fread("All_cobalt_ICPMS_data_cleaned_formatted_revised_mass_balance_summaries_included_20251001.txt", sep = "\t", header = "auto")

#Adjust the presentation aesthetics
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
        size = 34, angle = 0, hjust = .5, vjust = 0, face = "bold"),
        axis.title.y = element_text(color = "grey20", 
        size = 34, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 32, colour = "grey20", face = "bold"),
        strip.text.y = element_text(size = 32, colour = "grey20"),
        plot.title = element_text(size = 30, face = "bold", hjust = .5),
        legend.title=element_text(size=32, face = "bold"), 
        legend.text=element_text(size=32),
        panel.border = element_rect(colour = "black", fill=NA, size=1))


#Create custom display orders
All_cobalt_data_combined_revised$Abiotic_vs_biotic_ordered <- factor(All_cobalt_data_combined_revised$Abiotic_vs_biotic,
                                                             levels = c("Wild type",
                                                                        "∆phz",
                                                                        "∆hcn",
                                                                        "∆phz∆hcn",
                                                                        "Spent medium",
                                                                        "Spent medium from ∆phz",
                                                                        "Sterile medium"))


All_cobalt_data_combined_revised$Metabolite_treatment_ordered <- factor(All_cobalt_data_combined_revised$Metabolite_treatment,
                                                                 levels = c("Nothing added",
                                                                            "PCA",
                                                                            "MeOH"))
No_wire_scatterplot <- All_cobalt_data_combined_revised %>%
                       filter(Wire_treatment == "No wire") %>% #filter to no wire experiments
                       filter(DMSO_treatment == "No DMSO") %>% #use this to only focus on experiments that did not receive DMSO
                       filter(Sample_preservation == "Acidified, filtered") %>%
                       filter(!Abiotic_vs_biotic == "Sterile medium, anoxic" & !Abiotic_vs_biotic == "Spent medium from ∆phz, anoxic") %>%
                       group_by(Abiotic_vs_biotic, Wire_treatment, DMSO_treatment, Metabolite_treatment, Time) %>%
                       filter(Metabolite_treatment_ordered == "Nothing added") %>% #remove all of the other metabolite treatments
                       mutate(Abiotic_vs_biotic_ordered = fct_recode(Abiotic_vs_biotic_ordered, "Spent medium\nfrom ∆phz" = "Spent medium from ∆phz")) %>%
                      #plotting script
                      ggplot(aes(x=Time, y = Mean_cobalt_concentration_uM_corrected)) +
                      geom_errorbar(aes(ymin = Mean_cobalt_concentration_uM_corrected - Stdev_cobalt_concentration_uM_corrected,
                                        ymax = Mean_cobalt_concentration_uM_corrected + Stdev_cobalt_concentration_uM_corrected)) +
                      geom_point(shape = 21, color = "black", stroke = 0.8, size = 6) +
                      geom_line(aes(group = Bottle_number))+
                      geom_hline(yintercept = 1.68, linetype = "dashed", color = "orange", size = 1) +
                      scale_x_continuous(breaks = seq(0, 168, by = 24)) +
                      #ylim(0, 6500) +
                      facet_grid(~ Abiotic_vs_biotic_ordered, drop = FALSE) +
                      ylab("Co(2+) concentration (µM)") +
                      xlab("Time elapsed (hours)") +
                      theme +
                      theme(plot.margin = unit(c(1,1,1,1), "cm"),
                            axis.title.x = element_text(vjust = -2),
                            axis.title.y = element_text(vjust = 2),
                            legend.position = "none")
          
#Add a layer to angle the text
No_wire_scatterplot_angled <- No_wire_scatterplot + theme(axis.text.x = element_text(angle = 45))

#Save in different image formats at 300 dpi
ggsave("FigS1_Cobalt_ICPMS_data_no_wire_treaments_scatterplot_angled_20260202.jpg", dpi = 300, height=8, width=34, device= jpeg)
ggsave("FigS1_Cobalt_ICPMS_data_no_wire_treaments_scatterplot_angled_20260202.tiff", dpi = 300, height=8, width=34, device= tiff)
ggsave("FigS1_Cobalt_ICPMS_data_no_wire_treaments_scatterplot_angled_20260202.pdf", dpi = 300, height=8, width=34, device= cairo_pdf)
```
**Note** Spent medium from 2024-06-28 and a few others were incorrectly labelled as sterile medium, went and manually corrected in input file. Noticed because 72 hours was blank in the original. Seems like a mislabel that was manual in the original data.



# Figure S2: Boxplot of [Co(2+)] at 168 hours for abiotic metabolite treatments

1. Load the data frame from the compiled data and remove the outlier value from the sterile medium treatment that was contaminated.
```{r}
All_cobalt_data_combined_revised <- fread("All_cobalt_ICPMS_data_cleaned_formatted_revised_mass_balance_summaries_included_20251001.txt", sep = "\t", header = "auto")

#Create custom display orders
All_cobalt_data_combined_revised$Abiotic_vs_biotic_ordered <- factor(All_cobalt_data_combined_revised$Abiotic_vs_biotic,
                                                             levels = c("Wild type",
                                                                        "∆phz",
                                                                        "∆hcn",
                                                                        "∆phz∆hcn",
                                                                        "Spent medium",
                                                                        "Spent medium from ∆phz",
                                                                        "Sterile medium"))

All_cobalt_data_combined_revised$Metabolite_treatment_ordered <- factor(All_cobalt_data_combined_revised$Metabolite_treatment,
                                                                 levels = c("Nothing added",
                                                                            "PCA",
                                                                            "MeOH"))
Abiotic_no_outlier <- All_cobalt_data_combined_revised %>%
                      filter(DMSO_treatment == "No DMSO") %>% #use this to only focus on experiments that did not receive DMSO
                      filter(Sample_preservation == "Acidified, filtered") %>% #216 observations
                      filter(Time == 168 & Wire_treatment == "Wire included") %>% #Only look at no metabolite spikes + 168 hours
                      filter(Abiotic_vs_biotic_ordered == "Sterile medium")
```

2. Generate the linear models and verify assumptions for the ANOVA using type III sum of squares.
```{r}
##Generate model
Abiotic_model_outlier_removed <- lm(Cobalt_concentration_uM_corrected ~ Metabolite_treatment, 
                                    data = Abiotic_no_outlier, 
                                    contrasts = list(Metabolite_treatment = contr.sum), type =3)

Abiotic_model_outlier_removed
par(mfrow = c(2,2))
plot(Abiotic_model_outlier_removed)

#Run Breusch Pagan test for homoscedasticity
bptest(Abiotic_model_outlier_removed)

#Run Shapiro-Wilk normality test
shapiro.test(Abiotic_model_outlier_removed$residuals)

#Run Durbin-Watson test for autocorrelation
dwtest(Abiotic_model_outlier_removed)
```
## Figure S3 ANOVA Interpretation
- Homoscedasticity may be an issue, it looks like samples associated with higher cobalt concentrations ~ 240 µM have higher variance.
- Normal distribution of residuals looks good.
- BP test returns a p-value of *0.2725* --> we cannot reject the null hypothesis of homoscedasticity.
- SW test returns a p-value of *0.4557* --> we cannot reject the null hypothesis of normally distributed residuals.
- DW test returns a p-value of *0.9226* --> we cannot reject the null hypothesis that there is no autocorrelation between data points.

**All assumptions met, okay to proceed with ANOVA**

3. Run the ANOVA and the Tukey comparison.
```{r}
Abiotic_type_III_ANOVA_results <- Anova(Abiotic_model_outlier_removed, type = 3)
Abiotic_type_III_ANOVA_results
```
- Metabolite treatment did not emerge as a significant variable, so there is no significant differences between groups.
- F-value 0.8128
- p-value 0.4666

4. Generate the boxplot. There is no need to annotate the boxes later because there are no significant differences.
```{r}
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 16, angle = 45, hjust = 0.5, vjust = 0.55, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 16, angle = 0, hjust = 0.5, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 18, angle = 0, hjust = 0.5, vjust = -1, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 18, angle = 90, face = "plain"),
        strip.text.x = element_text(size = 20, colour = "grey20"),
        strip.text.y = element_text(size = 20, colour = "grey20"),
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5, vjust = 1),
        legend.title=element_text(size=18), 
        legend.text=element_text(size=18),
        legend.position = "right",
        panel.border = element_rect(colour = "black", fill=NA, size=1))

Abiotic_outlier_removed_boxplot <- Abiotic_no_outlier %>%
                                   ggplot(aes(x=Metabolite_treatment_ordered, y=Cobalt_concentration_uM_corrected)) + 
                                   geom_boxplot() +
                                   geom_jitter(fill = "grey", colour = "black", width = 0.1, pch =21, stroke =1)+
                                   ylim(0, 400) +
                                   labs(title = "" ,
                                   y = "Co(2+) concentration (µM)\ncorrected for recovery", 
                                   x = "Metabolic treatments in sterile medium") +
                                   theme +
                                   theme(plot.margin = unit(c(1,1,1,1), "cm"),
                                         axis.title.x = element_text(vjust = -2),
                                         axis.title.y = element_text(vjust = 2))
Abiotic_outlier_removed_boxplot
ggsave("FigS2_Cobalt_ICPMS_data_abiotic_treatments_no_DMSO_boxplot_20260202.jpg", dpi = 300, height=6, width=6, device=jpeg)
ggsave("FigS2_Cobalt_ICPMS_data_abiotic_treatments_no_DMSO_boxplot_20260202.tiff", dpi = 300, height=6, width=6, device=tiff)
ggsave("FigS2_Cobalt_ICPMS_data_abiotic_treatments_no_DMSO_boxplot_20260202.pdf", dpi = 300, height=6, width=6, device=cairo_pdf)
```
#  Biosynthetic gene cluster analyses

These analyses underpin *Figure S3,* *Table S2*, and *Table S3*, which were extracted from the notes in this R notebook.

## Run DRAM for annotations and distillation

1. Activate the conda environment containing DRAM and run the annotate command. Replace the directory and fasta file with the genome file for Pseudomonas chlororaphis subspecies aurefoaciens (or any genome or genoems of your choice)
```{bash}
DRAM.py annotate -i 'Pseudomonas_genome_analyses/*.fasta' --use_uniref -o Pseudomonas_annotation_with_uniref --threads 8
```

2. Once the annotation is complete, use the distillation command. Make sure to use the same output directory from the annotate step.
```{bash}
DRAM.py distill -i Pseudomonas_annotation_with_uniref/annotations.tsv -o Pseudomonas_genome_summaries --trna_path Pseudomonas_annotation_with_uniref/trnas.tsv --rrna_path Pseudomonas_annotation_with_uniref/rrnas.tsv
```

## Table S2 raw data: Manually search annotations for genes of interest corresponding to antiSMASH predictions and potential corrosion pathways

These files were used to manually create **Table S2**.

1. Load in the tab-delimited file containing the annotation data. Make sure to convert the MS Excel file to a flat format, i.e., .tsv manually.
```{r}
Pseudomonas_chlororaphis_aureofaciens_annotations <- fread("Supporting_Data_1B.txt", sep = "\t", header = "auto") %>% slice (-1)
colnames(Pseudomonas_chlororaphis_aureofaciens_annotations) <- as.character(Pseudomonas_chlororaphis_aureofaciens_annotations[1, ])
Pseudomonas_chlororaphis_aureofaciens_annotations <- Pseudomonas_chlororaphis_aureofaciens_annotations[-1,]
```

2. Manually search the annotation file using different strings of test for functions relevant to cobalt and microbially-induced corrosion.
```{r}
corrosion_pathways <- Pseudomonas_chlororaphis_aureofaciens_annotations %>%
                    
                   #Simple text search for "cobalt"
                   filter(
                     (if_any(c(kegg_hit:pfam_hits), ~ grepl("cobalt", .))) |
                     (if_any(c(kegg_hit:pfam_hits), ~ grepl("cob", .))) |
                     (if_any(c(kegg_hit:pfam_hits), ~ grepl("cyanide", .)))|
                     (if_any(c(kegg_hit:pfam_hits), ~ grepl("hcn", .)))|
                     (if_any(c(kegg_hit:pfam_hits), ~ grepl("cytochrome", .)))|
                     #(if_any(c(kegg_hit:pfam_hits), ~ grepl("hydrogenase", .)))| this is non-specific and picks up anything with dehydrogenase, comment out
                     (if_any(c(kegg_hit:pfam_hits), ~ grepl("siderophore", .)))) %>%
                    dplyr::select(scaffold,
                           gene_position,
                           start_position,
                           end_position,
                           ko_id,
                           kegg_hit,
                           uniref_id,
                           uniref_hit)

#Write the table to save the source data
write.table(corrosion_pathways, file = "SuppData_Pseudomonas_chlororaphis_aureofaciens_putative_corrosion_pathways.tsv", sep = "\t", row.names = FALSE)
``` 

*key biosynthetic genes* are denoted by asterisks in markdown format.

## Table S3 raw data: Cross-referencing annotations with potential metabolites

### Region 2 (hydrogen cyanide cluster): 2951483 to 2964455
```{r}
Region2 <- Pseudomonas_chlororaphis_aureofaciens_annotations %>% filter(start_position >= 2951483 & end_position <=2964455)
```

Genes in hydrogen cyanide-cluster:
- *ctg1_2591*: 2956483 - 2956800 --> hydrogen cyanide synthase HcnA [EC:1.4.99.5] (C grade, KEGG ID = K10814)
- *ctg1_2592*: 2956797 - 2958206 --> hydrogen cyanide synthase HcnB [EC:1.4.99.5] (B grade, KEGG ID = K10815)
- *ctg1_2593*: 2958199 - 2959455 --> hydrogen cyanide synthase HcnC [EC:1.4.99.5] (B grade, KEGG ID = K10816)
- ctg1_2594: 2959618 - 2960223 --> glutathione S-transferase [EC:2.5.1.18] (B grade, KEGG ID = K00799)

### Region 6 (NI-siderophore): 3832837 to 3869821
```{r}
Region6 <- Pseudomonas_chlororaphis_aureofaciens_annotations %>% filter(start_position >= 3832837 & end_position <=3869821)
```

Genes in NI-siderophore cluster:

- *ctg1_3372*: 3846837 - 3848735 --> *staphyloferrin* B synthase [EC:6.3.2.56] (B grade KEGG ID = K23375)
- ctg1_3373: 3848740 - 3849513 --> *staphyloferrin* B biosynthesis citrate synthase (B grade, KEGG ID = K23376)
- *ctg1_3374*: 3849507 - 3851369 --> 2-[(L-alanin-3-ylcarbamoyl)methyl]-3-(2-aminoethylcarbamoyl)-2-hydroxypropanoate synthase [EC:6.3.2.55] (B grade, KEGG ID = K23374)
- ctg1_3375: 3851394 - 3852812 --> UniRef90_A0A554P7Y5 DHA2 family efflux MFS transporter permease subunit n=30 Tax=Pseudomonadaceae TaxID=135621 RepID=A0A554P7Y5_9PSED (B grade, no hit in KEGG)
- ctg1_3376: 3852809 - 3854029 --> diaminopimelate decarboxylase [EC:4.1.1.20] (B grade, KEGG ID = K01586)
- *ctg1_3377*: 3854022 - 3855821 --> UniRef90_A0A0C5EL69 AcsD protein n=70 Tax=Pseudomonadaceae TaxID=135621 RepID=A0A0C5EL69_9PSED (B grade, no hit in KEGG)

### Region 7 (hydrogen cyanide cluster): 3888854 to 3901701
```{r}
Region7 <- Pseudomonas_chlororaphis_aureofaciens_annotations %>% filter(start_position >= 3888854 & end_position <=3901701)
```

- ctg1_3405: 3890984 - 3892441 --> succinate-semialdehyde dehydrogenase / glutarate-semialdehyde dehydrogenase [EC:1.2.1.16 1.2.1.79 1.2.1.20] (B grade, KEGG ID = K00135)
- ctg1_3406: 3892700 - 3893857 --> acetylornithine deacetylase [EC:3.5.1.16] (B grade, KEGG ID = K01438)
- *ctg1_3407*: 3893854 - 3895008 --> UniRef90_A0AA45XIN0 Sarcosine oxidase subunit beta n=5 Tax=Pseudomonas TaxID=286 RepID=A0AA45XIN0_9PSED (no KEGG hit, B grade, UniRef ID = A0AA45XIN0_9PSED)
  - Catalyzes oxidative demethylation of sarcosine (N-methylglycine) to yield glycine
- *ctg1_3408*: 3895005 - 3896411 --> UniRef90_A0A3G7K0E2 Opine oxidase subunit A n=39 Tax=Pseudomonas TaxID=286 RepID=A0A3G7K0E2_9PSED, (no KEGG hit, B grade, Uniref ID = A0A3G7K0E2_9PSED)
  - Cleaves opines into amino acids and pyruvate
- *ctg1_3409*: 3896393 - 3896701 --> UniRef90_A0A5M7CTD1 (2Fe-2S)-binding protein n=14 Tax=Pseudomonadaceae TaxID=135621 RepID=A0A5M7CTD1_9PSED (C grade, no KEGG hit, UniRef ID = A0A5M7CTD1_9PSED)
- ctg1_3410: 3896698 - 3898539 --> peptide/nickel transport system ATP-binding protein (B grade, KEGG ID = K02031)
- ctg1_3411: 3898541 - 3899416 --> UniRef90_A0A0C1ZSU6 DNA-directed RNA polymerase subunit alpha n=30 Tax=Pseudomonadaceae TaxID=135621 RepID=A0A0C1ZSU6_PSEFL (no KEGG hit, B grade, Uniref ID = A0A0C1ZSU6_PSEFL)
- ctg1_3412: 3899413 - 3900369 --> peptide/nickel transport system permease protein (B grade, KEGG ID = K02033)

### Region 10 (NRP-metallophore): 5056434 to 5155407
```{r}
Region10 <- Pseudomonas_chlororaphis_aureofaciens_annotations %>% filter(start_position >= 5056434 & end_position <=5155407)
```

- *ctg1_4474*: 5,076,434 - 5,077,768 --> L-ornithine N5-monooxygenase [EC:1.14.13.195 1.14.13.196] (B grade, KEGG ID = K10531)
- ctg1_4475: 5,077,900 - 5,078,496 --> UniRef90_A0A165YGX4 RNA polymerase subunit sigma n=4 Tax=Pseudomonas TaxID=286 RepID=A0A165YGX4_PSEFL (C grade, no KEGG ID)
- ctg1_4476: 5,078,716 - 5,079,888 --> membrane fusion protein, macrolide-specific efflux system (B grade, KEGG ID = K13888)
- ctg1_4477: 5,079,889 - 5,081,862 --> macrolide transport system ATP-binding/permease protein [EC:7.6.2.-] (B grade, KEGG ID = K05685)
- ctg1_4478: 5,081,870 - 5,083,279 --> outer membrane protein, multidrug efflux system (B grade, KEGG ID = K18139)
- *ctg1_4479*: 5,083,355 - 5,084,977 --> UniRef90_A0A554PAC0 Twin-arginine translocation signal domain-containing protein n=19 Tax=Pseudomonadaceae TaxID=135621 RepID=A0A554PAC0_9PSED (B grade, no KEGG hit)
- ctg1_4480: 5,085,186 - 5,086,595 --> membrane dipeptidase [EC:3.4.13.19] (B grade, KEGG ID = K01273)
- ctg1_4481: 5,086,617 - 5,087,915 --> UniRef90_A0A2C9EQN1 Isopenicillin N epimerase CefD n=56 Tax=Pseudomonadaceae TaxID=135621 RepID=A0A2C9EQN1_PSEPH (B grade, no KEGG hit)
- *ctg1_4482*: 5,087,959 - 5,088,855 --> UniRef90_Q4K997 Chromophore maturation protein *PvdO* n=55 Tax=Pseudomonadaceae TaxID=135621 RepID=Q4K997_PSEF5 (B grade, no KEGG hit)
- ctg1_4483: 5,088,910 - 5,089,773 --> UniRef90_Q4K996 phosphoribosylglycinamide formyltransferase 1 n=49 Tax=Pseudomonadaceae TaxID=135621 RepID=Q4K996_PSEF5 (B grade, no KEGG hit)
- ctg1_4484: 5,089,961 - 5,091,613 --> putative pyoverdin transport system ATP-binding/permease protein (B grade, KEGG ID = K06160)
- ctg1_4485: 5,091,714 - 5,094,182 --> outer-membrane receptor for ferric coprogen and ferric-rhodotorulic acid (B grade, KEGG ID = K16088)
- *ctg1_4486*: 5,094,570 - 5,107,805 --> UniRef90_A0A554PAB6 Amino acid adenylation domain-containing protein n=22 Tax=Pseudomonas TaxID=286 RepID=A0A554PAB6_9PSED (B grade, no KEGG hit)
- *ctg1_4487*: 5,107,818 - 5,114,414 --> UniRef90_A0A554PAC2 Non-ribosomal peptide synthetase n=26 Tax=Pseudomonas TaxID=286 RepID=A0A554PAC2_9PSED (B grade, no KEGG hit)
- *ctg1_4488*: 5,114,414 - 5,125,576 --> UniRef90_UPI002407C9DD amino acid adenylation domain-containing protein n=4 Tax=Pseudomonas chlororaphis TaxID=587753 RepID=UPI002407C9DD (B grade, no KEGG hit)

### Region 11 (NRPS): 5189456 to 5242472
```{r}
Region11 <- Pseudomonas_chlororaphis_aureofaciens_annotations %>% filter(start_position >= 5189456 & end_position <=5242472)
```

NRPS region here is really only one gene:
- *ctg1_4563*: 5209456 - 5222472 --> UniRef90_A0A554PEL7 Non-ribosomal peptide synthetase n=29 Tax=Pseudomonadaceae TaxID=135621 RepID=A0A554PEL7_9PSED (B grade, no KEGG hit)

# Figure 2: Wild type vs deletion cell growth in corrosion assays as measured by viable plate counts

1. Load the plate count data.
```{r}
Plate_counts_combined <- fread("Plate_counts_combined_20250915.txt", sep = "\t", header = "auto")
```

2. Add new variables to the data frame parsing the Comment variable that contains detailed informtion on the sample
```{r}
Plate_counts_combined_treatments_added <- Plate_counts_combined %>%
                                          #Add time as a numeric variable
                                          mutate(Time = case_when(grepl("0hr", Comment) ~ "0",
                                                                  grepl("24hr", Comment) ~ "24",
                                                                  grepl("72hr", Comment) ~ "72",
                                                                  grepl("168hr", Comment) ~ "168")) %>%
  
                                          mutate(Time = as.numeric(Time)) %>%
        
                                          #Add wire treatment as a dedicated variable and leave as character
                                          mutate(Wire_treatment = case_when(grepl("_Wire", Comment) ~ "Wire included",
                                                                            grepl("NoWire", Comment) ~ "No wire")) %>%
  
  
                                          #Add strain information
                                          mutate(Strain = case_when(grepl("+10%inoc_", Comment) ~ "Wild type",
                                                                    grepl("DMSO", Comment) ~ "Wild type",
                                                                    grepl("0%inoc∆phz", Comment) ~ "∆phz",
                                                                    grepl("0%inoc∆hcn", Comment) ~ "∆hcn",
                                                                    grepl("DD", Comment) ~ "∆phz∆hcn")) %>%
  
                                          #Add DMSO information
                                          mutate(DMSO_treatment = case_when(!grepl("DMSO", Comment) ~ "No DMSO",
                                                                            grepl("+DMSO", Comment) ~ "DMSO")) 

Plate_counts_combined_summary_statistics <- Plate_counts_combined_treatments_added %>%
                                            group_by(Strain, Wire_treatment, DMSO_treatment, Flask_number, Time) %>%
                                            filter(CFU_mL >0) %>% #remove any values that had 0 counts on the plate for graphing
            
                                            #Calculate plate counts using technical replicates for each flask at each time point
                                            mutate(Technical_mean_CFU_mL = mean(CFU_mL),
                                                   Technical_stdev_CFU_mL = sd(CFU_mL)) %>%
                              
                                             ungroup() %>%
                            
                                             #Calculate plate counts using average of technical replicates for all flasks (biological replicates)
                                             group_by(Strain, Wire_treatment, DMSO_treatment, Time) %>%
                                             
                                             mutate(Biological_mean_CFU_mL = mean(Technical_mean_CFU_mL),
                                                    Biological_stdev_CFU_mL = sd(Technical_mean_CFU_mL))

write.table(Plate_counts_combined_summary_statistics, file = "Plate_counts_combined_summary_statistics_20251009.txt", sep = "\t", 
            row.names = FALSE, col.names = TRUE) 
```

3. Plot the growth data, summarizing the means and standard deviations prior to the ggplot command.
```{r}
#Adjust the presentation aesthetics
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
        size = 34, angle = 0, hjust = .5, vjust = 0, face = "bold"),
        axis.title.y = element_text(color = "grey20", 
        size = 34, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 32, colour = "grey20", face = "bold"),
        strip.text.y = element_text(size = 32, colour = "grey20"),
        plot.title = element_text(size = 30, face = "bold", hjust = .5),
        legend.title=element_text(size=32, face = "bold"), 
        legend.text=element_text(size=32),
        panel.border = element_rect(colour = "black", fill=NA, size=1))

dark2_colors <- brewer.pal(n = 7, name = "Dark2") #generate colour palette with desired levels to match ICP-MS data

#Create custom display orders
Plate_counts_combined_treatments_added$Strain_ordered <- factor(Plate_counts_combined_treatments_added$Strain,
                                                             levels = c("Wild type",
                                                                        "∆phz",
                                                                        "∆hcn",
                                                                        "∆phz∆hcn"))


Growth_curves_log_scale <- Plate_counts_combined_treatments_added %>%
                           filter(DMSO_treatment == "No DMSO") %>% #use this to only focus on experiments that did not receive DMSO
                           group_by(Strain, Wire_treatment, DMSO_treatment, Flask_number, Time) %>%
  
                           filter(CFU_mL >0) %>% #remove any values that had 0 counts on the plate for graphing
            
                           #Calculate plate counts using technical replicates for each flask at each time point
                           mutate(Technical_mean_CFU_mL = mean(CFU_mL),
                                  Technical_stdev_CFU_mL = sd(CFU_mL)) %>%
            
                           ungroup() %>%
          
                           #Calculate plate counts using average of technical replicates for all flasks (biological replicates)
                           group_by(Strain, Wire_treatment, DMSO_treatment, Time) %>%
                           
                           mutate(Biological_mean_CFU_mL = mean(Technical_mean_CFU_mL),
                                  Biological_stdev_CFU_mL = sd(Technical_mean_CFU_mL)) %>%
  
                           
                           #plotting script
                      
                           ggplot(aes(x=Time, y = log10(Biological_mean_CFU_mL), fill = Strain_ordered)) +
                           geom_errorbar(aes(ymin = log10(Biological_mean_CFU_mL - Biological_stdev_CFU_mL),
                                             ymax = log10(Biological_mean_CFU_mL + Biological_stdev_CFU_mL),
                                             colour = Strain_ordered)) +
                           geom_point(shape = 21, color = "black", stroke = 0.8, size = 6) +
                           
                           geom_line(aes(group = Flask_number, color = Strain_ordered))+
                           scale_x_continuous(breaks = seq(0, 168, by = 24))+
                           facet_grid(~ Wire_treatment) +
                           ylab("Log10\n(CFU/mL)") +
                           scale_y_continuous(labels = label_math(), limits = c(6, 10.5))+
                           xlab("Time elapsed (hours)") + 
                           scale_color_manual(name = "Strain", 
                                               values = dark2_colors,
                                               labels = c("Wild type", 
                                                        expression(italic("∆phz")),
                                                        expression(italic("∆hcn")),
                                                        expression(italic("∆phz∆hcn"))))+
                           scale_fill_manual(name = "Strain", 
                                             values = dark2_colors,
                                             labels = c("Wild type", 
                                                      expression(italic("∆phz")),
                                                      expression(italic("∆hcn")),
                                                      expression(italic("∆phz∆hcn")))) +
                           theme +
                           theme(plot.margin = unit(c(1,1,1,1), "cm"),
                                 axis.title.x = element_text(vjust = -2),
                                 axis.title.y = element_text(vjust = 2))
          
Growth_curves_log_scale

Growth_curves_log_scale_no_legend_angled <- Growth_curves_log_scale + theme(axis.text.x = element_text(angle = 45))

ggsave("Fig2_plate_count_data_WT_phz_mutant_scatterplot_log10_20260202.jpg", dpi = 300, height=8, width=20, device=jpeg)
ggsave("Fig2_plate_count_data_WT_phz_mutant_scatterplot_log10_20260202.tiff", dpi = 300, height=8, width=20, device=tiff)
ggsave("Fig2_plate_count_data_WT_phz_mutant_scatterplot_log10_20260202.pdf", dpi = 300, height=8, width=20, device=cairo_pdf)
```

# Figure S4: Comparison of [Co(2+)] between acidify-filter and filter-acidify preservation with wild-type live cells

1. Use the code below to load in source data.
```{r}
All_cobalt_data_combined_revised <- fread("All_cobalt_ICPMS_data_cleaned_formatted_revised_mass_balance_summaries_included_20251001.txt", sep = "\t", header = "auto")
```

2. Use this code to create a similar figure to Figure 1 but facet it according to sample preservation and plot each replicate individually.
```{r}
#Adjust the presentation aesthetics
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
        size = 34, angle = 0, hjust = .5, vjust = 0, face = "bold"),
        axis.title.y = element_text(color = "grey20", 
        size = 34, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 32, colour = "grey20", face = "bold"),
        strip.text.y = element_text(size = 32, colour = "grey20"),
        plot.title = element_text(size = 30, face = "bold", hjust = .5),
        legend.title=element_text(size=32, face = "bold"), 
        legend.text=element_text(size=32),
        panel.border = element_rect(colour = "black", fill=NA, size=1))


#Graphing script with additional data manipulation
Acidify_filter_scatterplot <-  All_cobalt_data_combined_revised %>%
                               filter(Wire_treatment == "Wire included") %>% #use this to only focus on experiments that did not receive DMSO
                               filter(Abiotic_vs_biotic == "Wild type") %>%
                               filter(Metabolite_treatment == "Nothing added") %>%
                               filter(DMSO_treatment == "No DMSO") %>% #96 observations
                               filter(between(Replicate, 1, 4)) %>%
                              #plotting script
  
                               ggplot(aes(x=Time, 
                                          y = Cobalt_concentration_uM_corrected
                                          )) +
                                      geom_point(size = 6) +
                                      geom_line(aes(group = Bottle_number))+
                                      scale_x_continuous(breaks = seq(0, 168, by = 24)) +
                                      facet_grid(Replicate ~ Sample_preservation) +
                                      ylab("Co(2+) concentration (µM) corrected for recoveries") +
                                      xlab("Time elapsed (hours)") + 
                                      theme +
                                      theme(plot.margin = unit(c(1,1,1,1), "cm"),
                                            axis.title.x = element_text(vjust = -2),
                                            axis.title.y = element_text(vjust = 2),
                                            axis.text.x = element_text(angle = 45))

Acidify_filter_scatterplot

#Save the images in different formats
ggsave("FigS4_cobalt_ICPMS_data_sample_preservation_scatterplot_20260202.jpg", dpi = 300, height=16, width=12, device=jpeg)
ggsave("FigS4_cobalt_ICPMS_data_sample_preservation_scatterplot_20260202.tiff", dpi = 300, height=16, width=12, device=tiff)
ggsave("FigS4_cobalt_ICPMS_data_sample_preservation_scatterplot_20260202.pdf", dpi = 300, height=16, width=12, device=cairo_pdf)
```
# Figure S5. log(µmol Co(2+)/CFU) over Co(2+) exposure from spike concentrations in corrosion assays compared to dose-response experiments

**Note:** These analyses were conducted manually in MS Excel and are tied to the source files:
- "FigS5A_source_data.xlsx" on the sheet "Paired analyses"
- "FigS5B_source_data.xlsx" on the sheet "Paired analyses condensed"

# Figure S6. Cobalt dose-response curve reporting growth rate of P. chlororaphis subsp. aureofaciens in medium amended with CoCl2

3. Try loading the data where growth rates were manually calculated and refitting the dose response curve to see if it changes the conclusion.
```{r}
#Load the data
library(tidyverse)
library(growthcurver)
library(drc)
library(readxl)

Growth_data <- read_excel("FigS6_source_data.xlsx", sheet = "Sheet1")

Growth_rate_4_param_model_manual<- drm(Growth_rate ~ Total_cobalt_recovered_uM, 
                                data = Growth_data, 
                                fct = LL.4())

summary(Growth_rate_4_param_model_manual)

Growth_rate_dose_response_curve_summary_manual <- ED(Growth_rate_4_param_model_manual, c(5, 10, 50), interval = "delta")
Growth_rate_dose_response_curve_summary_manual

pdf("FigS6_Pseudomonas_chlororaphis_aureofaciens_cobalt_dose_response_curve_manual_calculation_20260202.pdf")   # open PDF device
Growth_rate_dose_response_plot_manual<- plot(Growth_rate_4_param_model_manual, broken = FALSE, type = "all",
                                      xlab = "Co(2+) concentration (µM)", xlim = c(0, 2000),
                                      ylab = "Growth rate (hr-1)", 
                                      main = "Pseudomonas chlororaphis aureofaciens\ngrowth curve in response to Co(2+)")
dev.off()  
```

#Figure S7: Boxplots of [Co(2+)] between A-F and F-A preservation in a Co(2+) dose-response curve with wild type cells

1. Import the source data.
```{r}
Cobalt_data_treatments_added <- fread("FigS7_source_data.txt", sep = "\t", header = "auto")
```

2. Create a boxplot with the full data set and another version with the extreme value from the 2000 µM treatment removed.
```{r}
Cobalt_boxplot_free <- ggplot(Cobalt_data_treatments_added,
                   aes(x = Sample_preservation,
                       y = Cobalt_concentration_uM,
                       fill = Sample_preservation)) +
  geom_boxplot(width = 0.6, outlier.shape = NA, alpha = 0.8) +
  
  geom_jitter(
    shape = 21,           # filled circle
    colour = "black",     # optional outline
    width = 0.15,
    height = 0,
    alpha = 0.6,
    size = 1.8
  ) +
  
  facet_wrap(
    ~ Spike_concentration,
    scales = "free_y",
    nrow = 1,
    labeller = labeller(
      Spike_concentration = function(x) paste0(x, " µM")
    )
  ) +
  labs(
    x = "Preservation Method",
    y = "Cobalt concentration (µM)",
    fill = "Preservation Method"
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, colour = "black", size = 18),
    axis.text.y = element_text(colour = "black", size = 18),
    axis.title.x = element_text(margin = margin(t = 16), size = 20),
    axis.title.y = element_text(margin = margin(r = 16), size = 20),
    panel.grid = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5),
    strip.background = element_rect(fill = "grey85", colour = "grey50"),
    strip.text = element_text(face = "bold", size = 22),
    legend.position = "none"
  )

Cobalt_boxplot_free

Cobalt_boxplot_free_outlier_removed <- Cobalt_data_treatments_added %>%
                                       dplyr::filter(!Sample_name == "2025-12-16_32hr_10%inoc+2000µMCo2+_3_F-A") %>% #remove outlier
  ggplot(aes(x = Sample_preservation,
                       y = Cobalt_concentration_uM,
                       fill = Sample_preservation)) +
  geom_boxplot(width = 0.6, outlier.shape = NA, alpha = 0.8) +
  
  geom_jitter(
    shape = 21,           # filled circle
    colour = "black",     # optional outline
    width = 0.15,
    height = 0,
    alpha = 0.6,
    size = 1.8
  ) +
  
  facet_wrap(
    ~ Spike_concentration,
    scales = "free_y",
    nrow = 1,
    labeller = labeller(
      Spike_concentration = function(x) paste0(x, " µM")
    )
  ) +
  labs(
    x = "Preservation Method",
    y = "Cobalt concentration (µM)",
    fill = "Preservation Method"
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, colour = "black", size = 18),
    axis.text.y = element_text(colour = "black", size = 18),
    axis.title.x = element_text(margin = margin(t = 16), size = 20),
    axis.title.y = element_text(margin = margin(r = 16), size = 20),
    panel.grid = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5),
    strip.background = element_rect(fill = "grey85", colour = "grey50"),
    strip.text = element_text(face = "bold", size = 22),
    legend.position = "none"
  )

Cobalt_boxplot_free_outlier_removed
ggsave("FigS7_Cobalt_recovered_sample_preservation_method_comparison_outlier_removed_20260202.pdf", dpi = 300, height=6, width=22, device= pdf)
```

# Figure S8: Comparison of [Co(2+)] between acidify-filter and filter-acidify preservation with sterile medium
```{r}
All_cobalt_data_combined_revised <- fread("All_cobalt_ICPMS_data_cleaned_formatted_revised_mass_balance_summaries_included_20251001.txt", sep = "\t", header = "auto")
```

2. Use this code to create a similar figure to Figure 1 but facet it according to sample preservation and plot each replicate individually.
```{r}
#Adjust the presentation aesthetics
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
        size = 34, angle = 0, hjust = .5, vjust = 0, face = "bold"),
        axis.title.y = element_text(color = "grey20", 
        size = 34, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 32, colour = "grey20", face = "bold"),
        strip.text.y = element_text(size = 32, colour = "grey20"),
        plot.title = element_text(size = 30, face = "bold", hjust = .5),
        legend.title=element_text(size=32, face = "bold"), 
        legend.text=element_text(size=32),
        panel.border = element_rect(colour = "black", fill=NA, size=1))


#Graphing script with additional data manipulation
Acidify_filter_scatterplot_sterile <-  All_cobalt_data_combined_revised %>%
                               filter(Wire_treatment == "Wire included") %>% #use this to only focus on experiments that did not receive DMSO
                               filter(Abiotic_vs_biotic == "Sterile medium") %>%
                               filter(Metabolite_treatment == "Nothing added") %>%
                               filter(DMSO_treatment == "No DMSO") %>% #96 observations
                               filter(between(Replicate, 2, 4)) %>%
                              #plotting script
  
                               ggplot(aes(x=Time, 
                                          y = Cobalt_concentration_uM_corrected
                                          )) +
                                      geom_point(size = 6) +
                                      geom_line(aes(group = Bottle_number))+
                                      scale_x_continuous(breaks = seq(0, 168, by = 24)) +
                                      facet_grid(Replicate ~ Sample_preservation) +
                                      ylab("Co(2+) concentration (µM) corrected for recoveries") +
                                      xlab("Time elapsed (hours)") + 
                                      theme +
                                      theme(plot.margin = unit(c(1,1,1,1), "cm"),
                                            axis.title.x = element_text(vjust = -2),
                                            axis.title.y = element_text(vjust = 2),
                                            axis.text.x = element_text(angle = 45))

Acidify_filter_scatterplot_sterile

#Save the images in different formats
ggsave("FigS8_cobalt_ICPMS_data_sample_preservation_sterile_only_scatterplot_20260202.jpg", dpi = 300, height=16, width=12, device=jpeg)
ggsave("FigS8_cobalt_ICPMS_data_sample_preservation_sterile_only_scatterplot_20260202.tiff", dpi = 300, height=16, width=12, device=tiff)
ggsave("FigS8_cobalt_ICPMS_data_sample_preservation_sterile_only_scatterplot_20260202.pdf", dpi = 300, height=16, width=12, device=cairo_pdf)
```
# Figure 4B: Boxplot of [Co(2+)] at 168 hours comparing oxic and anoxic treatments 

1. Load the data frame from the compiled data and remove the outlier value from the sterile medium treatment that was contaminated and trim the data to only include the oxic vs anoxic experiments conducted with sterile and spent medium 
```{r}
All_cobalt_data_combined_revised <- fread("All_cobalt_ICPMS_data_cleaned_formatted_revised_mass_balance_summaries_included_20251001.txt", sep = "\t", header = "auto")

Oxic_vs_anoxic_data <- All_cobalt_data_combined_revised %>%
                       filter(DMSO_treatment == "No DMSO") %>% #use this to only focus on experiments that did not receive DMSO
                       filter(Sample_preservation == "Acidified, filtered") %>%
                       filter(Metabolite_treatment == "Nothing added") %>%
                       filter(Time == 168 & Wire_treatment == "Wire included") %>%
                       filter(Abiotic_vs_biotic == "Sterile medium" |
                              Abiotic_vs_biotic == "Sterile medium, anoxic" |
                              Abiotic_vs_biotic == "Spent medium from ∆phz"|
                              Abiotic_vs_biotic == "Spent medium from ∆phz, anoxic")
```

2. Generate the linear models and verify assumptions for the ANOVA using type III sum of squares.
```{r}
##Generate model
Oxic_anoxic_model <- lm(Cobalt_concentration_uM_corrected ~ Abiotic_vs_biotic, 
                        data = Oxic_vs_anoxic_data, 
                        contrasts = list(Abiotic_vs_biotic = contr.sum), type =3)

Oxic_anoxic_model
par(mfrow = c(2,2))
plot(Oxic_anoxic_model)

#Run Breusch Pagan test for homoscedasticity
bptest(Oxic_anoxic_model)

#Run Shapiro-Wilk normality test
shapiro.test(Oxic_anoxic_model$residuals)

#Run Durbin-Watson test for autocorrelation
dwtest(Oxic_anoxic_model)
```
## Figure 4B ANOVA Interpretation
- Homoscedasticity may be an issue, it looks like samples associated with higher cobalt concentrations ~ 240 µM have higher variance.
- Normal distribution of residuals looks good.
- BP test returns a p-value of *0.2706* --> we cannot reject the null hypothesis of homoscedasticity.
- SW test returns a p-value of *0.4928* --> we cannot reject the null hypothesis of normally distributed residuals.
- DW test returns a p-value of *0.8272* --> we cannot reject the null hypothesis that there is no autocorrelation between data points.

**All assumptions met, okay to proceed with ANOVA**

3. Run the ANOVA and the Tukey comparison.
```{r}
Oxic_anoxic_type_III_ANOVA_results <- Anova(Oxic_anoxic_model, type = 3)
Oxic_anoxic_type_III_ANOVA_results
```
- F-value = 136.12
- p-value = 1.361e-10 ***
- There is a *significant* effect of the different abiotic vs biotic treatments, which in this case, are sterile vs spent medium oxic vs anoxic

4. Run a Tukey post-hoc test to figure out where the significant differences are.
```{r}
#First you need to convert the model to an aov object
Oxic_anoxic_model_converted <- aov(Oxic_anoxic_model)
Tukey_comparison_oxic_vs_anoxic <- TukeyHSD(Oxic_anoxic_model_converted, conf.level = 0.95)
Tukey_comparison_oxic_vs_anoxic
Tukey_comparison_oxic_vs_anoxic_data_frame <- as.data.frame(Tukey_comparison_oxic_vs_anoxic$Abiotic_vs_biotic)
```

5. Generate the boxplot. There is no need to annotate the boxes later because there are no significant differences.
```{r}
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 16, angle = 0, hjust = 0.5, vjust = 0.55, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 16, angle = 0, hjust = 0.5, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 18, angle = 0, hjust = 0.5, vjust = -1, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 18, angle = 90, face = "plain"),
        strip.text.x = element_text(size = 20, colour = "grey20"),
        strip.text.y = element_text(size = 20, colour = "grey20"),
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5, vjust = 1),
        legend.title=element_text(size=18), 
        legend.text=element_text(size=18),
        legend.position = "right",
        panel.border = element_rect(colour = "black", fill=NA, size=1))


Oxic_vs_anoxic_boxplot <- Oxic_vs_anoxic_data %>%
                          mutate(Abiotic_vs_biotic = fct_recode(Abiotic_vs_biotic, "Spent medium\nfrom ∆phz, anoxic" = "Spent medium from ∆phz, anoxic")) %>%
                          mutate(Abiotic_vs_biotic = fct_recode(Abiotic_vs_biotic, "Sterile medium,\nanoxic" = "Sterile medium, anoxic")) %>%
                          mutate(Abiotic_vs_biotic = fct_recode(Abiotic_vs_biotic, "Spent medium\nfrom ∆phz" = "Spent medium from ∆phz")) %>%
                          ggplot(aes(x=Abiotic_vs_biotic, y=Cobalt_concentration_uM_corrected)) + 
                          geom_boxplot() +
                          geom_jitter(fill = "grey", colour = "black", width = 0.1, pch =21, stroke =1)+
                          #ylim(0, 400) +
                          labs(title = "" ,
                          y = "Co(2+) concentration (µM)\ncorrected for recovery", 
                          x = "Oxic vs anoxic treatments in sterile and spent media") +
                          theme +
                          theme(plot.margin = unit(c(1,1,1,1), "cm"),
                                axis.title.x = element_text(vjust = -2),
                                axis.title.y = element_text(vjust = 2))
Oxic_vs_anoxic_boxplot

ggsave("Fig4B_Cobalt_ICPMS_data_oxic_vs_anoxic_sterile_spent_medium_comparison_boxplot_20260202.jpg", dpi = 300, height=6, width=11, device=jpeg)
ggsave("Fig4B_Cobalt_ICPMS_data_oxic_vs_anoxic_sterile_spent_medium_comparison_boxplot_20260202.tiff", dpi = 300, height=6, width=11, device=tiff)
ggsave("Fig4B_Cobalt_ICPMS_data_oxic_vs_anoxic_sterile_spent_medium_comparison_boxplot_20260202.pdf", dpi = 300, height=6, width=11, device=cairo_pdf)
```

## Figure 4B Tukey comparison results

$Abiotic_vs_biotic
                                                            diff         lwr        upr     p adj
Spent medium from ∆phz, anoxic-Spent medium from ∆phz -990.74815 -1160.73133 -820.76497 0.0000000
Sterile medium-Spent medium from ∆phz                 -832.72603  -971.51672 -693.93535 0.0000000
Sterile medium, anoxic-Spent medium from ∆phz         -989.60484 -1159.58802 -819.62166 0.0000000
Sterile medium-Spent medium from ∆phz, anoxic          158.02211    19.23143  296.81280 0.0236234
Sterile medium, anoxic-Spent medium from ∆phz, anoxic    1.14331  -168.83987  171.12649 0.9999972
Sterile medium, anoxic-Sterile medium                 -156.87880  -295.66949  -18.08812 0.0247129

### Figure 4B Letters to add to plot
- Spent medium from ∆phz = a
- Spent medium from ∆phz, anoxic = b
- Sterile medium = c
- Sterile medium, anoxic = b

# Figure S20: Boxplot of [Co(2+)] at 168 hours across DMSO biotic treatments

1. Load the source data and trim it to the desired samples to carry out the comparison for the effect of DMSO.
```{r}
All_cobalt_data_combined_revised <- fread("All_cobalt_ICPMS_data_cleaned_formatted_revised_mass_balance_summaries_included_20251001.txt", sep = "\t", header = "auto")

#Subet DMSO data
DMSO_comparison_data <- All_cobalt_data_combined_revised %>%
                        filter(Wire_treatment == "Wire included") %>%
                        filter(Sample_preservation == "Acidified, filtered") %>%
                        filter(Time == 168) %>%
                        filter(Abiotic_vs_biotic == "Wild type")
```

2. Use the subset data to generate a linear model that for an ANOVA comparison. Use this code to verify the assumptions of linear model fit.
```{r}
DMSO_comparison_model <- lm(Cobalt_concentration_uM_corrected ~ DMSO_treatment, data = DMSO_comparison_data,
                            contrasts = list(DMSO_treatment = contr.sum), type =3) #need to do this for unabalance sample size
DMSO_comparison_model
par(mfrow = c(2,2))
plot(DMSO_comparison_model)
```

### Interpretation
- Looks like we may have issues with equal variance in the absence of DMSO.
- Normality looks pretty good excpet for the 5th and 7th data points.

3. Run formal tests to verify the assumptions of linear models.
```{r}
#Run Breusch Pagan test for homoscedasticity
bptest(DMSO_comparison_model)

#Run Shapir-Wilk normality test
shapiro.test(DMSO_comparison_model$residuals)

#Run Durbin-Watson test for autocorrelation
dwtest(DMSO_comparison_model)
```

### Interpretation

studentized Breusch-Pagan test

data:  DMSO_comparison_model
BP = 6.5497, df = 3, p-value = 0.04823 *p < 0.05* meaning we reject the null hypothesis of homoscedasticity, this is good (just barely)


Shapiro-Wilk normality test

data:  DMSO_comparison_model$residuals
W = 0.72204, p-value = 0.6735 *p >0.05*, meaning we cannot reject the null hypothesis that are normally distributed.


Durbin-Watson test

data:  DMSO_comparison_model
DW = 1.9783, p-value = 0.1346

3. Regenerate the model with log10 transformed data.
```{r}
DMSO_comparison_model_log10 <- lm(log10(Cobalt_concentration_uM_corrected) ~ DMSO_treatment, data = DMSO_comparison_data,
                            contrasts = list(DMSO_treatment = contr.sum), type =3) #need to do this for unabalance sample size
DMSO_comparison_model_log10
par(mfrow = c(2,2))
plot(DMSO_comparison_model_log10)

#Formal tests

#Run Breusch Pagan test for homoscedasticity
bptest(DMSO_comparison_model_log10)

#Run Shapir-Wilk normality test
shapiro.test(DMSO_comparison_model_log10$residuals)

#Run Durbin-Watson test for autocorrelation
dwtest(DMSO_comparison_model_log10)
```
studentized Breusch-Pagan test

data:  DMSO_comparison_model
BP = 6.5497, df = 3, p-value = 0.2456 *p > 0.05* meaning we reject the null hypothesis of homoscedasticity, this is good (just barely)


Shapiro-Wilk normality test

data:  DMSO_comparison_model$residuals
W = 0.72204, p-value = 0.5072 *p >0.05*, meaning we cannot reject the null hypothesis that are normally distributed.


Durbin-Watson test

data:  DMSO_comparison_model
DW = 1.9783, p-value = 0.06003 *p > 0.05*

**All linear model assumptions met**

4. Run the ANOVA and the Tukey comparison.
```{r}
DMSO_type_III_ANOVA_results_log10 <- Anova(DMSO_comparison_model_log10, type = 3)
DMSO_type_III_ANOVA_results_log10
```

### Interpretation
- F-value = 0.06003
- p- value = 3.879e-05 ***
- DMSO treatment had a significant effect on cobalt concentration at 168 hours for wild type cell experiments.

5. Create a boxplot to illustrate this result. There is no need to annotate because there are only two boxes and the independent variable was significant. 
```{r}
theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", 
        size = 32, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
        size = 34, angle = 0, hjust = .5, vjust = 0, face = "bold"),
        axis.title.y = element_text(color = "grey20", 
        size = 34, angle = 90, face = "bold"),
        strip.text.x = element_text(size = 32, colour = "grey20", face = "bold"),
        strip.text.y = element_text(size = 32, colour = "grey20"),
        plot.title = element_text(size = 30, face = "bold", hjust = .5),
        legend.title=element_text(size=32, face = "bold"), 
        legend.text=element_text(size=32),
        panel.border = element_rect(colour = "black", fill=NA, size=1))

DMSO_boxplot <- DMSO_comparison_data %>%
                ggplot(aes(x=DMSO_treatment, y=log10(Cobalt_concentration_uM_corrected))) + 
                geom_boxplot() +
                geom_jitter(fill = "grey", colour = "black", width = 0.1, pch =21, stroke =1)+
                #ylim(0, 3000) +
                labs(title = "" ,
                y = "log10[Co(2+) concentration (µM)]\ncorrected for recoveries", 
                x = "DMSO treatment") +
                theme +
                theme(plot.margin = unit(c(1,1,1,1), "cm"),
                      axis.title.x = element_text(vjust = -2),
                      axis.title.y = element_text(vjust = 2),
                      axis.text.x = element_text(angle = 0))

DMSO_boxplot

#Save different image formats
ggsave("FigS20_cobalt_ICPMS_data_DMSO_20260202.jpg", dpi = 300, height=8, width=10, device=jpeg)
ggsave("FigS20_cobalt_ICPMS_data_DMSO_20260202.tiff", dpi = 300, height=8, width=10, device=tiff)
ggsave("FigS20_cobalt_ICPMS_data_DMSO_20260202.pdf", dpi = 300, height=8, width=10, device=cairo_pdf)
```
# Analysis for genetic basis of DET 

## Run Nanophase to assemble nanopore long read sequencing data on mutants

1. Make a directory to house the pre-assembled genomes and raw fasqt.gz files from the sequencing centre. Copy the files over to the target directories.
```{bash}
mkdir Pseudomonas_deletion_mutant_long_read_sequencing_data
mkdir Raw_fasta_files
mkdir Preassembled_genomes

#Copy pre-assembled genomes
scp daniel@134.117.183.97:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/FlowGenomics_data/Pca*.fasta ~/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Preassembled_genomes

#Copy raw fastq files
scp daniel@134.117.183.97:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/FlowGenomics_data/4074f4bc/fastqs/*.fastq.gz ~/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Raw_fasta_files
```

2. Start a screen, navigate to the proper directory, and activate nanophase to run the assembly in each raw fastq.gz file.
```{bash}
tmux new -s nanophase_run #you can change this name to whatever you want
cd ~/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Raw_fasta_files
conda activate nanophase

#Wild type
nanophase isolate -l Pca-WT.fastq.gz -t 40 -o ~/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_WT_nanophase_assembly

#∆phz deletion mutant
nanophase isolate -l Pca-phz.fastq.gz -t 40 -o ~/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_phz_nanophase_assembly

#∆hcn deletion mutant
nanophase isolate -l Pca-hcn.fastq.gz -t 40 -o ~/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_hcn_nanophase_assembly
```

3. Rename files (nanophase provides generic names). Export and copy over required files to computer.
```{bash}
cd /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_WT_nanophase_assembly/03-Polishing/Final-bins

mv bin.fasta Pca_WT_nanophase.fasta

mv nanophase.ont.genome.summary Pca_WT_nanophase.ont.genome.summary

cd /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_phz_nanophase_assembly/03-Polishing/Final-bins

mv bin.fasta Pca_phz_nanophase.fasta

mv nanophase.ont.genome.summary Pca_phz_nanophase.ont.genome.summary

cd /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_hcn_nanophase_assembly/03-Polishing/Final-bins

mv bin.fasta Pca_hcn_nanophase.fasta

mv nanophase.ont.genome.summary Pca_hcn_nanophase.ont.genome.summary

scp /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_WT_nanophase_assembly/03-Polishing/Pca_WT_nanophase.ont.genome.summary daniel@134.117.183.120:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/

scp /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_phz_nanophase_assembly/03-Polishing/Pca_phz_nanophase.ont.genome.summary daniel@134.117.183.120:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/

scp /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_hcn_nanophase_assembly/03-Polishing/Pca_hcn_nanophase.ont.genome.summary daniel@134.117.183.120:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/
```


## Re-run DRAM on the pre-assembled and nanophase assembled genomes

1. Copy all of the input .fasta files to a dedicated directory to facilitate running DRAM.
```{bash}
cd /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data

mkdir DRAM_analyses
mkdir DRAM_analyses/Input_genomes

cp /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Preassembled_genomes/*.fasta /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/DRAM_analyses/Input_genomes

cp /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_WT_nanophase_assembly/03-Polishing/Final-bins/*.fasta /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/DRAM_analyses/Input_genomes

cp /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_phz_nanophase_assembly/03-Polishing/Final-bins/*.fasta /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/DRAM_analyses/Input_genomes

cp /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/Pca_hcn_nanophase_assembly/03-Polishing/Final-bins/*.fasta /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/DRAM_analyses/Input_genomes
```

2. Create a screen to run DRAM in and activate the conda environment. Run DRAM with the --use_uniref flag.
```{bash}
cd /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/DRAM_analyses

tmux new -s Pseudomonas_DRAM
conda activate DRAM_conda
DRAM.py annotate -i 'Input_genomes/*.fasta' -o Pseudomonas_genome_DRAM_annotation_uniref90_20250916 --use_uniref --threads 40
```

3. Create the distillate to obtain all of the summary files.
```{bash}
DRAM.py distill -i Pseudomonas_genome_DRAM_annotation_uniref90_20250916/annotations.tsv -o Pseudomonas_genome_DRAM_annotation_uniref90_20250916/Pseudomonas_genome_DRAM_genome_summaries_20250916 --trna_path Pseudomonas_genome_DRAM_annotation_uniref90_20250916/trnas.tsv --rrna_path Pseudomonas_genome_DRAM_annotation_uniref90_20250916/rrnas.tsv
```

4. Copy the whole directory over to your computer.
```{bash}
scp -r /files/home/dgregoire/Pseudomonas_genome_analyses/Pseudomonas_deletion_mutant_long_read_sequencing_data/DRAM_analyses daniel@134.117.183.120:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/
```

## Verify long-read assemblies for phz and hcn genes.

1. Load the annotation file for all genomes output by DRAM.
```{r}
#Change the directory to load the data
setwd("/Users/daniel/Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/DRAM_analyses/Pseudomonas_genome_DRAM_annotation_uniref90_20250916")
Pseudomonas_long_read_genome_annotation <- fread("annotations.tsv", sep = "\t", header = "auto")
```

2. Separate each genome into a separate annotation data frame that includes the genome from the sequencing centre and the ones asembled with nanophase.
```{r}
Pca_WT_genomes <- Pseudomonas_long_read_genome_annotation %>% filter(fasta == "Pca-WT" | fasta == "Pca_WT_nanophase")
Pca_phz_genomes <- Pseudomonas_long_read_genome_annotation %>% filter(fasta == "Pca-phz" | fasta == "Pca_phz_nanophase")
Pca_hcn_genomes <- Pseudomonas_long_read_genome_annotation %>% filter(fasta == "Pca-hcn" | fasta == "Pca_hcn_nanophase")
```

3. Manually search the expected gene positions for the phz and hcn genes in Proksee (separate figures manually downloaded).

## Fig S22: Corrosion gene annotation

UniProt KB entries were manually collected by using the literature to identify genes directly involved in direct electron transfer in Geobacter metallireducens GS15, Geobacter sulfurreducens PCA (by way of gene deletion mutants using this as the parent strain), Methanosarcina acetivorans C2A (by way of gene deletion mutants using this as the parent strain), and Shewanella oneidensis MR-1. 

The UniProtKB accession numbers were then used to create a concatanated .fasta file that would be used as a custom annotation database with DRAM. This annotation run would apply to the Pseudomonas chlororaphis subspecieis aureofaciens genome obtained via long-read sequencing alongside the strains noted above to ensure that the genes that should be present in the parent genomes are indeed there. Genomes for the model organisms noted above were manually retrieved by searching for the organism names in KEGG and clicking through links to obtain the raw accession numbers for download via NCBI.

1. To create the gene database, use the UniProtKB website to do a batch download, copy and pasting the raw accession number collected in the associated spreadsheet that was used to collect these data. Download the associated metadata manually as well.

This provided *47 unique entries *which were exported as a .fasta file using the amino acid alphabet.

2. Create a dedicated directory to house all of the input genomes and the data base file. Copy over all source files required from your computer (code removed for privacy)
```{bash}
mkdir Model_corroder_genomes
scp daniel@134.117.183.226:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/Model_corroder_genomes/*.fna ~/Model_corroder_genomes

scp daniel@134.117.183.226:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/Model_corroder_genomes/*.fasta ~/Model_corroder_genomes

scp daniel@134.117.183.226:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/Model_corroder_genomes/Corrosion_gene_database.fasta /datastore/userdata/daniel/Custom_annotation_databases/

#Rename everything to .fasta extension
for f in *.fna; do
    mv "$f" "${f%.fna}.fasta"
done
```


3. Create a screen to run the process in the background and run DRAM calling uniref90 but also providing a path to the custom database. 
```{bash}
tmux new -s Corrosion_gene_annotation
conda activate DRAM_conda

DRAM.py annotate -i 'Model_corroder_genomes/*.fasta' -o Model_corroder_annotations_uniref90_20260108 --use_uniref --custom_fasta_loc /datastore/userdata/daniel/Custom_annotation_databases/Corrosion_gene_database.fasta --custom_db_name Corrosion_genes --threads 4
```

Started 2026-01-08 at 16:15

Got an error message at 20:45: 

/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/database_handler.py:218: UserWarning: No descriptions were found for your id's. Does this GT2_Glycos_transf_2 look like an id from dbcan_description           
  warnings.warn(
/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/database_handler.py:218: UserWarning: No descriptions were found for your id's. Does this GT2_Glyco_tranf_2_3 look like an id from dbcan_description           
  warnings.warn(
2026-01-08 18:05:53,932 - Getting hits from Corrosion_genes
2026-01-08 18:05:53,933 - Getting forward best hits from Corrosion_genes
2026-01-08 18:06:08,313 - Getting reverse best hits from Corrosion_genes
2026-01-08 18:06:08,549 - Getting descriptions of hits from Corrosion_genes
Traceback (most recent call last):
  File "/datastore/tools/gregoire/envs/DRAM_conda/bin/DRAM.py", line 211, in <module>
    args.func(**args_dict)
  File "/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/annotate_bins.py", line 1531, in annotate_bins                                                                                                         
    all_annotations = annotate_fastas(
  File "/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/annotate_bins.py", line 1410, in annotate_fastas                                                                                                       
    annotate_fasta(
  File "/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/annotate_bins.py", line 1255, in annotate_fasta                                                                                                        
    annotations = annotate_orfs(
  File "/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/annotate_bins.py", line 1159, in annotate_orfs                                                                                                         
    do_blast_style_search(
  File "/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/annotate_bins.py", line 859, in do_blast_style_search                                                                                                  
    hits = formater(hits, header_dict)
  File "/datastore/tools/gregoire/envs/DRAM_conda/lib/python3.10/site-packages/mag_annotator/annotate_bins.py", line 181, in get_basic_description                                                                                                  
    header = header_dict[hit]
KeyError: 'Q8EBS0'

May need to clean up the leading text in the headers to facilitate look up.

4. Tidy up headers in a new file and try again.
```{bash}
#Clean up sp and tr leading the headers
sed -E 's/^>(sp|tr)\|/>/' Corrosion_gene_database.fasta > Corrosion_gene_database_cleaned.fasta

DRAM.py annotate -i 'Model_corroder_genomes/*.fasta' -o Model_corroder_annotations_uniref90_20260108 --use_uniref --custom_fasta_loc /datastore/userdata/daniel/Custom_annotation_databases/Corrosion_gene_database_cleaned.fasta --custom_db_name Corrosion_genes --threads 4
```

5. Run the distillation command for an additional summary (this file won't be used in the final analyses, but is helpful for other applications).
```{bash}
DRAM.py distill -i ~/Model_corroder_annotations_uniref90_20260108/annotations.tsv -o ~/Model_corroder_annotations_uniref90_20260108/genome_summaries --trna_path ~/Model_corroder_annotations_uniref90_20260108/trnas.tsv --rrna_path ~/Model_corroder_annotations_uniref90_20260108/rrnas.tsv
```

6. Copy the files over to your computer to begin inspecting in R.
```{bash}
scp -r ~/Model_corroder_annotations_uniref90_20260108 daniel@134.117.183.34:Dropbox/Carleton/Research_program/E_waste_biogeochemistry/Corrosion_project/Pseudomonas_genome_annotation/Model_corroder_genomes/
```

7. Import the annotation data into R and trim the file to the relevant genomes and columns of interest.
```{r}
#Load the data frame from scratch to test
Pseudomonas_annotations <- fread("Supporting_data_1A_corrosion_gene_annotations.tsv", sep = "\t", header = "auto") %>% 
                           filter(fasta == "Pseudomonas_chlororaphis_aureofaciens_WT") 

write.table(Pseudomonas_annotations, file = "Supporting_data_1_Pseudomonas_chlororaphis_subsp_aureofaciens_genome_annotation_20260129.txt", sep = "\t", 
            row.names = FALSE, col.names = TRUE) 

Corrosion_gene_annotations <- fread("Supporting_data_1A_corrosion_gene_annotations.tsv", sep = "\t", header = "auto") %>% 
                              filter(!fasta == "Cupriavidus_metallidurans_CH34") %>%
                              dplyr::select(c(fasta:uniref_eVal, Corrosion_genes_id:heme_regulatory_motif_count)) %>%
                              filter(!Corrosion_genes_id == "") %>%
                              mutate(Gene_name = str_extract(Corrosion_genes_hit, "(?<=GN=).*?(?=\\sPE=)")) %>%
                              relocate(Gene_name, .after = uniref_eVal)

write.table(Corrosion_gene_annotations, file = "Supporting_data_1A_Corrosion_genome_annotations_20260129.txt", sep = "\t", 
            row.names = FALSE, col.names = TRUE)

#Quickly inspect density plots for bit score and percent identity

ggplot(Corrosion_gene_annotations, aes(x = Corrosion_genes_identity)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  labs(
    title = "Density Plot of Corrosion_genes_identity",
    x = "Corrosion genes identity",
    y = "Density"
  ) +
  theme_minimal()

ggplot(Corrosion_gene_annotations, aes(x = log10(Corrosion_genes_bitScore))) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  labs(
    title = "Density Plot of Corrosion_genes_identity",
    x = "Corrosion genes identity",
    y = "Density"
  ) +
  theme_minimal()
```
This looks like setting a simple bitScore cutoff of 100 would get rid of a lot of the hits that align over a small portion of the gene (i.e., < 50%)

8. Filter the data to group by strain, and then corrosion gene id, and then take the highest similarity scores for each one.
```{r}
Corrosion_gene_annotations_max_similarity <- Corrosion_gene_annotations %>%
                                             group_by(fasta, Corrosion_genes_id) %>%
                                             slice_max(Corrosion_genes_identity, n = 1, with_ties = FALSE) %>%
                                             ungroup() #87 observations

ggplot(Corrosion_gene_annotations_max_similarity, aes(x = Corrosion_genes_identity)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  labs(
    title = "Density Plot of Corrosion_genes_identity",
    x = "Corrosion genes identity",
    y = "Density"
  ) +
  theme_minimal()

ggplot(Corrosion_gene_annotations_max_similarity, aes(x = log10(Corrosion_genes_bitScore))) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  labs(
    title = "Density Plot of Corrosion_genes_identity",
    x = "Corrosion genes identity",
    y = "Density"
  ) +
  theme_minimal()
```

9. Plot a heat map that uses a custom order for the genomes and shows the percent homology.
```{r}
custom_order <- c(
  "*Shewanella oneidensis MR1*",
  "*Methanosarcina acetivorans C2A*",
  "*Geobacter sulfurreducens PCA*",
  "*Geobacter metallireducens GS15*",
  "*Pseudomonas chlororaphis aureofaciens WT*"
  )


Corrosion_gene_heatmap <- Corrosion_gene_annotations_max_similarity %>%
                          filter(!Gene_name == "recA") %>%
                          mutate(Corrosion_genes_identity_percent = Corrosion_genes_identity*100) %>%
                          mutate(
                          Genome = paste0(
                            "*",
                            gsub("_", " ", fasta),
                            "*"
                          )
                        ) %>%
                        mutate(Genome_ordered = factor(Genome, levels = custom_order)) %>%
                          ggplot(aes(x = Gene_name, y = Genome_ordered, fill = Corrosion_genes_identity_percent)) +
                          geom_tile(color = "grey80") +
                          scale_fill_viridis_c(
                            option = "viridis",
                            na.value = "white"
                          ) +
                          theme_minimal() +
                          theme(
                            panel.grid = element_blank(),            # remove grid lines
                            axis.text.x = element_text(
                              angle = 90,
                              vjust = 0.5,
                              hjust = 1,
                              size = 20
                            ),
                            axis.text.y = element_markdown(size = 20),
                            legend.text = element_text(size = 22),
                            legend.title = element_text(size = 24, margin = margin(b = 20))
                          ) +
                          labs(
                            x = NULL,
                            y = NULL,
                            fill = "% similarity"
                          )

Corrosion_gene_heatmap
ggsave("FigS22_Corrosion_gene_annotation_heatmap_20260202.jpg", 
       dpi = 300, height=6, width=22, device= jpeg)
ggsave("FigS22_Corrosion_gene_annotation_heatmap_20260202.tiff", 
       dpi = 300, height=6, width=22, device= tiff)
ggsave("FigS22_Corrosion_gene_annotation_heatmap_20260202.pdf", 
       dpi = 300, height=6, width=22, device= cairo_pdf)
```


# Figure S31: Stacked barplots for phenazine measurements via LC-HRMS

1. Load the formatted LC-HRMS data.
```{r}
LC_MS_data_combined_labelled_long <- fread("LC_MS_data_combined_20250419.txt", sep = "\t", header = "auto")
```

2. Plot the stacked bar graphs using variables with a custom order.
```{r}
barplot_theme<-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
                 panel.background = element_blank(), axis.line = element_line(colour = "black"),
                 axis.text.x = element_text(color = "black", size = 32),
                 axis.text.y = element_text(color = "black", size = 32),
                 axis.title.x = element_text(color = "black", size = 34, vjust = -3, face = "bold"),
                 axis.title.y = element_text(color = "black", size = 34, vjust = 6, face = "bold"),
                 strip.text.x = element_text(color = "black", size = 32, face = "bold"),
                 strip.text.y = element_text(color = "black", size = 32, face = "bold"),
                 legend.title=element_text(size=40, vjust = 2, face = "bold"), 
                 legend.text=element_text(size=34),
                 legend.key = element_rect(fill = "white", color = NA),
                 plot.title = element_text(size = 68, face = "bold", hjust = 0.5),
                 panel.border = element_rect(colour = "black", fill=NA, size=1),
                 plot.margin = unit(c(2, 2, 2, 2), "cm"))

LC_MS_data_combined_labelled_long$Abiotic_vs_biotic_ordered <- factor(LC_MS_data_combined_labelled_long$Abiotic_vs_biotic, 
                                                                      levels = c("Wild type",
                                                                            "Spent medium",
                                                                            "∆phz"))

LC_MS_data_combined_labelled_long$Metabolite_ordered <- factor(LC_MS_data_combined_labelled_long$Metabolite, 
                                                                      levels = c("2-OHPCA",
                                                                            "PCA",
                                                                            "2-OHPHZ"))


Phenazine_barplot <-   LC_MS_data_combined_labelled_long %>%
                       group_by(Comment) %>%
                       mutate(Mean_concentration_ug_mL = mean(Concentration_ug_mL)) %>%
                       ggplot(aes(x= Time, y = Concentration_ug_mL, fill = Metabolite_ordered)) +
                       geom_bar(aes(group = Replicate), position="stack", stat="identity") +
                       facet_nested(Replicate~Wire_treatment + Abiotic_vs_biotic_ordered)+
                       scale_x_continuous(breaks = seq(0, 168, by = 24)) +
                       scale_fill_manual(name = "Metabolite", 
                                         values = c("2-OHPCA" = "#FA1F1E",   # Red
                                                    "PCA" = "#EBA602",       # Gold
                                                    "2-OHPHZ" = "#027AB0"    # Blue
                                                    ))+
                       #scale_fill_brewer(name = "Metabolite", palette = "YlOrRd")+ 
                       ylim(0,15) +
                       theme(axis.text.x=element_text(angle=0, hjust=0.5),
                       axis.title.y = element_text(color = "black", size = 24, vjust = 0))+
                       labs(x = "Time elapsed (hours)" , y = "Concentration (µg/mL)") +
                       barplot_theme + 
                       theme(legend.key.size = unit(2, 'cm'))

Phenazine_barplot

Phenazine_barplot_angled <- Phenazine_barplot + theme(axis.text.x = element_text(angle = 45, vjust = 0.5))

#Save in different image formats
ggsave("FigS31_phenazine_LCMS_data_stacked_barplot_faceted_20260202.jpg", dpi = 300, height=12, width=40, device=jpeg)
ggsave("FigS31_phenazine_LCMS_data_stacked_barplot_faceted_20260202.tiff", dpi = 300, height=12, width=40, device=tiff)
ggsave("FigS31_phenazine_LCMS_data_stacked_barplot_faceted_20260202.pdf", dpi = 300, height=12, width=40, device=cairo_pdf)
```